### 2023-12-03 data/sleep and HBP.Rmd
update

---
title: "Preparing the NHANES Data for the Replication of: Association between sleep duration on workdays and blood pressure in non‑overweight/obese population in NHANES: a public database research"
subtitle: "By Yingjie Su and published in Nature Scientific Reports"
date: "  `r format(Sys.time(), '%d %B, %Y')` "
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{fvextra}
- \usepackage{float}
- \usepackage{wrapfig}
- \usepackage{amsmath}
- \usepackage{threeparttable} %used to fix table with notes
- \usepackage{float}
- \usepackage{graphicx}
- \usepackage{microtype}
- \usepackage{setspace}
- \usepackage[font=singlespacing]{caption} #can change font here for captions here!!
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines, commandchars=\\\{\}}
- \singlespacing
fontsize: 10pt
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    toc: false
    toc_depth: 4
    number_sections: false
    keep_md: false
link-citations: yes
colorlinks: yes
linkcolor: blue
urlcolor: blue
#bibliography: refs.bib
bibliostyle: apalike
editor_options: 
  markdown: 
    wrap: 72
---

## Introduction
In this sample, we are interested in replicating Table 1 in the paper by Su et al. (2022) using data from the NHANES database. We will start by first loading some of the packages we will be using in our data extraction as well as replication of the table.

```{r warning=FALSE, message=FALSE}
library(nhanesA)
library(knitr)
library(kableExtra)
library(tidyverse)
library(gtsummary)
```

We will first extract the demographic variables from the 2015-2016 cycle. To check how the data table for this specific cycle has been coded on the CDC website, we can run the following chunk which queries the database and returns the associated name of the data group associated with the 2015 - 2016 survey cycle using the `nhanesTables()` function. For more info, you can type `?nhanesTables()` in the console to get a more detailed review of the function and how to use it.

```{r}
## demographic data 2015/2016
#race, gender, age, education level, marital status, income, religion, occupation, home ownership, 
kable(nhanesTables(data_group = 'DEMO', year = "2016"))
```

We now know that the demographic variables we are interested in acquiring are stored within the `DEMO_I` data group. We can then explicitly query the database for all demographic variables for this survey cycle using the following chunk. From the results, we are interested in the two-year survey weights for the cycle, `SDMVPSU` --	Masked variance unit pseudo-PSU variable for variance estimation, `SDMVSTRA`	-- Masked variance unit pseudo-stratum variable for variance estimation, all of which we need to work with the complex survey design aspect of the survey. We also include `SEQN` which unquely identifies a respondent to be able to link the files together for each individual. In addition, we will extract variables related to race, gender, age, education level, marital status, the number of reported children aged below 5 years, and income range. We have added education level, marital status, and income range while thinking about the possibility of confounding by these variables.

```{r}
## get the descriptions of all the demographic variables in nhanes 2015/2016
kable(nhanesTableVars(data_group = "DEMO", nh_table = "DEMO_I"), format = "latex")%>% 
  kable_styling(latex_options = c("scale_down"))
```

## Extracting NHANES 2015/16 cycle

Here, we extract the variables of interest in the NHANES dataset for the 2015/2016 cycle. All variables are based on the reports in the paper in the methods section. While there might be an easier way to work with these data, the variables here are extracted step-by-step for each survey cycle. The CDC variable search tool can be helpful while thinking about getting information on what variables are available (https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx).

```{r warning=FALSE, message=FALSE}
## select all demographic variables that might be of interest in the study
demo <- nhanes("DEMO_I")

## subset to only vars of interest
demo1 <- demo[,c("DMDCITZN", # citizenship status
                "DMDEDUC2", # highest grade/education level
                "DMDHHSZA", # children aged < 5 yrs
                "RIDRETH3", # reported race
                "SDMVPSU",  # unit psedudo psu
                "SDMVSTRA", # pseudo-stratum vriable
                "RIAGENDR", # gender of the participant
                "RIDAGEYR", # age in yrs
                "INDFMIN2", # total family income range
                "DMDMARTL",  # marital status
                "SEQN",     # resp. seq. number
                "WTINT2YR" # full two-year interview weights
                )]
demo_vars <- names(demo1)
demo2 <- nhanesTranslate("DEMO_I", demo_vars, data = demo1)

## alcohol use
alcohol <- nhanes("ALQ_I")
alcohol1 <- alcohol[, c("SEQN", "ALQ120Q")]
alc_var <- names(alcohol1)
alcohol2 <- nhanesTranslate("ALQ_I", alc_var, data = alcohol1)

## smoking: do you smoke now?
sm <- nhanes("SMQ_I")
sm1 <- sm[c("SEQN", "SMQ040")]
sm_vars <- names(sm1)
sm2 <- nhanesTranslate("SMQ_I", sm_vars, data = sm1)

## diabetes: have you ever been told you have diabetes?
db <- nhanes("DIQ_I")
db1 <- db[c("SEQN", "DIQ010")]
db_vars <- names(db1)
db2 <- nhanesTranslate("DIQ_I", db_vars, data = db1)

## hypertension:
bp <- nhanes("BPQ_I") %>% dplyr::select(SEQN, BPQ020)
bp_vars <- names(bp)
bp2 <-nhanesTranslate("BPQ_I", bp_vars, data = bp)

## use hypertensive medication (?) included in the exclusion criteria but not mentioned in the list of variables
hbp_med <- nhanes("BPQ_I")
hbp_med1 <- hbp_med[c("SEQN", "BPQ050A")]
hbp_medvar <- names(hbp_med1)
hbp_med2 <- nhanesTranslate("BPQ_I", hbp_medvar, data = hbp_med1)

## snort/stop breathing
snort <- nhanes("SLQ_I")
snort1 <- snort[c("SEQN", "SLQ040")]
snort_vars <- names(snort1)
snort2 <- nhanesTranslate("SLQ_I", snort_vars, data = snort1)

## creatinine
creat <- nhanes("ALB_CR_I")
kable(nhanesTableVars(data_group = "LAB", nh_table = "ALB_CR_I"))
creat1 <- creat[c("SEQN", "URXCRS" # creatinine (umol/L)
                )]
creat_var <- names(creat1)
creat2 <- nhanesTranslate("ALB_CR_I", creat_var, data = creat1)

## albumin (g/L)
albu <- nhanes("BIOPRO_I")
albu1 <- albu[c("SEQN", "LBDSALSI", # albumin
                "LBXSASSI" # AST (U/L)
                )]
alb_vars <- names(albu1)
albu2 <- nhanesTranslate("BIOPRO_I", alb_vars, data = albu1)

## hemoglobin (g/dL)
hemo <- nhanes("CBC_I")
hemo1 <- hemo[c("SEQN", "LBXHGB")]
hemo_var <-names(hemo1)
hemo2 <- nhanesTranslate("CBC_I", hemo_var, data = hemo1)

## total cholestrol (TC mmol/L)
tc <- nhanes("TCHOL_I")
tc1 <- tc[c("SEQN", "LBDTCSI")]
tc_var <- names (tc1)
tc2 <- nhanesTranslate("TCHOL_I", tc_var, data = tc1)

## blood pressure (take three and average to get DBP in mmHg)
dbp <- nhanes("BPX_I")
dbp1 <- dbp[c("SEQN", "BPXDI1", "BPXDI2", "BPXDI3", "BPXDI4", # dbp
              "BPXSY1", "BPXSY2", "BPXSY3", "BPXSY4" # sbp
              )] # 1st, 2nd, and 3rd measurements
dbp_vars <- names(dbp1)
dbp2 <- nhanesTranslate("BPX_I", dbp_vars, data = dbp1)

## high density lipoprotein (HDL mmol/L)
hdl <- nhanes("HDL_I")
hdl1 <- hdl[c("SEQN", "LBDHDDSI")]
hdl_vars <- names(hdl1)
hdl2 <- nhanesTranslate("HDL_I", hdl_vars, data = hdl1)

## BMI (kg/m^2)
bmi <- nhanes("BMX_I")
bmi1 <- bmi[c("SEQN", "BMXBMI")]
bmi_var <- names(bmi1)
bmi2 <- nhanesTranslate("BMX_I", bmi_var, data = bmi1)

## sleep variable (number of hrs of sleep during weekdays or workdays) hrs
sleep <- nhanes("SLQ_I")
sleep1 <- sleep[c("SEQN", "SLD012")]
sleep_var <- names(sleep1)
sleep2 <- nhanesTranslate("SLQ_I", sleep_var, data = sleep1)

## combine all the files into one data frame
nhanes2015 <- plyr::join_all(list(sleep2, bmi2, hdl2, dbp2, tc2, hemo2, albu2, creat2, snort2, bp2, hbp_med2,  db2, sm2, alcohol2, demo2), by = "SEQN", type = "full")
dim(nhanes2015)
## save file
save(nhanes2015, file = "../Sleep and HBP/nhanes2015.RData")
```

We can then check the dimensions of the data we downloaded like in the chunk and note that we have 9971 observations on 34 variables. We have saved these data in our working directory for merging with the 2017/2018 cycle of the NHANES data which was also used in combination to the already downloaded data.

```{r}
dim(nhanes2015)
```

## Extracting NHANES 2017/18 Cycle Data

Here we extract the same variables for the 2017/2018 cycle and save the individuals and merged data files separately for analysis. We begin by querying the data to see its abbreviation in the 2017/2018 series. We are only interested in the same variables as the ones we have downloaded and therefore only query for these using the steps as seen above.

```{r}
## clear the workspace
rm(list = ls())
## NHANES 2017/2018 cycle has abbreviation J.
kable(nhanesTables(data_group = 'DEMO', year = "2018"))
```

Now that we know the abbreviation used in the 2017/18 NHANES series, we can proceed to get the variables defined in the previous section. For the demographics, the variables available can be listed like so:

```{r}
## select all demographic variables that might be of interest in the study
demo <- nhanes("DEMO_J")
kable(nhanesTableVars(data_group = "DEMO", nh_table = "DEMO_J"), format = "latex") %>% 
  kable_styling(latex_options = c("scale_down"))
```

```{r warning=FALSE, message=FALSE}
## subset to only vars of interest
demo1 <- demo[,c("DMDCITZN", # citizenship status
                "DMDEDUC2", # highest grade/education level
                "DMDHHSZA", # children aged < 5 yrs
                "RIDRETH3", # reported race
                "SDMVPSU",  # unit psedudo psu
                "SDMVSTRA", # pseudo-stratum vriable
                "RIAGENDR", # gender of the participant
                "RIDAGEYR", # age in yrs
                "INDFMIN2", # total family income range
                "DMDMARTL",  # marital status
                "SEQN",     # resp. seq. number
                "WTINT2YR" # full two-year interview weights
                )]
demo_vars <- names(demo1)
demo2 <- nhanesTranslate("DEMO_J", demo_vars, data = demo1)

## alcohol use
alcohol <- nhanes("ALQ_J")
alcohol1 <- alcohol[, c("SEQN", "ALQ121")]
alc_var <- names(alcohol1)
alcohol2 <- nhanesTranslate("ALQ_J", alc_var, data = alcohol1)

## smoking: do you smoke now?
sm <- nhanes("SMQ_J")
sm1 <- sm[c("SEQN", "SMQ040")]
sm_vars <- names(sm1)
sm2 <- nhanesTranslate("SMQ_J", sm_vars, data = sm1)

## diabetes: have you ever been told you have diabetes?
db <- nhanes("DIQ_J")
db1 <- db[c("SEQN", "DIQ010")]
db_vars <- names(db1)
db2 <- nhanesTranslate("DIQ_J", db_vars, data = db1)

## hypertension:
bp <- nhanes("BPQ_J") %>% dplyr::select(SEQN, BPQ020)
bp_vars <- names(bp)
bp2 <-nhanesTranslate("BPQ_J", bp_vars, data = bp)

## use hypertensive medication (?) included in the exclusion criteria but not mentioned in the list of variables
hbp_med <- nhanes("BPQ_J")
hbp_med1 <- hbp_med[c("SEQN", "BPQ050A")]
hbp_medvar <- names(hbp_med1)
hbp_med2 <- nhanesTranslate("BPQ_J", hbp_medvar, data = hbp_med1)

## snort/stop breathing
snort <- nhanes("SLQ_J")
snort1 <- snort[c("SEQN", "SLQ040")]
snort_vars <- names(snort1)
snort2 <- nhanesTranslate("SLQ_J", snort_vars, data = snort1)

## creatinine
creat <- nhanes("ALB_CR_J")
kable(nhanesTableVars(data_group = "LAB", nh_table = "ALB_CR_J"))
creat1 <- creat[c("SEQN", "URXCRS" # creatinine (umol/L)
                )]
creat_var <- names(creat1)
creat2 <- nhanesTranslate("ALB_CR_J", creat_var, data = creat1)

## albumin (g/L)
albu <- nhanes("BIOPRO_J")
albu1 <- albu[c("SEQN", "LBDSALSI", # albumin
                "LBXSASSI" # AST (U/L)
                )]
alb_vars <- names(albu1)
albu2 <- nhanesTranslate("BIOPRO_J", alb_vars, data = albu1)

## hemoglobin (g/dL)
hemo <- nhanes("CBC_J")
hemo1 <- hemo[c("SEQN", "LBXHGB")]
hemo_var <-names(hemo1)
hemo2 <- nhanesTranslate("CBC_J", hemo_var, data = hemo1)

## total cholestrol (TC mmol/L)
tc <- nhanes("TCHOL_J")
tc1 <- tc[c("SEQN", "LBDTCSI")]
tc_var <- names (tc1)
tc2 <- nhanesTranslate("TCHOL_J", tc_var, data = tc1)

## blood pressure (take three and average to get DBP in mmHg)
dbp <- nhanes("BPX_J")
dbp1 <- dbp[c("SEQN", "BPXDI1", "BPXDI2", "BPXDI3", "BPXDI4",# dbp
              "BPXSY1", "BPXSY2", "BPXSY3", "BPXSY4" # sbp
              )] # 1st, 2nd, and 3rd measurements
dbp_vars <- names(dbp1)
dbp2 <- nhanesTranslate("BPX_J", dbp_vars, data = dbp1)

## high density lipoprotein (HDL mmol/L)
hdl <- nhanes("HDL_J")
hdl1 <- hdl[c("SEQN", "LBDHDDSI")]
hdl_vars <- names(hdl1)
hdl2 <- nhanesTranslate("HDL_J", hdl_vars, data = hdl1)

## BMI (kg/m^2)
bmi <- nhanes("BMX_J")
bmi1 <- bmi[c("SEQN", "BMXBMI")]
bmi_var <- names(bmi1)
bmi2 <- nhanesTranslate("BMX_J", bmi_var, data = bmi1)

## sleep variable (number of hrs of sleep during weekdays or workdays) hrs
sleep <- nhanes("SLQ_J")
sleep1 <- sleep[c("SEQN", "SLD012")]
sleep_var <- names(sleep1)
sleep2 <- nhanesTranslate("SLQ_J", sleep_var, data = sleep1)

## combine all the files into one data frame
nhanes2017 <- plyr::join_all(list(sleep2, bmi2, hdl2, dbp2, tc2, hemo2, albu2, creat2, snort2, bp2, hbp_med2,  db2, sm2, alcohol2, demo2), by = "SEQN", type = "full")
dim(nhanes2017)

## save the file to local computer
save(nhanes2017, file = "../Sleep and HBP/nhanes2017.RData")
```

The downloaded dataset has 9254 rows and 33 columns. We can then merge the data frames together after correcting any naming conflicts and attaching a survey cycle indicator. This can be useful when thinking about performing the same analysis stratified by the survey cycle to check, for example, if the same effects presented in the paper hold.

```{r}
rm(list = ls())
# load the 2015 cycle and add an indicator of survey cycle
load("../Sleep and HBP/nhanes2015.RData")
load("../Sleep and HBP/nhanes2017.RData")
nhanes2015$cycle <- 0
nhanes2017$cycle <- 1

## check what names do not match between the two surveys
setdiff(names(nhanes2015), names(nhanes2017))

## only column ALQ120Q is not the same and we can change that label in the 2015 cycle to match and merge
nhanes2015 <- nhanes2015 %>% dplyr::rename(ALQ121 = ALQ120Q)

## we can confirm that the names are now identical
identical(names(nhanes2015), names(nhanes2017))

## append the two data files together
df <- rbind(nhanes2015, nhanes2017)

## get the dimension of combined file
dim(df)
```

Finally we have a working data set with more variables than in the original file so we can play around with it and we have the same sample size as in the paper. We can save the merged file together.

```{r}
## save all the three files
#save(df, file = "../Sleep and HBP/df.RData")
```

## Data preprocessing

Now that we have a dataset together, we can go ahead and clean the variables following the preprocessing steps employed in the paper. We begin by re-coding all the factor variables into the formats presented in the table. We first rename the variables into more reasonable identities. One thing about the variables that the paper does not mention is the question used to define use of hypertensive medication.

```{r}
## rename variable names
dt <- df %>% 
  dplyr::rename("seq_no" = SEQN, "psu" = SDMVPSU, "strata" = SDMVSTRA, "gender" = RIAGENDR, "age_yr" = RIDAGEYR, "marital_status" = DMDMARTL, "weights" = WTINT2YR, "income_range" = INDFMIN2, "race" = RIDRETH3, "children <5" = DMDHHSZA, "educ_level" = DMDEDUC2, "citizenship_status" = DMDCITZN, "sleep" = SLD012, "bmi" = BMXBMI, "alcohol" = ALQ121, "hdl" = LBDHDDSI, "albumin" = LBDSALSI, "diabetes" = DIQ010, "hypertension" = BPQ020, "creatinine" = URXCRS, "snort" = SLQ040, "total_chol" = LBDTCSI,"hemoglobin" = LBXHGB, "AST" = LBXSASSI, "smoke" = SMQ040, "hyper_meds" = BPQ050A)

## drop missing sleep values
dt <- dt %>% tidyr::drop_na(sleep)
dim(dt)

## combine the systolic and diastolic blood pressure by averaging the four measurements available
dt$dbp <- rowMeans(subset(dt, select = c(BPXDI1, BPXDI2, BPXDI3, BPXDI4)), na.rm = TRUE)
dt$sbp <- rowMeans(subset(dt, select = c(BPXSY1, BPXSY2, BPXSY3, BPXSY4)), na.rm = TRUE)

## we can then delete the variables related to dbp and sbp from the data set and drop missing values for either dbp or sbp
dt <- dt %>% dplyr::select(-c(BPXDI1, BPXDI2, BPXDI3, BPXDI4, BPXSY1, BPXSY2, BPXSY3, BPXSY4)) %>% 
  tidyr::drop_na(dbp, sbp)
dim(dt)

## then we exclude individuals on antihypertension meds
dt = dt %>% dplyr::filter(is.na(hyper_meds)|hyper_meds == "No")
dim(dt)
## we also need to exclude individuals with missing BMI and then finally subset to only underweight or obese
dt <- dt %>% tidyr::drop_na(bmi) %>% 
  dplyr::filter(bmi < 25)
dim(dt)

## we can save this files for use later
#save(dt, file = "../Sleep and HBP/cleaned_df.RData")
```

Now that we have filtered the data to have the same sample size as used in the paper and also followed the prescribed exclusion criteria, we can go on and clean up the categories of the categorical variables so that they match what has been presented in the paper or are at least reasonable enough. Not the categorizations for the alcohol variable differ from what was presented in the paper since the variable was differently measured across survey cycles.

```{r include=FALSE}
## sleep duration
dt$sleep2 <- as.factor(ifelse(dt$sleep < 6, "<6hrs", 
                              ifelse((dt$sleep >= 6) & (dt$sleep < 8), "6-8hrs", ">8hrs")))

#race
table(dt$race)
dt$race2 <- as.factor(ifelse(dt$race == "Mexican American", "Mexican American",
                            ifelse(dt$race == "Non-Hispanic White", "White",
                                   ifelse(dt$race == "Non-Hispanic Black", "Black", "Other"))))
## smoking
dt$smoke2 <- as.factor(ifelse(dt$smoke == "Every day" | dt$smoke == "Some days", "Smoking",
                               ifelse(dt$smoke == "Not at all", "Not smoking", "Not recorded")))

## diabetes
table(dt$diabetes)

## hypertension
table(dt$hypertension)

## snort
table(dt$snort)
dt$snort2 <- as.factor(ifelse(dt$snort == "Never", "No",
                             ifelse(dt$snort == "Rarely - 1-2 nights a week"| dt$snort == "Occasionally - 3-4 nights a week" | dt$snort == "Frequently - 5 or more nights a", "Yes", "Not recorded")))

## gender
table(dt$gender)

## alcohol consumption
table(dt$alcohol, useNA = "ifany")
dt[dt == 999] <- NA ## convert missing to NA
## assume individuals with 0 drinks have never drank and all others with info have drank, and then we have NA
dt$alcohol2 <- as.factor(ifelse(dt$alcohol == 0, "No drinking", 
                                ifelse(is.na(dt$alcohol), "Not recorded", "Drinking")))
table(dt$alcohol2)
```

## Table 1 Replication

The table below shows our sample replication. Even though weights are not used in this table, we perform pretty well. The p-values show effects in the same direction as those reported in the paper except for age, alcohol consumption, sleep and smoking status. Overall, replicating this paper was relatively difficult considering no information is presented in the paper about variables they used; given that some variables occur multiply across the cycles. This could have had the effect of reducing the quality of these replication. An extension could consider adding a survey cycle indicator, including some of the additional confounding variables into the fitted model and seeing how the results change.

```{r}
## select the variables in Table 1
dt2 <- dt %>% dplyr::select(age_yr, albumin, creatinine, sbp, dbp, hemoglobin, total_chol, AST, hdl, bmi, alcohol2, diabetes, hypertension, snort2, smoke2, race2, sleep2, gender)
#str(dt2)

table1 <- dt2 %>% gtsummary::tbl_summary(by = gender,
                                 statistic = list(all_continuous() ~ "{mean} ({sd})", 
                                          all_categorical() ~ "{p}"),
                         percent = "column",
                         missing = "no") %>% 
  bold_labels() %>% 
  add_overall() %>% 
  add_p() %>% 
  as_kable(format = "pipe", caption = "Baseline participant characteristics", digits = 2) #%>% 
  #kable_styling(latex_options = "scale_down")
table1
```

### 2023-12-03 manuscript/paper.Rmd
update

---
title: "Preparing the NHANES Data for the Replication of: Association between sleep duration on workdays and blood pressure in non‑overweight/obese population in NHANES: a public database research"
thanks: "Replication files are available on the author's Github account (http://github.com/svmiller/svm-r-markdown-templates). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
subtitle: "Short title goes here!"
author:
- Amos Okutse, Brown University
- Second Author, Another Uni
- Graduate Student, Graduate School
abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec sit amet libero justo. Pellentesque eget nibh ex. Aliquam tincidunt egestas lectus id ullamcorper. Proin tellus orci, posuere sed cursus at, bibendum ac odio. Nam consequat non ante eget aliquam. Nulla facilisis tincidunt elit. Nunc hendrerit pellentesque quam, eu imperdiet ipsum porttitor ut. Interdum et malesuada fames ac ante ipsum primis in faucibus. Suspendisse potenti. Duis vitae nibh mauris. Duis nec sem sit amet ante dictum mattis. Suspendisse diam velit, maximus eget commodo at, faucibus et nisi. Ut a pellentesque eros, sit amet suscipit eros. Nunc tincidunt quis risus suscipit vestibulum. Quisque eu fringilla massa."
keywords: "pandoc, r markdown, knitr"
date: "  `r format(Sys.time(), '%d %B, %Y')` "
geometry: margin=1in
#mainfont: cochineal
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage{fvextra}
- \usepackage{float}
- \usepackage{wrapfig}
- \usepackage{amsmath}
- \usepackage{threeparttable} %used to fix table with notes
- \usepackage{float}
- \usepackage{graphicx}
- \usepackage{microtype}
- \usepackage{setspace}
- \usepackage[font=singlespacing]{caption} #can change font here for captions here!!
- \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines, commandchars=\\\{\}}
- \singlespacing
fontsize: 10pt
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    dev: cairo_pdf
    toc: false
    toc_depth: 4
    number_sections: false
    keep_md: false
link-citations: yes
colorlinks: yes
linkcolor: blue
urlcolor: blue
#bibliography: refs.bib
bibliostyle: apalike
editor_options: 
  markdown: 
    wrap: 72
---

# Introduction

- importance of reproducibility
- As a case on point ...?
- Underscores the importance of reproducibility within the frame of statistical analysis. We use the study by to ...
The current study builds on the paper by .. and starts in the manner of a replication study and extends the analyses therein. We focus on statistical analyses of these data and compare models based on varied specification of functional forms of the covariates.

# Methods

## Study population

Data utilized in this paper is sourced from the National Health and Nutrition Examination Survey (NHANES), a comprehensive nationwide survey administered by the National Center for Health Statistics (NCHS) via the Centers for Disease Control and Prevention (CDC). The survey assesses the health and nutrition of the entire non-institutionalized US population, spanning all ages and residing in all 50 states as well as Washington D.C. As such, the survey provides a cross-sectional view of a representative sample of the US population. Further information about NHANES can be found at www.cdc.gov/nchs/nhanes.

## Data

Our current analyses combine the 2015 - 2018 NHANES survey cycles to yield $n = 19225$ observations on 35 covariates. This sample size was comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and 2017/2018 survey cycles, respectively. Analyses excluded individuals with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055), and body mass index (BMI) or those with BMI > 25 kg/$m^2$ (overweight) (n = 5521). Individuals that reported being on anti-hypertensive medication were also excluded from further analyses (n = 2944).

### Outcome definition: 
The outcome is

The trained and certified examiners used the standardized protocols and calibrated equipment to get the blood
pressure readings. Three consecutive BP readings were acquired via ausculatory means. If a BP measurement
was not successfully completed, a fourth measurement was implemented. The average of all available measures
was used.

### Exposure:
Sleep duration on workdays was evaluated by the questionnaire with the following questions:
“Number of hours usually sleep on weekdays or workdays”. Sleep duration was divided into three groups, which
were < 6 h, 6–8 h, ≥ 8 h respectively, of which 6–8 h was used as the reference group.

### Covariates:

Table goes here for the covariate distributional questions

## Statistical modeling

Model 1
Model 3, 4, 5

## Methods 1

## Methods 2

## Methods 3

# Results

## Results 1

### Results 1a

# Discussion and conclusion

# References

<div id="refs"></div>

### 2023-12-03 data/script.R

@@ -1,0 +1,141 @@
+#############################
+## Sleep and Blood Pressure #
+## Amos Okutse              #
+## Monica Colon-Vargas      #
+#############################
+
+## load required libraries
+library(nhanesA)
+library(knitr)
+library(kableExtra)
+library(tidyverse)
+library(gtsummary)
+library(survey)
+
+## load the 2015 cycle and add an indicator of survey cycle
+load("../ProjectFinal/data/nhanes2015.RData")
+load("../ProjectFinal/data/nhanes2017.RData")
+nhanes2015$cycle <- 0
+nhanes2017$cycle <- 1
+
+## check what names do not match between the two surveys
+setdiff(names(nhanes2015), names(nhanes2017))
+
+## only column ALQ120Q is not the same and we can change that label in the 2015 cycle to match and merge
+nhanes2015 <- nhanes2015 %>% 
+  dplyr::rename(ALQ121 = ALQ120Q)
+
+## we can confirm that the names are now identical
+identical(names(nhanes2015), names(nhanes2017))
+
+## append the two data files together
+df <- rbind(nhanes2015, nhanes2017)
+
+## get the dimension of combined file
+dim(df)
+
+## rename variable names
+dt <- df %>% 
+  dplyr::rename("seq_no" = SEQN, 
+                "psu" = SDMVPSU, 
+                "strata" = SDMVSTRA, 
+                "gender" = RIAGENDR, 
+                "age_yr" = RIDAGEYR, 
+                "marital_status" = DMDMARTL, 
+                "weights" = WTINT2YR, 
+                "income_range" = INDFMIN2, 
+                "race" = RIDRETH3, 
+                "children <5" = DMDHHSZA, 
+                "educ_level" = DMDEDUC2, 
+                "citizenship_status" = DMDCITZN, 
+                "sleep" = SLD012, 
+                "bmi" = BMXBMI, 
+                "alcohol" = ALQ121, 
+                "hdl" = LBDHDDSI, 
+                "albumin" = LBDSALSI, 
+                "diabetes" = DIQ010, 
+                "hypertension" = BPQ020, 
+                "creatinine" = URXCRS, 
+                "snort" = SLQ040, 
+                "total_chol" = LBDTCSI,
+                "hemoglobin" = LBXHGB, 
+                "AST" = LBXSASSI, 
[…]

### 2023-12-03 data/sleep and HBP.Rmd
update

@@
 subtitle: "By Yingjie Su and published in Nature Scientific Reports"
+author: "Amos Okutse, Brown University"
 date: "  `r format(Sys.time(), '%d %B, %Y')` "

### 2023-12-05 data/script.R
update

@@ -4,6 +4,9 @@
 ## Monica Colon-Vargas      #
 #############################
 
+##Set Working Directory
+setwd("C:/Users/monic/OneDrive/Desktop/LinearModelsProject/")
+
 ## load required libraries
 library(nhanesA)
 library(knitr)
@@ -13,10 +16,14 @@ library(gtsummary)
 library(survey)
 
 ## load the 2015 cycle and add an indicator of survey cycle
-load("../ProjectFinal/data/nhanes2015.RData")
-load("../ProjectFinal/data/nhanes2017.RData")
+load("nhanes2015.RData")
+load("nhanes2017.RData")
 nhanes2015$cycle <- 0
 nhanes2017$cycle <- 1
+
+
+nhanes2015 %>% head()
+nhanes2017 %>% head()
 
 ## check what names do not match between the two surveys
 setdiff(names(nhanes2015), names(nhanes2017))
@@ -65,6 +72,21 @@ dt <- df %>%
 
 ## Table of questions used when extracting variables from the NHANES database
 
+## Check missing values per variable
+#Count missing data accros each variable
+missing_table <- dt %>% 
+  summarise(across(everything(), ~ sum(is.na(.x)))) %>%
+  t() %>% 
+  as.data.frame() %>%
+  mutate(missing = V1) %>%
+  select(missing) %>% mutate(percent = round(missing/dim(dt)[1],4)*100)
+
+missing_table %>%
+  filter(missing > 0) %>%
+  kable(caption = "Count of Missing Data across all variables",
+        booktabs=T,
+        col.names = c("Number of Missing","Percent")) %>% 
+  kable_styling(full_width=T, latex_options = c('HOLD_position'))
 
 ## drop missing sleep values
 dt <- dt %>% tidyr::drop_na(sleep)
@@ -77,10 +99,11 @@ dt$sbp <- rowMeans(subset(dt, select = c(BPXSY1, BPXSY2, BPXSY3, BPXSY4)), na.rm
 ## we can then delete the variables related to dbp and sbp from the data set and drop missing values for either dbp or sbp
 dt <- dt %>% 
[…]

### 2023-12-06 data/script.R
update

@@ -191,7 +191,7 @@ library(ggpubr)
 library(glmnet)
 
 vis_miss(dt)
-dt <- dt %>% select(-seq_no,-hyper_meds,-sleep,-race,-smoke,-snort,-alcohol)
+dt <- dt %>% select(-hyper_meds,-sleep,-race,-smoke,-snort,-alcohol)
 dt$cycle <- as.factor(dt$cycle)
 dt <- dt %>%
   rename(sleep = sleep2,
@@ -200,6 +200,7 @@ dt <- dt %>%
          snort = snort2,
          alcohol = alcohol2)
 
+
 mod1 <- lm(dbp ~. ,data = dt)
 summary(mod1)
 
@@ -207,6 +208,14 @@ summary(mod1)
 #After removing some vairables and filtering some NAs and Changin NA to not recorded
 dt <- na.omit(dt)
 
+dt
+
+
+## declare data as survey data
+design = svydesign(id = ~psu, weights = ~weights, strata = ~strata, 
+          nest = TRUE, survey.lonely.psu = "adjust", data = dt)
+mod <- svyglm()
+View(dt)
 
 ##EXPLORATORY DATA ANALYSIS
 ggcorrplot(cor(dt %>% select_if(is.numeric)))
@@ -297,6 +306,11 @@ dt <- dt %>%
     educ_level %in% c("Dont know", "Refused") ~ "Unknown/Refused")) %>%
   mutate(educ_level = as.factor(educ_level))
 
+## SAve
+#write.csv(dt, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data", row.names = F)
+write.csv(dt, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
+
+##########DBP
 
 
 
@@ -305,7 +319,7 @@ set.seed(123)
 folds <- sample(1:10,nrow(dt),replace = T)
 
 ## Create the matrix of the model
-X <- model.matrix(dbp~., data = dt %>% select(-sbp))[,-1]
+X <- model.matrix(dbp~., data = dt %>% select(-sbp,-weights,-strata))[,-1]
 y <- dt$dbp
 
 ## Do cross validation
@@ -319,7 +333,9 @@ plot(lasso.mod.cv) #maybe choose one larger lambda that not adds more mse?
 ## Select nonzero coefficients
 coefficients <- coef(lasso.mod)
 nonzero.coef <- coefficients[,1] != 0
-nonzero.vars <- rownames(coefficients)[nonzero.coef]
[…]

### 2023-12-08 data/script.R
update

@@ -308,7 +308,7 @@ dt <- dt %>%
 
 ## SAve
 #write.csv(dt, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data", row.names = F)
-write.csv(dt, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
+#write.csv(dt, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
 
 ##########DBP
 
@@ -319,8 +319,8 @@ set.seed(123)
 folds <- sample(1:10,nrow(dt),replace = T)
 
 ## Create the matrix of the model
-X <- model.matrix(dbp~., data = dt %>% select(-sbp,-weights,-strata))[,-1]
-y <- dt$dbp
+X <- model.matrix(sbp~., data = dt %>% select(-dbp,-weights,-strata))[,-1]
+y <- dt$sbp
 
 ## Do cross validation
 lasso.mod.cv <- cv.glmnet(X, y, alpha=1, nfolds = 10, foldid = folds)
@@ -333,9 +333,9 @@ plot(lasso.mod.cv) #maybe choose one larger lambda that not adds more mse?
 ## Select nonzero coefficients
 coefficients <- coef(lasso.mod)
 nonzero.coef <- coefficients[,1] != 0
-nonzero.vars.lasso <- rownames(coefficients)[nonzero.coef]
+nonzero.vars.lasso_sbp <- rownames(coefficients)[nonzero.coef]
 
-write.csv(nonzero.vars.lasso, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res.csv")
+write.csv(nonzero.vars.lasso_sbp, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_sbp.csv")
 
 ## Filter dataframe to include only non-zero variables
 modified_dt <- dt %>% 
@@ -355,7 +355,7 @@ library(leaps)
 
 k <- 10
 n <- nrow(dt)
-p <- ncol(dt %>% select(-sbp,-weights,-strata)) -1
+p <- ncol(dt %>% select(-dbp,-weights,-strata)) -1
 
 set.seed(1)
 #folds <- sample(1:k, n, replace=TRUE)
@@ -363,23 +363,25 @@ set.seed(1)
 #ISSUES  for stepwise selection 
 cv.errors <- matrix(NA,k,p,
[…]

### 2023-12-08 data/script.R
up

@@ -139,3 +139,4 @@ table1 <- dt2 %>% gtsummary::tbl_summary(by = gender,
   as_kable(format = "pipe", caption = "Baseline participant characteristics", digits = 2) #%>% 
 #kable_styling(latex_options = "scale_down")
 table1
+

### 2023-12-08 data/script.R
update

@@ -312,20 +312,21 @@ dt <- dt %>%
 
 ##########DBP
 
-
+dt <- dt %>%
+  mutate_if(is.character,as.factor)
 
 ## Set seed for folds
 set.seed(123)
 folds <- sample(1:10,nrow(dt),replace = T)
 
 ## Create the matrix of the model
-X <- model.matrix(sbp~., data = dt %>% select(-dbp,-weights,-strata))[,-1]
-y <- dt$sbp
+X <- model.matrix(dbp~., data = dt %>% select(-sbp,-weights,-strata,-seq_no))[,-1]
+y <- dt$dbp
 
 ## Do cross validation
 lasso.mod.cv <- cv.glmnet(X, y, alpha=1, nfolds = 10, foldid = folds)
 lasso.mod.cv$lambda.min
-lasso.mod <- glmnet(X,y,alpha = 1,lambda = lasso.mod.cv$lambda.min)
+lasso.mod <- glmnet(X,y,alpha = 1,lambda = lasso.mod.cv$lambda.1se)
 
 ## Plot cross validation results
 plot(lasso.mod.cv) #maybe choose one larger lambda that not adds more mse?
@@ -333,9 +334,9 @@ plot(lasso.mod.cv) #maybe choose one larger lambda that not adds more mse?
 ## Select nonzero coefficients
 coefficients <- coef(lasso.mod)
 nonzero.coef <- coefficients[,1] != 0
-nonzero.vars.lasso_sbp <- rownames(coefficients)[nonzero.coef]
+nonzero.vars.lasso_dbp_1se <- rownames(coefficients)[nonzero.coef]
 
-write.csv(nonzero.vars.lasso_sbp, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_sbp.csv")
+write.csv(nonzero.vars.lasso_dbp_1se, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_dbp_1se.csv")
 
 ## Filter dataframe to include only non-zero variables
 modified_dt <- dt %>% 
@@ -355,7 +356,7 @@ library(leaps)
 
 k <- 10
 n <- nrow(dt)
-p <- ncol(dt %>% select(-dbp,-weights,-strata)) -1
+p <- ncol(dt %>% select(-dbp,-weights,-strata,-seq_no)) -1
 
 set.seed(1)
 #folds <- sample(1:k, n, replace=TRUE)

### 2023-12-08 data/scrip2.R
update

@@ -17,7 +17,7 @@ dt <- dt %>%
 
 
 ## Get Summary Statistics
-dt %>%  dplyr::select(-seq_no) %>%
+dt %>%  dplyr::select(-seq_no,-strata,-psu,-weights) %>%
   tbl_summary(digits = list(everything() ~ c(2)),
               statistic = list(all_continuous() ~ "{mean} ({sd})"),
               by = gender) %>%
@@ -27,20 +27,14 @@ dt %>%  dplyr::select(-seq_no) %>%
 
 ## Systolic Blood Pressure
 ## Load variables selected by LASSO
-lasso.coef.min <- read.csv("../SleepBP/data/lasso_res_sbp.csv")
+lasso.coef.min <- read.csv("../SleepBP/data/lasso_res_sbp_min.csv")
 lasso.coef.min$x
 names.lasso.min <- c("bmi", "hdl", "total_chol", "hemoglobin", "albumin", "AST",
                      "hypertension", "creatinine", "diabetes", "citizenship_status",
                      "educ_level", "children..5", "gender", "age_yr", "marital_status",
-                     "cycle", "sleep", "race", "smoke", "snort", "alcohol", 
-                     "income_category")
+                     "sleep", "race", "smoke", "snort", "alcohol", "income_category")
 dt %>% select(all_of(names.lasso.min))
 
-lasso.coef.1se <- read.csv("../SleepBP/data/lasso_res_sbp_1se.csv")
-lasso.coef.1se$x
-names.lasso.1se <- c("total_chol", "hypertension", "educ_level", "gender",
-                     "age_yr", "marital_status", "race", "income_category")
-dt %>% select(all_of(names.lasso.1se),sbp)
 
 ## Fit lm
 ## With min
@@ -50,12 +44,6 @@ summary(modlasso.sbp.min)
 pred <- predict(modlasso.sbp.min, newdata = dt %>% select(all_of(names.lasso.min)))
 sum((pred-dt$sbp)^2)/nrow(dt)
 
-# With 1se
-modlasso.sbp.1se <- lm(sbp ~. ,data = dt %>% select(all_of(names.lasso.1se),sbp)
-)
-summary(modlasso.sbp.1se)
-pred <- predict(modlasso.sbp.1se, newdata = dt %>% select(all_of(names.lasso.1se)))
-sum((pred-dt$sbp)^2)/nrow(dt)
 
 ## Dyastolic Blood Pressure
[…]

### 2023-12-08 data/script.R

@@ -320,13 +320,13 @@ set.seed(123)
 folds <- sample(1:10,nrow(dt),replace = T)
 
 ## Create the matrix of the model
-X <- model.matrix(dbp~., data = dt %>% select(-sbp,-weights,-strata,-seq_no))[,-1]
-y <- dt$dbp
+X <- model.matrix(sbp~., data = dt %>% select(-dbp,-weights,-strata,-seq_no,-psu))[,-1]
+y <- dt$sbp
 
 ## Do cross validation
 lasso.mod.cv <- cv.glmnet(X, y, alpha=1, nfolds = 10, foldid = folds)
 lasso.mod.cv$lambda.min
-lasso.mod <- glmnet(X,y,alpha = 1,lambda = lasso.mod.cv$lambda.1se)
+lasso.mod <- glmnet(X,y,alpha = 1,lambda = lasso.mod.cv$lambda.min)
 
 ## Plot cross validation results
 plot(lasso.mod.cv) #maybe choose one larger lambda that not adds more mse?
@@ -334,9 +334,9 @@ plot(lasso.mod.cv) #maybe choose one larger lambda that not adds more mse?
 ## Select nonzero coefficients
 coefficients <- coef(lasso.mod)
 nonzero.coef <- coefficients[,1] != 0
-nonzero.vars.lasso_dbp_1se <- rownames(coefficients)[nonzero.coef]
+nonzero.vars.lasso_sbp_min <- rownames(coefficients)[nonzero.coef]
 
-write.csv(nonzero.vars.lasso_dbp_1se, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_dbp_1se.csv")
+write.csv(nonzero.vars.lasso_sbp_min, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_sbp_min.csv")
 
 ## Filter dataframe to include only non-zero variables
 modified_dt <- dt %>% 
@@ -356,7 +356,7 @@ library(leaps)
 
 k <- 10
 n <- nrow(dt)
-p <- ncol(dt %>% select(-dbp,-weights,-strata,-seq_no)) -1
+p <- ncol(dt %>% select(-dbp,-weights,-strata,-seq_no, -psu)) -1
 
 set.seed(1)
 #folds <- sample(1:k, n, replace=TRUE)

### 2023-12-08 manuscript/paper.Rmd
up

@@
 - Graduate Student, Graduate School
-abstract: "Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Donec sit amet libero justo.
-Pellentesque eget nibh ex.
-Aliquam tincidunt egestas lectus id ullamcorper.
-Proin tellus orci, posuere sed cursus at, bibendum ac odio.
+abstract: "**Introduction:** Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec sit amet libero justo. Pellentesque eget nibh ex. Aliquam tincidunt egestas lectus id ullamcorper.
+
+**Methods:** Proin tellus orci, posuere sed cursus at, bibendum ac odio.
 Nam consequat non ante eget aliquam.
@@
 -   As a case on point ...?
-- Underscores the importance of reproducibility within the frame of statistical analysis. We use the study by to ...
-The current study builds on the paper by ..
-and starts in the manner of a replication study and extends the analyses therein.
-We focus on statistical analyses of these data and compare models based on varied specification of functional forms of the covariates.
+-   Underscores the importance of reproducibility within the frame of
+    statistical analysis. We use the study by to ... The current study
+    builds on the paper by .. and starts in the manner of a replication
+    study and extends the analyses therein. We focus on statistical
+    analyses of these data and compare models based on varied
+    specification of functional forms of the covariates.
 
@@
 
-Data utilized in this paper is sourced from the National Health and Nutrition Examination Survey (NHANES), a comprehensive nationwide survey administered by the National Center for Health Statistics (NCHS) via the Centers for Disease Control and Prevention (CDC).
-The survey assesses the health and nutrition of the entire non-institutionalized US population, spanning all ages and residing in all 50 states as well as Washington D.C.
-As such, the survey provides a cross-sectional view of a representative sample of the US population.
-Further information about NHANES can be found at www.cdc.gov/nchs/nhanes.
+Data utilized in this paper is sourced from the National Health and
+Nutrition Examination Survey (NHANES), a comprehensive nationwide survey
+administered by the National Center for Health Statistics (NCHS) via the
+Centers for Disease Control and Prevention (CDC). The survey assesses
+the health and nutrition of the entire non-institutionalized US
+population, spanning all ages and residing in all 50 states as well as
+Washington D.C. As such, the survey provides a cross-sectional view of a
+representative sample of the US population. Further information about
+NHANES can be found at www.cdc.gov/nchs/nhanes.
 
@@
 
-Our current analyses combine the 2015 - 2018 NHANES survey cycles to yield $n = 19225$ observations on 35 covariates.
-This sample size was comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and 2017/2018 survey cycles, respectively.
-Analyses excluded individuals with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055), and body mass index (BMI) or those with BMI > 25 kg/$m^2$ (overweight) (n = 5521).
-Individuals that reported being on anti-hypertensive medication were also excluded from further analyses (n = 2944).
+Our current analyses combine the 2015 - 2018 NHANES survey cycles to
+yield $n = 19225$ observations on 35 covariates. This sample size was
+comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and
+2017/2018 survey cycles, respectively. Analyses excluded individuals
+with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055),
+and body mass index (BMI) or those with BMI \> 25 kg/$m^2$ (overweight)
+(n = 5521). Individuals that reported being on anti-hypertensive
+medication were also excluded from further analyses (n = 2944).
 
 ### Outcome definition:
+
 The outcome is
 
-The trained and certified examiners used the standardized protocols and calibrated equipment to get the blood
-pressure readings. Three consecutive BP readings were acquired via ausculatory means. If a BP measurement
-was not successfully completed, a fourth measurement was implemented. The average of all available measures
-was used.
+The trained and certified examiners used the standardized protocols and
+calibrated equipment to get the blood pressure readings. Three
+consecutive BP readings were acquired via ausculatory means. If a BP
+measurement was not successfully completed, a fourth measurement was
+implemented. The average of all available measures was used.
 
 ### Exposure:
-Sleep duration on workdays was evaluated by the questionnaire with the following questions:
-“Number of hours usually sleep on weekdays or workdays”. Sleep duration was divided into three groups, which
-were < 6 h, 6–8 h, ≥ 8 h respectively, of which 6–8 h was used as the reference group.
+
+Sleep duration on workdays was evaluated by the questionnaire with the
+following questions: "Number of hours usually sleep on weekdays or
+workdays". Sleep duration was divided into three groups, which were \< 6
+h, 6--8 h, ≥ 8 h respectively, of which 6--8 h was used as the reference
+group.
 
@@
 
-Model 1
-Model 3, 4, 5
+Model 1 Model 3, 4, 5
 
@@
 
-<div id="refs"></div>
+::: {#refs}
+:::

### 2023-12-08 data/scrip2.R

@@ -7,6 +7,7 @@
 library(dplyr)
 library(gtsummary)
 library(survey)
+library(ggcorrplot)
 
 ## Load Data
 dt <- read.csv("../SleepBP/data/dt.csv")
@@ -43,10 +44,10 @@ modlasso.sbp.min <- lm(sbp ~. ,data = dt %>% select(all_of(names.lasso.min),sbp)
 summary(modlasso.sbp.min)
 pred <- predict(modlasso.sbp.min, newdata = dt %>% select(all_of(names.lasso.min)))
 sum((pred-dt$sbp)^2)/nrow(dt)
-
+tbl_regression(modlasso.sbp.min, include = everything()) %>%
+  modify_column_unhide(column = std.error)
 
 ## Dyastolic Blood Pressure
-## Load variables selected by LASSO
 ## Load variables selected by LASSO
 lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_dbp_min.csv")
 lasso.coef.min$x
@@ -58,22 +59,20 @@ names.lasso.min <- c("bmi","total_chol", "hemoglobin", "albumin", "AST",
 dt %>% select(all_of(names.lasso.min))
 
 ## Fit lm
-## With min
-modlasso.dbp.min <- lm(dbp ~. ,data = dt %>% select(all_of(names.lasso.min),dbp)
-)
+## With Min
+modlasso.dbp.min <- lm(dbp ~. ,data = dt %>% select(all_of(names.lasso.min),dbp))
 summary(modlasso.dbp.min)
 pred <- predict(modlasso.dbp.min, newdata = dt %>% select(all_of(names.lasso.min)))
 sum((pred-dt$dbp)^2)/nrow(dt)
 
 
+tab1 <- tbl_regression(modlasso.sbp.min, include = everything()) %>%
+  modify_column_unhide(column = std.error)
 
-## With survery 
-## For sbp
-svydesign(id = ~psu, weights = ~weights, strata = ~strata, 
-          nest = TRUE, survey.lonely.psu = "adjust", 
-          data = dt %>% select(all_of(names.lasso.min),sbp, weights, strata, psu))
+tab2 <- tbl_regression(modlasso.dbp.min, include = everything()) %>%
+  modify_column_unhide(column = std.error)
 
[…]

### 2023-12-08 data/script.R

@@ -462,3 +462,106 @@ ggplot(coef_values, aes(x = variables, y = V1)) +
   labs(x = "Variables", y = "Variable Importance") +
   theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
+
+
+
+
+
+
+
+
+
+
+
+
+##############
+#SYSTOLIC
+library(leaps)
+# Perform forward subset selection
+# Cross-Validation
+
+k <- 10
+n <- nrow(dt)
+p <- ncol(dt %>% select(-dbp,-weights,-strata,-seq_no, -psu,-hypertension)) -1
+set.seed(123)
+folds <- sample(1:k, n, replace=TRUE)
+
+#ISSUES  for stepwise selection 
+cv.errors.sbp <- matrix(NA,k,p,
+                    dimnames = list(NULL, paste(1:p))) #This takes too long, maybe use only variables of lasso??
+modified_dt <- dt %>% dplyr::select(-dbp,-weights,-strata,-seq_no, -psu, - hypertension)
+for (j in 1:k) {
+  best.fit <- regsubsets(sbp ~., 
+                         data = modified_dt[folds != j, ],
+                         nvmax = p, 
+                         really.big = T)
+  print(j) 
+  for (i in 1:p) {
+    coefi <- coef(best.fit, id = i)
+    test.mat <- model.matrix(sbp ~., data = modified_dt[folds == j,])
+    pred <- test.mat[,names(coefi)] %*% coefi
+    cv.errors.sbp[j,i]  <- mean((modified_dt[folds == j,]$sbp - pred)^2)
+    print(i)
+  }
+}
+errors.sbp <- data.frame(subset_size = 1:p, error = colMeans(cv.errors.sbp))
+best_k.sbp <- which.min(errors.sbp$error)
+ggplot(errors.sbp, aes(x = subset_size, y = error, color = subset_size == best_k.sbp)) +geom_point()+theme_minimal()+
+  geom_line()+
+  scale_color_manual(values = c("TRUE" = "red", "FALSE" = "black"))+
+  scale_x_continuous(breaks = 1:p)+
+  guides(color = "none")+ggtitle("Cross Valiation for Foward Setsize")
+
+
+## Get coefficients with best_k
+reg.best.sbp <- regsubsets(sbp~., data = modified_dt, nvmax = p)
+
+
+
+##########
+##DYstolic
+k <- 10
+n <- nrow(dt)
[…]

### 2023-12-09 data/scrip2.R
update

@@ -9,11 +9,10 @@ library(gtsummary)
 library(survey)
 
 ## Load Data
-dt <- read.csv("../SleepBP/data/dt.csv")
+dt <- read.csv("../ProjectFinal/data/dt.csv")
 dt <- dt %>% select(-X)
 dt <- dt %>%
   mutate_if(is.character,as.factor)
-#View(dt)
 
 
 ## Get Summary Statistics

### 2023-12-09 data/script.R

@@ -4,9 +4,6 @@
 ## Monica Colon-Vargas      #
 #############################
 
-##Set Working Directory
-setwd("C:/Users/monic/OneDrive/Desktop/LinearModelsProject/")
-
 ## load required libraries
 library(nhanesA)
 library(knitr)
@@ -14,6 +11,7 @@ library(kableExtra)
 library(tidyverse)
 library(gtsummary)
 library(survey)
+library(leaps)
 
 ## load the 2015 cycle and add an indicator of survey cycle
 load("nhanes2015.RData")
@@ -72,8 +70,17 @@ dt <- df %>%
 
 ## Table of questions used when extracting variables from the NHANES database
 
+tbl_one <- data.frame(Variable = as.character(), Name = as.character(), Description = as.character()) %>% 
+  add_row(Variable = "SEQN", Name = "sequence number", Description = "Respondent sequence number") %>% 
+  add_row(Variable = "SDMVPSU", Name = "psu", Description = "Masked variance unit pseudo-PSU variable for variance estimation") %>% 
+  add_row(Variable = "SDMVSTRA", Name = "strata", Description = "Masked variance unit pseudo-stratum variable for variance estimation") %>% 
+  add_row(Variable = "RIAGENDR", Name = "gender", Description = "Gender of the participant") %>% 
+  add_row(Variable = "", )
+  
+
+
 ## Check missing values per variable
-#Count missing data accros each variable
+#Count missing data across each variable
 missing_table <- dt %>% 
   summarise(across(everything(), ~ sum(is.na(.x)))) %>%
   t() %>% 
@@ -176,13 +183,6 @@ table1 <- dt2 %>% gtsummary::tbl_summary(by = gender,
 #kable_styling(latex_options = "scale_down")
 table1
 
-<<<<<<< HEAD
-=======
-
-
-
-
-
 #age alcohol sleep and smoking status
 
 library(tidyverse)
@@ -283,9 +283,7 @@ ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,p13,p14,ncol = 3,nrow = 5,legen
 
 
 dt$income_range %>% levels()
-# Recategorizing income ranges
-
-
+# Re-categorizing income ranges
 dt$income_category <- case_when(
   dt$income_range %in% c("$ 0 to $ 4,999", "$ 5,000 to $ 9,999") ~ "Low income",
[…]

### 2023-12-09 manuscript/paper.Rmd
update

@@
 title: "Preparing the NHANES Data for the Replication of: Association between sleep duration on workdays and blood pressure in non‑overweight/obese population in NHANES: a public database research"
-thanks: "Replication files are available on the author's Github account (http://github.com/svmiller/svm-r-markdown-templates).
+thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP).
 **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
@@
 
-Our current analyses combine the 2015 - 2018 NHANES survey cycles to
-yield $n = 19225$ observations on 35 covariates. This sample size was
-comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and
-2017/2018 survey cycles, respectively. Analyses excluded individuals
-with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055),
-and body mass index (BMI) or those with BMI \> 25 kg/$m^2$ (overweight)
-(n = 5521). Individuals that reported being on anti-hypertensive
-medication were also excluded from further analyses (n = 2944).
+Our current analyses combine the 2015 - 2018 NHANES survey cycles to yield $n = 19225$ observations on 35 covariates.
+This sample size was comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and 2017/2018 survey cycles, respectively.
+Analyses excluded individuals with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055), and body mass index (BMI) or those with BMI > 25 kg/$m^2$ (overweight) (n = 5521).
+Individuals that reported being on anti-hypertensive medication were also excluded from further analyses (n = 2944).
 
@@
 following questions: "Number of hours usually sleep on weekdays or
-workdays". Sleep duration was divided into three groups, which were \< 6
+workdays". Sleep duration was divided into three groups, which were < 6
 h, 6--8 h, ≥ 8 h respectively, of which 6--8 h was used as the reference
@@
 # References
-
-::: {#refs}
-:::

### 2023-12-09 manuscript/paper.Rmd
update

@@
 title: "Preparing the NHANES Data for the Replication of: Association between sleep duration on workdays and blood pressure in non‑overweight/obese population in NHANES: a public database research"
-thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP).
-**Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
+thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
 subtitle: "Short title goes here!"
@@
 
+```{r source-file, echo=FALSE, include=FALSE}
+# source file 
+source(file = "/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/ProjectFinal/data/scrip2.R")
+```
 # Introduction
@@
 
-The outcome is
-
-The trained and certified examiners used the standardized protocols and
-calibrated equipment to get the blood pressure readings. Three
-consecutive BP readings were acquired via ausculatory means. If a BP
-measurement was not successfully completed, a fourth measurement was
-implemented. The average of all available measures was used.
+We defined our outcome as systolic and diastolic blood pressure.
+These variables are measured by trained examiners using standardized procedures.
+Given that systolic and diastolic blood pressure measurements are taken at least four times on an individual, our definition of these outcome is based on an average of the first three blood pressure measurements.
 
@@
 
-Table goes here for the covariate distributional questions
+In addition to the exposure and outcome variables, our analyses included the following as covariates: Race divided into four groups as Mexican American, white, black and other race.
+Alcohol consumption was grouped into drinking, no drinking, not recorded.
+Smoking status as smoking, not smoking, not recorded.
+Diabetes was defined as yes, no, borderline, or not recorded.
+Hypertension was defined as yes, no, or not
+recorded.
+Snorting was defined as yes, no, and not recorded.
+US citizenship status was defined as citizen by birth or naturalization, don't know, not a citizen, or refused to answer.
+Education level was grouped into four categories including graduate studies, high school, less than grade 12 or some college.
+Additional covariates included marital status, gender, age, albumin, creatinine,hemoglobin, total cholesterol (TC), aspertate aminotransferase (AST), high-density lipoprotein (HDL), and body mass index (BMI).
+Details about these variables can be found at https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx.
+Age, albumin, creatinine, hemoglobin, TC, AST, HDL, and BMI were analyzed as continuous variables whereas gender, alcohol consumption, diabetes, smoking,
+race, hypertension, and snorting were analyzed as categorical variables. Table \@ref{tab:tabone} highlights the variable names and descriptions as utilized in this study.
+
+```{r tabone, echo=FALSE}
+tab.one <- kable(vars, caption = "Variable names and descriptions", 
+                 format = "latex") %>% 
+  kable_styling(latex_options = c("HOLD_position")) %>% 
+  row_spec(c(0), bold = TRUE) %>% 
+  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>% 
+  column_spec(3, width = "8cm")
+tab.one
+```
 
@@
 
-Model 1 Model 3, 4, 5
-
-## Methods 1
-
-## Methods 2
-
-## Methods 3
-
 # Results
 
-## Results 1
+Table \@ref(tab:tabtwo) summarizes participant characteristics in the 2015 - 2018 NHANES survey cycles stratified by gender.
+Among the participants, the proportion of males and females were 44.81% (n = 1378) and 55.19% (n = 1509),
+respectively. With ethnicity, the proportion of Mexican American, White and Black were 6.27%, 64.62% and
+9.94%, respectively. Overall, the mean (SD) values for age, albumin, creatinine, SBP, DBP, hemoglobin, TC, AST,
+HDL, BMI were 38.54 (17.72) years, 43.48 (3.58) g/L, 73.21 (19.73) umol/L, 115.33 (14.56) mmHg, 68.47 (10.08)
+mmHg, 14.07(1.44)g/dL, 4.67 (1.01) mmol/L, 23.18 (13.01) IU/L, 1.59 (0.43) mmol/L, and 21.97(2.06) kg/m2,
+respectively. Among the participants, 66.01% were consumed alcohol drinkers, 2.61% were diabetes, 3.16% were
+hypertension, 6.72% have snort or stop breathing. 17.99% were smokers. Sleep duration was divided into three
+groups, which were < 6 h, 6–8 h, ≥ 8 h, each with a proportion of 6.26%, 40.53%, 53.21%, respectively.
 
-### Results 1a
+Male and female participants differed significantly on characteristics including
+
+The univariate
+analysis of potential confounding factors is shown in Table 2.
+
+```{r tabtwo, echo=FALSE}
+tabsum %>% 
+  as_kable_extra(format = "latex", caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", digits = 2, longtable = T) %>% 
+  kable_styling(latex_options = c("HOLD_position"))
+```
 

### 2023-12-09 data/scrip2.R

@@ -8,70 +8,104 @@ library(dplyr)
 library(gtsummary)
 library(survey)
 library(ggcorrplot)
+library(kableExtra)
 
 ## Load Data
-dt <- read.csv("../ProjectFinal/data/dt.csv")
+dt <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/ProjectFinal/data/dt.csv")
+
+## load the variable name file
+vars <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/ProjectFinal/data/vars.csv")
+
+## some pre-processing
 dt <- dt %>% select(-X)
 dt <- dt %>%
-  mutate_if(is.character,as.factor)
+  mutate_if(is.character, as.factor)
 
+## edit the citizenship levels: Citizen, not citizen/other
+## Marital status: single/never married, married or living with partner, divorced or separated, widowed 
 
-## Get Summary Statistics
-dt %>%  dplyr::select(-seq_no,-strata,-psu,-weights) %>%
+## summarize the data by gender
+tabsum <- dt %>%  
+  dplyr::select(-seq_no,-strata,-psu,-weights) %>%
   tbl_summary(digits = list(everything() ~ c(2)),
               statistic = list(all_continuous() ~ "{mean} ({sd})"),
-              by = gender) %>%
+              by = gender,
+              label = list(bmi ~ "BMI",
+                           hdl ~ "HDL",
+                           total_chol ~ "TC",
+                           hemoglobin ~ "Hemoglobin",
+                           albumin ~ "Albumin",
+                           creatinine ~ "Creatinine",
+                           hypertension ~ "Hypertension",
+                           diabetes ~ "Diabetes",
+                           citizenship_status ~ "Citizenship",
+                           educ_level ~ "Education",
+                           children..5 ~ "Children > 5 yrs",
+                           age_yr ~ "Age (yrs)",
+                           marital_status ~ "Marital status",
+                           cycle ~ "Survey cycle",
+                           dbp ~ "DBP",
+                           sbp ~ "SBP",
+                           sleep ~ "Sleep",
+                           race ~ "Race",
[…]

### 2023-12-09 data/script.R
update

@@ -320,23 +320,31 @@ set.seed(123)
 folds <- sample(1:10,nrow(dt),replace = T)
 
 ## Create the matrix of the model
-X <- model.matrix(sbp~., data = dt %>% select(-dbp,-weights,-strata,-seq_no,-psu))[,-1]
-y <- dt$sbp
+X <- model.matrix(dbp~., 
+                  data = dt %>% select(-sbp,-weights,-strata,-seq_no,-psu,-hypertension))[,-1]
+y <- dt$dbp
 
 ## Do cross validation
 lasso.mod.cv <- cv.glmnet(X, y, alpha=1, nfolds = 10, foldid = folds)
 lasso.mod.cv$lambda.min
 lasso.mod <- glmnet(X,y,alpha = 1,lambda = lasso.mod.cv$lambda.min)
 
+(191.9156-190.1472)/190.1472*100 #29 
+lambda.log = -1.305027684
+lambda = exp(lambda.log)
+## Check for best lambda
+cbind(log(lasso.mod.cv$lambda),lasso.mod.cv$cvup, lasso.mod.cv$nzero)
+ 
+
 ## Plot cross validation results
 plot(lasso.mod.cv) #maybe choose one larger lambda that not adds more mse?
-
+title("Diastolic Blood Pressure", line = 2.5)
 ## Select nonzero coefficients
 coefficients <- coef(lasso.mod)
 nonzero.coef <- coefficients[,1] != 0
-nonzero.vars.lasso_sbp_min <- rownames(coefficients)[nonzero.coef]
+nonzero.vars.lasso_dbp_min <- rownames(coefficients)[nonzero.coef]
 
-write.csv(nonzero.vars.lasso_sbp_min, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_sbp_min.csv")
+write.csv(nonzero.vars.lasso_dbp_min, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_dbp_min.csv")
 
 ## Filter dataframe to include only non-zero variables
 modified_dt <- dt %>% 
@@ -523,7 +531,7 @@ reg.best.sbp <- regsubsets(sbp~., data = modified_dt, nvmax = p)
 k <- 10
 n <- nrow(dt)
 p <- ncol(dt %>% select(-sbp,-weights,-strata,-seq_no, -psu,-hypertension)) -1
-set.seed(123)
[…]

### 2023-12-09 data/scrip2.R
update

@@ -11,10 +11,10 @@ library(ggcorrplot)
 library(kableExtra)
 
 ## Load Data
-dt <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/ProjectFinal/data/dt.csv")
+dt <- read.csv("data/dt.csv")
 
 ## load the variable name file
-vars <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/ProjectFinal/data/vars.csv")
+vars <- read.csv("data/vars.csv")
 
 ## some pre-processing
 dt <- dt %>% select(-X)
@@ -59,53 +59,105 @@ tabsum <- dt %>%
   add_p()
 
 
-# ## Systolic Blood Pressure
-# ## Load variables selected by LASSO
-# lasso.coef.min <- read.csv("../ProjectFinal/SleepBP/data/lasso_res_sbp_min.csv")
-# lasso.coef.min$x
-# names.lasso.min <- c("bmi", "hdl", "total_chol", "hemoglobin", "albumin", "AST",
-#                      "hypertension", "creatinine", "diabetes", "citizenship_status",
-#                      "educ_level", "children..5", "gender", "age_yr", "marital_status",
-#                      "sleep", "race", "smoke", "snort", "alcohol", "income_category")
-# dt %>% select(all_of(names.lasso.min))
-# 
-# 
-# ## Fit lm
-# ## With min
-# modlasso.sbp.min <- lm(sbp ~. ,data = dt %>% select(all_of(names.lasso.min),sbp)
-# )
-# summary(modlasso.sbp.min)
-# pred <- predict(modlasso.sbp.min, newdata = dt %>% select(all_of(names.lasso.min)))
-# sum((pred-dt$sbp)^2)/nrow(dt)
-# tbl_regression(modlasso.sbp.min, include = everything()) %>%
-#   modify_column_unhide(column = std.error)
-# 
-# ## Dyastolic Blood Pressure
-# ## Load variables selected by LASSO
-# lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_dbp_min.csv")
-# lasso.coef.min$x
-# names.lasso.min <- c("bmi","total_chol", "hemoglobin", "albumin", "AST",
-#                      "hypertension", "diabetes", "citizenship_status",
[…]

### 2023-12-09 data/script.R

@@ -564,16 +564,33 @@ ggplot(errors.dbp, aes(x = subset_size, y = error, color = subset_size == best_k
 ## Get coefficients with best_k
 reg.best.dbp <- regsubsets(dbp~., data = modified_dt, nvmax = p)
 
+
+
+
+best_k.dbp
+best_k.sbp
 reg.best.dbp <- as.data.frame(coef(reg.best.dbp,id=best_k.dbp))
 reg.best.sbp <- as.data.frame(coef(reg.best.sbp,id=best_k.sbp))
+
+reg.best.dbp$K = best_k.dbp
+reg.best.sbp$K <- best_k.sbp
 
 write.csv(reg.best.dbp, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/reg.dbp.csv")
 write.csv(reg.best.sbp, file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/reg.sbp.csv")
 
 
 
+summary(reg.best.dbp)
 
 
 
+## Get coefficients with best_k
+reg.best <- regsubsets(sbp~., data = modified_dt, nvmax = p)
+
+reg.best.dbp
+summary(reg.best.dbp)
+reg.best.dbp$call
 
 
+coef_reg_bestk_sbp <- coef(reg.best, best_k) #levels of variables but not whole variables
+ coef(reg.best.dbp, id=13) 

### 2023-12-09 manuscript/paper.Rmd
upd

@@
 # source file 
-source(file = "/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/ProjectFinal/data/scrip2.R")
+source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
 ```
+
 # Introduction
@@
 
-Our current analyses combine the 2015 - 2018 NHANES survey cycles to yield $n = 19225$ observations on 35 covariates.
-This sample size was comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and 2017/2018 survey cycles, respectively.
-Analyses excluded individuals with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055), and body mass index (BMI) or those with BMI > 25 kg/$m^2$ (overweight) (n = 5521).
-Individuals that reported being on anti-hypertensive medication were also excluded from further analyses (n = 2944).
+Our current analyses combine the 2015 - 2018 NHANES survey cycles to
+yield $n = 19225$ observations on 35 covariates. This sample size was
+comprised $n = 9971$ and $n = 9254$ observations from the 2015/2016 and
+2017/2018 survey cycles, respectively. Analyses excluded individuals
+with missing data on sleep ($n = 6818$), blood pressure (BP) (n = 1055),
+and body mass index (BMI) or those with BMI \> 25 kg/$m^2$ (overweight)
+(n = 5521). Individuals that reported being on anti-hypertensive
+medication were also excluded from further analyses (n = 2944).
 
@@
 
-We defined our outcome as systolic and diastolic blood pressure.
-These variables are measured by trained examiners using standardized procedures.
-Given that systolic and diastolic blood pressure measurements are taken at least four times on an individual, our definition of these outcome is based on an average of the first three blood pressure measurements.
+We defined our outcome as systolic and diastolic blood pressure. These
+variables are measured by trained examiners using standardized
+procedures. Given that systolic and diastolic blood pressure
+measurements are taken at least four times on an individual, our
+definition of these outcome is based on an average of the first three
+blood pressure measurements.
 
@@
 following questions: "Number of hours usually sleep on weekdays or
-workdays". Sleep duration was divided into three groups, which were < 6
+workdays". Sleep duration was divided into three groups, which were \< 6
 h, 6--8 h, ≥ 8 h respectively, of which 6--8 h was used as the reference
@@
 
-In addition to the exposure and outcome variables, our analyses included the following as covariates: Race divided into four groups as Mexican American, white, black and other race.
-Alcohol consumption was grouped into drinking, no drinking, not recorded.
-Smoking status as smoking, not smoking, not recorded.
-Diabetes was defined as yes, no, borderline, or not recorded.
-Hypertension was defined as yes, no, or not
-recorded.
-Snorting was defined as yes, no, and not recorded.
-US citizenship status was defined as citizen by birth or naturalization, don't know, not a citizen, or refused to answer.
-Education level was grouped into four categories including graduate studies, high school, less than grade 12 or some college.
-Additional covariates included marital status, gender, age, albumin, creatinine,hemoglobin, total cholesterol (TC), aspertate aminotransferase (AST), high-density lipoprotein (HDL), and body mass index (BMI).
-Details about these variables can be found at https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx.
-Age, albumin, creatinine, hemoglobin, TC, AST, HDL, and BMI were analyzed as continuous variables whereas gender, alcohol consumption, diabetes, smoking,
-race, hypertension, and snorting were analyzed as categorical variables. Table \@ref{tab:tabone} highlights the variable names and descriptions as utilized in this study.
+In addition to the exposure and outcome variables, our analyses included
+the following as covariates: Race divided into four groups as Mexican
+American, white, black and other race. Alcohol consumption was grouped
+into drinking, no drinking, not recorded. Smoking status as smoking, not
+smoking, not recorded. Diabetes was defined as yes, no, borderline, or
+not recorded. Hypertension was defined as yes, no, or not recorded.
+Snorting was defined as yes, no, and not recorded. US citizenship status
+was defined as citizen by birth or naturalization, don't know, not a
+citizen, or refused to answer. Education level was grouped into four
+categories including graduate studies, high school, less than grade 12
+or some college. Additional covariates included marital status, gender,
+age, albumin, creatinine,hemoglobin, total cholesterol (TC), aspertate
+aminotransferase (AST), high-density lipoprotein (HDL), and body mass
+index (BMI). Details about these variables can be found at
+<https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx>. Age, albumin,
+creatinine, hemoglobin, TC, AST, HDL, and BMI were analyzed as
+continuous variables whereas gender, alcohol consumption, diabetes,
+smoking, race, hypertension, and snorting were analyzed as categorical
+variables. Table \@ref{tab:tabone} highlights the variable names and
+descriptions as utilized in this study.
 
 ```{r tabone, echo=FALSE}
-tab.one <- kable(vars, caption = "Variable names and descriptions", 
-                 format = "latex") %>% 
-  kable_styling(latex_options = c("HOLD_position")) %>% 
-  row_spec(c(0), bold = TRUE) %>% 
-  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>% 
-  column_spec(3, width = "8cm")
-tab.one
+#tab.one <- as_kable_extra(vars, caption = "Variable names and descriptions", 
+#                 format = "latex", booktabs = TRUE) %>% 
+#  kable_styling(latex_options = c("HOLD_position") %>% 
+#  row_spec(c(0), bold = TRUE) %>% 
+#  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>% 
+#  column_spec(3, width = "8cm")
+#tab.one
+
+vars %>% kable(booktabs = T) %>% kable_styling(latex_options = "scale_down")
+
 ```
@@
 
+In order to select variables impacting blood pressure, we proceed by
+doing some variable selection with regularization techniques. We first
+employ best subset selection by using 10-fold cross validation to select
+the best number of subsets. For syastolic blood pressure, forward best
+subset selection choose 13 variables and for diastolic blood pressure 19
+variables. These variables are showed, along with its significance in
+Table \@ref(tab:tabreg)
+
+```{r tabref, echo=FALSE}
+tab.ref %>% as_kable_extra(booktabs = T, caption ="Coefficients for Foward Best Subset Selection") %>% kable_styling(latex_options = "scale_down")
+
+```
+
 # Results
 
-Table \@ref(tab:tabtwo) summarizes participant characteristics in the 2015 - 2018 NHANES survey cycles stratified by gender.
-Among the participants, the proportion of males and females were 44.81% (n = 1378) and 55.19% (n = 1509),
-respectively. With ethnicity, the proportion of Mexican American, White and Black were 6.27%, 64.62% and
-9.94%, respectively. Overall, the mean (SD) values for age, albumin, creatinine, SBP, DBP, hemoglobin, TC, AST,
-HDL, BMI were 38.54 (17.72) years, 43.48 (3.58) g/L, 73.21 (19.73) umol/L, 115.33 (14.56) mmHg, 68.47 (10.08)
-mmHg, 14.07(1.44)g/dL, 4.67 (1.01) mmol/L, 23.18 (13.01) IU/L, 1.59 (0.43) mmol/L, and 21.97(2.06) kg/m2,
-respectively. Among the participants, 66.01% were consumed alcohol drinkers, 2.61% were diabetes, 3.16% were
-hypertension, 6.72% have snort or stop breathing. 17.99% were smokers. Sleep duration was divided into three
-groups, which were < 6 h, 6–8 h, ≥ 8 h, each with a proportion of 6.26%, 40.53%, 53.21%, respectively.
+Table \@ref(tab:tabtwo) summarizes participant characteristics in the
+2015 - 2018 NHANES survey cycles stratified by gender. Among the
+participants, the proportion of males and females were 44.81% (n = 1378)
+and 55.19% (n = 1509), respectively. With ethnicity, the proportion of
+Mexican American, White and Black were 6.27%, 64.62% and 9.94%,
+respectively. Overall, the mean (SD) values for age, albumin,
+creatinine, SBP, DBP, hemoglobin, TC, AST, HDL, BMI were 38.54 (17.72)
+years, 43.48 (3.58) g/L, 73.21 (19.73) umol/L, 115.33 (14.56) mmHg,
+68.47 (10.08) mmHg, 14.07(1.44)g/dL, 4.67 (1.01) mmol/L, 23.18 (13.01)
+IU/L, 1.59 (0.43) mmol/L, and 21.97(2.06) kg/m2, respectively. Among the
+participants, 66.01% were consumed alcohol drinkers, 2.61% were
+diabetes, 3.16% were hypertension, 6.72% have snort or stop breathing.
+17.99% were smokers. Sleep duration was divided into three groups, which
+were \< 6 h, 6--8 h, ≥ 8 h, each with a proportion of 6.26%, 40.53%,
+53.21%, respectively.
 
-Male and female participants differed significantly on characteristics including
+Male and female participants differed significantly on characteristics
+including
 
-The univariate
-analysis of potential confounding factors is shown in Table 2.
+The univariate analysis of potential confounding factors is shown in
+Table 2.
 
 ```{r tabtwo, echo=FALSE}
+#tabsum %>% 
+#  as_kable_extra(format = "latex", caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", digits = 2, longtable = T) %>% 
+#  kable_styling(latex_options = c("HOLD_position"))
+
 tabsum %>%
-  as_kable_extra(format = "latex", caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", digits = 2, longtable = T) %>% 
-  kable_styling(latex_options = c("HOLD_position"))
+  as_kable_extra(booktabs = TRUE, 
+                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", 
+                 longtable = TRUE) %>%
+  kable_styling(latex_options = "scale_down")
 ```

### 2023-12-09 data/scrip2.R

@@ -11,10 +11,10 @@ library(ggcorrplot)
 library(kableExtra)
 
 ## Load Data
-dt <- read.csv("data/dt.csv")
+dt <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
 
 ## load the variable name file
-vars <- read.csv("data/vars.csv")
+vars <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/vars.csv")
 
 ## some pre-processing
 dt <- dt %>% select(-X)
@@ -61,7 +61,7 @@ tabsum <- dt %>%
 
  ## Systolic Blood Pressure
  ## Load variables selected by LASSO
-lasso.coef.min <- read.csv("data/lasso_res_sbp_min.csv")
+lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_sbp_min.csv")
 lasso.coef.min$x
 names.lasso.min <- c("bmi", "hdl", "total_chol","AST",
                      "creatinine", "diabetes", "citizenship_status",
@@ -114,7 +114,7 @@ AIC(modlasso.dbp.min, modlasso.sbp.min)
 ## Best Subset
 ## Systolic
 ## Load selected coefficients
-reg.coef.sbp <- read.csv("data/reg.sbp.csv") #K=13
+reg.coef.sbp <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/reg.sbp.csv") #K=13
 reg.coef.sbp$X
 names.reg <- c("bmi","total_chol", "creatinine", "educ_level", "gender", "age_yr", 
                "marital_status","sleep", "race", "income_category")
@@ -130,7 +130,7 @@ sum((pred-dt$sbp)^2)/nrow(dt)
 ## Best Subset
 ## Diastolic
 ## Load selected coefficients
-reg.coef.dbp <- read.csv("data/reg.dbp.csv") #K=19
+reg.coef.dbp <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/reg.dbp.csv") #K=19
 reg.coef.dbp$X
 names.reg <- c("bmi","total_chol", "hemoglobin", "albumin", "AST", "diabetes",
                "children..5", "marital_status",
@@ -154,10 +154,10 @@ tab2.reg <- tbl_regression(modref.dbp, include = everything()) %>%
[…]

### 2023-12-09 manuscript/paper.Rmd
up

@@
 
-In order to select variables impacting blood pressure, we proceed by
-doing some variable selection with regularization techniques. We first
-employ best subset selection by using 10-fold cross validation to select
-the best number of subsets. For syastolic blood pressure, forward best
-subset selection choose 13 variables and for diastolic blood pressure 19
-variables. These variables are showed, along with its significance in
-Table \@ref(tab:tabreg)
+With the purpose of identifying influential variables impacting blood
+pressure, a variable selection process was conducted. The initial
+approach involved employing best subset selection with 10-fold
+cross-validation to ascertain the optimal number of subsets. Notably,
+for systolic blood pressure, the forward best subset selection method
+discerned 13 variables, whereas diastolic blood pressure exhibited 19
+selected variables. However, a limitation inherent in forward best
+subset selection is its inability to eliminate previously selected
+features, potentially disregarding their relevance in light of newly
+added variables. Consequently, this methodology may yield suboptimal
+variable selections due to its lack of adaptability.
 
-```{r tabref, echo=FALSE}
-tab.ref %>% as_kable_extra(booktabs = T, caption ="Coefficients for Foward Best Subset Selection") %>% kable_styling(latex_options = "scale_down")
-
-```
+In an effort to mitigate this limitation and introduce regularization,
+LASSO regularization was employed. This involved a 10-fold
+cross-validation process to determine the optimal penalty parameter,
+$\lambda$ . For diastolic blood pressure, the $\lambda$ value associated
+with the lowest mean squared error (MSE) led to the selection of 25
+variables. Conversely, when considering systolic blood pressure, the
+$\lambda$ value minimizing the MSE resulted in the exclusion of a sole
+variable. Subsequently, a $\lambda$ value was strategically chosen to
+ensure a negligible increase of no more than 1% in MSE, ultimately
+leading to the inclusion of 29 variables in the model.
 
@@
 
+### Foward Stepwise Best Subset
+
+```{=tex}
+\begin{align*}
+\text{sbp} = &\, 85.65 + 0.47 \times \text{BMI} + 1.49 \times \text{total\_chol} - 0.00 \times \text{creatinine} \\
+&+ 3.06 \times \text{educ\_levelHighschool} + 2.00 \times \text{educ\_levelLess12grade} + 1.31 \times \text{educ\_levelsomCollege} \\
+&+ 4.59 \times \text{genderMale} + 0.47 \times \text{age\_yr} + 2.05 \times \text{marital\_statusLiving with partner} \\
+&- 0.12 \times \text{marital\_statusMarried} + 3.30 \times \text{marital\_statusNever married} \\
+&- 0.99 \times \text{marital\_statusSeparated} + 7.01 \times \text{marital\_statusWidowed} - 1.30 \times \text{sleep>8hrs} \\
+&- 2.82 \times \text{sleep6-8hrs} - 4.60 \times \text{raceMexican American} - 3.78 \times \text{raceOther} \\
+&- 5.03 \times \text{raceWhite income\_categoryLower-middle income} - 3.02 \times \text{income\_categoryMiddle income} \\
+&- 2.88 \times \text{income\_categoryUnknown/Refused} - 2.70 \times \text{income\_categoryUpper-middle income} \\
+&- 4.30 \times \text{income\_categoryVaried/High income}
+\end{align*}
+```
+```{=tex}
+\begin{align*}
+\text{dbp} = &\, 53.55 + 0.22 \times \text{BMI} + 0.78 \times \text{total\_chol} + 0.85 \times \text{hemoglobin} + 0.02 \times \text{albumin} \\
+&+ 0.03 \times \text{AST} - 0.89 \times \text{diabetesNo} - 3.37 \times \text{diabetesYes} - 1.06 \times \text{children..51} \\
+&- 1.67 \times \text{children..52} - 1.83 \times \text{children..53 or more} - 1.90 \times \text{marital\_statusLiving with partner} \\
+&+ 0.00 \times \text{marital\_statusMarried} - 2.78 \times \text{marital\_statusNever married} + 0.91 \times \text{marital\_statusSeparated} \\
+&+ 1.72 \times \text{marital\_statusWidowed} - 2.40 \times \text{sleep>8hrs} - 2.37 \times \text{sleep6-8hrs} - 2.76 \times \text{raceMexican American} \\
+&+ 0.14 \times \text{raceOther} - 1.98 \times \text{raceWhite} - 2.16 \times \text{snortNot recorded} \\
+&- 0.57 \times \text{snortYes} - 2.33 \times \text{alcoholNo drinking} - 1.79 \times \text{alcoholNot recorded}
+\end{align*}
+```
+### LASSO
+
+```{=tex}
+\begin{align*}
+\text{sbp} = &\, 84.13 \times \text{(Intercept)} + 0.52 \times \text{bmi} + 0.79 \times \text{hdl} + 1.43 \times \text{total\_chol} + 0.04 \times \text{AST} \\
+&- 0.17 \times \text{diabetesNo} + 1.88 \times \text{diabetesYes} - 14.02 \times \text{citizenship\_statusDon't Know} \\
+&- 0.59 \times \text{citizenship\_statusNot a citizen of the US} - 12.06 \times \text{citizenship\_statusRefused} \\
+&+ 3.09 \times \text{educ\_levelHighschool} + 2.19 \times \text{educ\_levelLess12grade} + 1.16 \times \text{educ\_levelsomeCollege} \\
+&- 0.99 \times \text{children..51} - 1.17 \times \text{children..52} + 0.90 \times \text{children..53 or more} \\
+&+ 4.42 \times \text{genderMale} + 0.44 \times \text{age\_yr} + 1.77 \times \text{marital\_statusLiving with partner} \\
+&+ 0.01 \times \text{marital\_statusMarried} + 2.95 \times \text{marital\_statusNever married} - 1.05 \times \text{marital\_statusSeparated} \\
+&+ 7.19 \times \text{marital\_statusWidowed} - 1.47 \times \text{sleep>8hrs} - 2.95 \times \text{sleep6-8hrs} \\
+&- 4.11 \times \text{raceMexican American} - 3.43 \times \text{raceOther} - 5.14 \times \text{raceWhite} \\
+&+ 1.75 \times \text{smokeNot Smoking} + 0.22 \times \text{smokeSmoking} - 2.74 \times \text{snortNot recorded} \\
+&- 0.61 \times \text{snortYes} - 1.69 \times \text{alcoholNo drinking} - 0.28 \times \text{alcoholNot recorded} \\
+&- 1.62 \times \text{income\_categoryLower-middle income} - 3.04 \times \text{income\_categoryMiddle income} \\
+&- 2.64 \times \text{income\_categoryUnknown/Refused} - 2.71 \times \text{income\_categoryUpper-middle income} \\
+&- 4.54 \times \text{income\_categoryVaried/High income}
+\end{align*}
+```
+```{=tex}
+\begin{align*}
+\text{dbp} = &\, 47.99 + 0.24 \times \text{bmi} + 0.80 \times \text{total\_chol} + 0.58 \times \text{hemoglobin} + 0.17 \times \text{albumin} \\
+&+ 0.04 \times \text{AST} - 0.42 \times \text{diabetesNo} - 3.18 \times \text{diabetesYes} + 1.36 \times \text{citizenship\_statusDon't Know} \\
+&- 0.60 \times \text{citizenship\_statusNot a citizen of the US} - 1.78 \times \text{citizenship\_statusRefused} \\
+&- 0.11 \times \text{educ\_levelHighschool} - 0.92 \times \text{educ\_levelLess12grade} - 0.39 \times \text{educ\_levelsomeCollege} \\
+&- 0.99 \times \text{children..51} - 1.39 \times \text{children..52} - 2.08 \times \text{children..53 or more} \\
+&+ 0.59 \times \text{genderMale} + 0.01 \times \text{age\_yr} - 1.77 \times \text{marital\_statusLiving with partner} \\
+&+ 0.12 \times \text{marital\_statusMarried} - 2.67 \times \text{marital\_statusNever married} + 0.94 \times \text{marital\_statusSeparated} \\
+&+ 1.69 \times \text{marital\_statusWidowed} + 2.21 \times \text{cycle} - 2.21 \times \text{sleep>8hrs} \\
+&- 2.25 \times \text{sleep6-8hrs} - 2.08 \times \text{raceMexican American} + 0.31 \times \text{raceOther} \\
+&- 2.09 \times \text{raceWhite} + 0.44 \times \text{smokeNot Smoking} + 0.75 \times \text{smokeSmoking} \\
+&- 2.15 \times \text{snortNot recorded} - 0.50 \times \text{snortYes} - 1.28 \times \text{alcoholNo drinking} \\
+&- 1.17 \times \text{alcoholNot recorded}
+\end{align*}
+```
 # Discussion and conclusion

### 2023-12-09 manuscript/paper.Rmd
table

@@
 
+| Variable | Name                    | Description                                                                                                                                  |
+|-------------|---------------|---------------------------------------------|
+| SEQN     | sequence number         | Respondent number                                                                                                                            |
+| SDMVPSU  | psu                     | Masked variance unit pseudo-PSU variable for variance estimation                                                                             |
+| WTINT2YR | weights                 | Full sample 2-year interview weights                                                                                                         |
+| SDMVSTRA | strata                  | Masked variance unit pseudo-stratum variable for variance estimation                                                                         |
+| RIAGENDR | gender                  | Respondent's number                                                                                                                          |
+| RIDAGEYR | age (yrs)               | Respondent's age in years                                                                                                                    |
+| DMDMARTL | marital status          | Marital status                                                                                                                               |
+| INDFMIN2 | income category         | Total family income (reported as a range value in dollars)                                                                                   |
+| RIDRETH3 | race                    | Recode of reported race and Hispanic origin information, with Non-Hispanic Asian Category                                                    |
+| DMDHHSZA | children \<5            | Number of children aged 5 years or younger in the household                                                                                  |
+| DMDEDUC2 | education level         | What is the highest grade or level of school {you have/SP has} completed or the highest degree {you have/s/he has} received?                 |
+| DMDCITZN | citizenship status      | {Are you/Is SP} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv             |
+| SLD012   | sleep                   | Number of hours usually sleep on weekdays or workdays                                                                                        |
+| BMXBMI   | bmi                     | Body mass index                                                                                                                              |
+| ALQ121   | alcohol use             | In the past 12 months, how often did you drink any type of alcoholic beverage?                                                               |
+| LBDHDDSI | hdl                     | High density lipoprotein                                                                                                                     |
+| LBDSALSI | albumin                 | Albumin (g/L)                                                                                                                                |
+| DIQ010   | diabetes                | Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?                                         |
+| BPQ020   | hypertension            | {Have you/Has SP} ever been told by a doctor or other health professional that {you/s/he} had hypertension, also called high blood pressure? |
+| URXCRS   | creatinine              | Creatinine, urine (umol/L)                                                                                                                   |
+| SLQ040   | snort                   | In the past 12 months, how often did {you/SP} snort, gasp, or stop breathing while {you were/s/he was} asleep?                               |
+| LBDTCSI  | total cholestrol levels | Total cholesterol (mmol/L)                                                                                                                   |
+| LBXHGB   | hemoglobin              | Hemoglobin (g/dL)                                                                                                                            |
+| LBXSASSI | AST                     | Aspartate aminotransferase                                                                                                                   |
+| SMQ040   | smoke                   | Do you now smoke cigarettes?                                                                                                                 |
+| BPXDI    | blood pressure          | Systolic and diastolic blood pressure taken as the average of the first three measurements                                                   |
+
 ```{r tabone, echo=FALSE}
@@
 
-vars %>% kable(booktabs = T) %>% kable_styling(latex_options = "scale_down")
+#vars %>% kable(booktabs = T) %>% kable_styling(latex_options = "scale_down")
 

### 2023-12-10 data/scrip2.R
update

@@ -11,11 +11,11 @@ library(ggcorrplot)
 library(kableExtra)
 
 ## Load Data
-dt <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
+dt <- read.csv("../SleepBP/data/dt.csv")
 
 ## load the variable name file
-vars <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/vars.csv")
-
+vars <- read.csv("../SleepBP/data/vars.csv")
+dt$cycle <- as.factor(dt$cycle)
 ## some pre-processing
 dt <- dt %>% select(-X)
 dt <- dt %>%
@@ -58,10 +58,169 @@ tabsum <- dt %>%
   bold_labels() %>% 
   add_p()
 
+####################################
+## Statistical Modeling
+####################################
 
+
+###################################
+## UNIVARIATE ANALYSIS
+###################################
+## create the survey design object
+design = svydesign(id = ~psu, weights = ~weights, strata = ~strata, 
+                   nest = TRUE, survey.lonely.psu = "adjust", data = dt)
+
+
+##-------------------------------
+## Univariate regressions (1) sbp
+dt.sbp <- dt %>% select(-c("seq_no", "psu", "strata", "dbp", "weights", "hypertension"))
+sbp.univariate <- tbl_uvregression(data = dt.sbp, y = sbp, method = lm,
+                                   hide_n = TRUE,
+                                   label = list(bmi ~ "BMI",
+                              hdl ~ "HDL",
+                              total_chol ~ "TC",
+                              hemoglobin ~ "Hemoglobin",
+                              albumin ~ "Albumin",
+                              creatinine ~ "Creatinine",
+                              diabetes ~ "Diabetes",
+                              citizenship_status ~ "Citizenship",
+                              educ_level ~ "Education",
+                              children..5 ~ "Children > 5 yrs",
+                              age_yr ~ "Age (yrs)",
+                              marital_status ~ "Marital status",
+                              cycle ~ "Survey cycle",
+                              sleep ~ "Sleep",
+                              race ~ "Race",
+                              smoke ~ "Smoking status",
+                              snort ~ "Snort",
+                              alcohol ~ "Alcohol",
+                              income_category ~ "Income",
+                              gender ~ "Gender"
+                 )) %>% 
[…]

### 2023-12-10 data/scrip2.R
update

@@ -204,16 +204,21 @@ dbpfull
 merged.mult <- tbl_merge(tbls = list(sbpfull, dbpfull), tab_spanner = c("SBP", "DBP"))
 merged.mult
 
-## 
-## Interacted regressions including all two-way dependencies
+##------------------------------------
+## Interacted regressions including all two-way dependencies: saved only results where p < 0.05
 ## sbp
 int.sbp <- lm(sbp ~ .^2, data = dt.sbp)
-summary(int.sbp)
+coefficients <- summary(int.sbp)$coefficients
+coef_df <- as.data.frame(coefficients)
+significant_results <- coef_df[coef_df$`Pr(>|t|)` < 0.05, ]
+print(significant_results)
 
 ## dbp
 int.dbp <- lm(dbp ~ .^2, data = dt.dbp)
-summary(int.dbp)
-
+coefficients <- summary(int.dbp)$coefficients
+coef_df <- as.data.frame(coefficients)
+significant_results <- coef_df[coef_df$`Pr(>|t|)` < 0.05, ]
+print(significant_results)
 
 ##-------------------------------------
 ## Variable-selection based regressions

### 2023-12-10 manuscript/paper.Rmd
update

@@
 | Variable | Name                    | Description                                                                                                                                  |
-|-------------|---------------|---------------------------------------------|
+|-----------------|-----------------|--------------------------------------|
 | SEQN     | sequence number         | Respondent number                                                                                                                            |
@@
 features, potentially disregarding their relevance in light of newly
-added variables. Consequently, this methodology may yield suboptimal
+added variables. Consequently, this methodology may yield sub-optimal
 variable selections due to its lack of adaptability.
@@
 
-The univariate analysis of potential confounding factors is shown in
+The uni-variate analysis of potential confounding factors is shown in
 Table 2.

### 2023-12-10 manuscript/paper.Rmd
update

@@
 
--   importance of reproducibility
--   As a case on point ...?
--   Underscores the importance of reproducibility within the frame of
-    statistical analysis. We use the study by to ... The current study
-    builds on the paper by .. and starts in the manner of a replication
-    study and extends the analyses therein. We focus on statistical
-    analyses of these data and compare models based on varied
-    specification of functional forms of the covariates.
+The clinical importance of systolic blood pressure (SBP) stands as a
+cornerstone extensively acknowledged and documented across the continuum
+of medical literature's data collection. Elevated SBP stands as an
+precursor linked with a spectrum of critical health conditions, notably
+encompassing cardiovascular disease, stroke, and renal impairment, among
+others. This nexus between heightened SBP and these significant health
+adversities underscores the gravity of SBP as a predictive marker for
+adverse health outcomes. Noteworthy is the pivotal concern surrounding
+the trajectory of SBP in the natural aging process, a trajectory
+observed ubiquitously across diverse populations.
+
+Within Western societies aged 40 years and above, a demonstrable pattern
+emerges, revealing an approximate elevation of 7 mmHg in SBP per decade
+[**CITE**]. This discernible and consistent increment in SBP with
+advancing age accentuates its profound impact within the broader context
+of aging-related health dynamics. Concurrently, diastolic blood pressure
+(DBP), while demonstrating a concurrent rise correlating with age,
+presents a substantially lower rate of increase when juxtaposed against
+the ascending trajectory observed in SBP. Thus, in light of the
+significance attached to SBP dynamics within aging populations and the
+intricate interplay between age and blood pressure alterations, our
+investigative focus is oriented towards the exploration of variables
+that exert potential influence on systolic blood pressure (SBP). This
+deliberate focus aims to elucidate the multifaceted nature of factors
+contributing to SBP variability, enabling a more comprehensive
+understanding of its determinants within the framework of health and
+aging.
+
+Beyond age, the basal metabolic rate (BMI) emerges as a recognized
+contributor to systolic blood pressure (SBP), with individuals
+exhibiting higher BMI values demonstrating a propensity for elevated
+SBP. Gender also holds a discernible association with SBP dynamics.
+Empirical evidence in CITE demonstrates that women typically manifest
+lower SBP than men until the onset of menopause, after which women
+exhibit higher SBP levels compared to men. Additionally, heightened
+levels of cholesterol are closely linked to vascular damage. The
+accumulation of cholesterol deposits and calcium within arteries
+precipitates arterial stiffening, necessitating augmented cardiac
+efforts to circulate blood, potentially culminating in elevated blood
+pressure. This study endeavors to examine the interrelationships between
+age and cholesterol levels, emphasizing consideration of potential
+confounding variables. Table 1 comprehensively delineates the array of
+variables under scrutiny, ensuring a comprehensive analysis of their
+collective impact
 
@@
 | Variable | Name                    | Description                                                                                                                                  |
-|-----------------|-----------------|--------------------------------------|
+|------------------|------------------|-------------------------------------|
 | SEQN     | sequence number         | Respondent number                                                                                                                            |
@@
 # References
+
+\@article{Gurven_Blackwell_Rodríguez_Stieglitz_Kaplan_2012a, title={Does
+blood pressure inevitably rise with age?}, volume={60},
+url={[https://doi.org/10.1161/hypertensionaha.111.189100},](https://doi.org/10.1161/hypertensionaha.111.189100},)
+DOI={[10.1161/hypertensionaha.111.189100](https://doi.org/10.1161/hypertensionaha.111.189100)},
+number={1}, journal={Hypertension}, author={Gurven, Michael and
+Blackwell, Aaron D. and Rodríguez, Daniel and Stieglitz, Jonathan and
+Kaplan, Hillard}, year={2012}, month=jul, pages={25--33} }

### 2023-12-10 data/scrip2.R

@@ -11,10 +11,10 @@ library(ggcorrplot)
 library(kableExtra)
 
 ## Load Data
-dt <- read.csv("../SleepBP/data/dt.csv")
+dt <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
 
 ## load the variable name file
-vars <- read.csv("../SleepBP/data/vars.csv")
+vars <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/vars.csv")
 dt$cycle <- as.factor(dt$cycle)
 ## some pre-processing
 dt <- dt %>% select(-X)
@@ -225,7 +225,7 @@ print(significant_results)
 ##-------------------------------------
  ## Systolic Blood Pressure
  ## Load variables selected by LASSO
-lasso.coef.min <- read.csv("../SleepBP/data/lasso_res_sbp_min.csv")
+lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_sbp_min.csv")
 lasso.coef.min$x
 names.lasso.min <- c("bmi", "hdl", "total_chol","AST",
                      "creatinine", "diabetes", "citizenship_status",

### 2023-12-10 manuscript/paper.Rmd
update

--- 2023-12-09 manuscript/paper.Rmd
@@
 # source file 
-source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
+source(file = "../SleepBP/data/scrip2.R")
 ```
@@
 (n = 5521). Individuals that reported being on anti-hypertensive
-medication were also excluded from further analyses (n = 2944).
+medication were also excluded from further analyses (n = 2944). Our final analyses were based on a sample of n = 1977 observations on 28 covariates.
 
@@
 
-### Exposure:
+### Exposures:
 
@@
 following questions: "Number of hours usually sleep on weekdays or
-workdays". Sleep duration was divided into three groups, which were \< 6
-h, 6--8 h, ≥ 8 h respectively, of which 6--8 h was used as the reference
-group.
+workdays". We then categorized this variable into three groups, that is, \< 6
+h, 6--8 h, ≥ 8 h respectively, and used < 6 h as the reference
+group in our analysis. We also explored the association between total cholestrol (in mmol/L) level on systolic blood pressure including potential effect modification by other factors.
 

### 2023-12-10 data/scrip2.R

@@ -204,34 +204,96 @@ dbpfull
 merged.mult <- tbl_merge(tbls = list(sbpfull, dbpfull), tab_spanner = c("SBP", "DBP"))
 merged.mult
 
-##------------------------------------
-## Interacted regressions including all two-way dependencies: saved only results where p < 0.05
+
+##--------------------------------------
+## sbp full model with age and cholestrol interaction
+sbp.fullc <- lm(sbp ~. + total_chol:age_yr, data = dt.sbp)
+summary(sbp.fullc)
+sbpfullc <- tbl_regression(sbp.fullc,
+                          label = list(bmi ~ "BMI",
+                                       hdl ~ "HDL",
+                                       total_chol ~ "TC",
+                                       hemoglobin ~ "Hemoglobin",
+                                       albumin ~ "Albumin",
+                                       creatinine ~ "Creatinine",
+                                       #hypertension ~ "Hypertension",
+                                       diabetes ~ "Diabetes",
+                                       citizenship_status ~ "Citizenship",
+                                       educ_level ~ "Education",
+                                       children..5 ~ "Children > 5 yrs",
+                                       age_yr ~ "Age (yrs)",
+                                       marital_status ~ "Marital status",
+                                       cycle ~ "Survey cycle",
+                                       sleep ~ "Sleep",
+                                       race ~ "Race",
+                                       smoke ~ "Smoking status",
+                                       snort ~ "Snort",
+                                       alcohol ~ "Alcohol",
+                                       income_category ~ "Income",
+                                       gender ~ "Gender"
+                          )) %>% 
+  modify_header(label ~ "**Variable**") %>%
+  bold_labels()
+sbpfullc
+
+##--------------------------------------
[…]

### 2023-12-11 manuscript/paper.Rmd
update

--- 2023-12-10 manuscript/paper.Rmd
@@
 ---
-title: "Preparing the NHANES Data for the Replication of: Association between sleep duration on workdays and blood pressure in non‑overweight/obese population in NHANES: a public database research"
+title: "Weekday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population"
 thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
-subtitle: "Short title goes here!"
+#subtitle: "Short title goes here!"
 author:
 - Amos Okutse, Brown University
-- Second Author, Another Uni
-- Graduate Student, Graduate School
+- Monic Colon-Vargas, Brown University
 abstract: "**Introduction:** Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec sit amet libero justo. Pellentesque eget nibh ex. Aliquam tincidunt egestas lectus id ullamcorper.
@@
 urlcolor: blue
-#bibliography: refs.bib
+bibliography: bibfile.bib
 bibliostyle: apalike
@@
 # source file 
-source(file = "../SleepBP/data/scrip2.R")
+# source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
+source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
 ```
@@
 
--   importance of reproducibility
--   As a case on point ...?
--   Underscores the importance of reproducibility within the frame of
-    statistical analysis. We use the study by to ... The current study
-    builds on the paper by .. and starts in the manner of a replication
-    study and extends the analyses therein. We focus on statistical
-    analyses of these data and compare models based on varied
-    specification of functional forms of the covariates.
+The clinical importance of systolic blood pressure (SBP) stands as a
+cornerstone extensively acknowledged and documented across the continuum
+of medical literature's data collection [@gurven2012does; @su2022association]. Elevated SBP stands as a
+precursor linked with a spectrum of critical health conditions, 
+encompassing cardiovascular diseases, stroke, and renal impairment, among
+others. This nexus between heightened SBP and these significant health
+adversities underscores the gravity of SBP as a predictive marker for
+adverse health outcomes. Noteworthy is the pivotal concern surrounding
+the trajectory of SBP in the natural aging process, a trajectory
+observed ubiquitously across diverse populations.
+
+In western societies aged 40 years and above, a demonstrable pattern
+emerges, revealing an approximate elevation of 7 mmHg in SBP per decade among individuals above 40 years. 
+This discernible and consistent increment in SBP with
+advancing age accentuates its profound impact within the broader context
+of aging-related health dynamics. Similarly, while diastolic blood pressure
+(DBP), demonstrates a concurrent rise correlating with age,
+presents a substantially lower rate of increase when juxtaposed against
+the ascending trajectory observed in SBP [@gurven2012does; @su2022association]. 
+Thus, in light of the
+significance attached to SBP dynamics within aging populations and the
+intricate interplay between age, sleep, cholestrol levels and blood pressure alterations, our
+investigative focus is oriented towards the exploration of variables
+that exert potential influence on systolic blood pressure (SBP). This
+deliberate focus aims to elucidate the multifaceted nature of factors
+contributing to SBP variability, enabling a more comprehensive
+understanding of its determinants within the framework of health and
+aging.
+
+<!-- Beyond age, the basal metabolic rate (BMI) emerges as a recognized -->
+<!-- contributor to systolic blood pressure (SBP), with individuals -->
+<!-- exhibiting higher BMI values demonstrating a propensity for elevated -->
+<!-- SBP. Gender also holds a discernible association with SBP dynamics. -->
+<!-- Empirical evidence in CITE demonstrates that women typically manifest -->
+<!-- lower SBP than men until the onset of menopause, after which women -->
+<!-- exhibit higher SBP levels compared to men. Additionally, heightened -->
+<!-- levels of cholesterol are closely linked to vascular damage. The -->
+<!-- accumulation of cholesterol deposits and calcium within arteries -->
+<!-- precipitates arterial stiffening, necessitating augmented cardiac -->
+<!-- efforts to circulate blood, potentially culminating in elevated blood -->
+<!-- pressure. This study endeavors to examine the interrelationships between -->
+<!-- age and cholesterol levels, emphasizing consideration of potential -->
+<!-- confounding variables. Table 1 comprehensively delineates the array of -->
+<!-- variables under scrutiny, ensuring a comprehensive analysis of their -->
+<!-- collective impact -->
 
@@
 (n = 5521). Individuals that reported being on anti-hypertensive
-medication were also excluded from further analyses (n = 2944). Our final analyses were based on a sample of n = 1977 observations on 28 covariates.
+medication were also excluded from further analyses (n = 2944). 
+Our final analyses were based on a sample of n = 1977 observations on 28 covariates.
 
@@
 
-| Variable | Name                    | Description                                                                                                                                  |
-|-------------|---------------|---------------------------------------------|
-| SEQN     | sequence number         | Respondent number                                                                                                                            |
-| SDMVPSU  | psu                     | Masked variance unit pseudo-PSU variable for variance estimation                                                                             |
-| WTINT2YR | weights                 | Full sample 2-year interview weights                                                                                                         |
-| SDMVSTRA | strata                  | Masked variance unit pseudo-stratum variable for variance estimation                                                                         |
-| RIAGENDR | gender                  | Respondent's number                                                                                                                          |
-| RIDAGEYR | age (yrs)               | Respondent's age in years                                                                                                                    |
-| DMDMARTL | marital status          | Marital status                                                                                                                               |
-| INDFMIN2 | income category         | Total family income (reported as a range value in dollars)                                                                                   |
-| RIDRETH3 | race                    | Recode of reported race and Hispanic origin information, with Non-Hispanic Asian Category                                                    |
-| DMDHHSZA | children \<5            | Number of children aged 5 years or younger in the household                                                                                  |
-| DMDEDUC2 | education level         | What is the highest grade or level of school {you have/SP has} completed or the highest degree {you have/s/he has} received?                 |
-| DMDCITZN | citizenship status      | {Are you/Is SP} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv             |
-| SLD012   | sleep                   | Number of hours usually sleep on weekdays or workdays                                                                                        |
-| BMXBMI   | bmi                     | Body mass index                                                                                                                              |
-| ALQ121   | alcohol use             | In the past 12 months, how often did you drink any type of alcoholic beverage?                                                               |
-| LBDHDDSI | hdl                     | High density lipoprotein                                                                                                                     |
-| LBDSALSI | albumin                 | Albumin (g/L)                                                                                                                                |
-| DIQ010   | diabetes                | Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?                                         |
-| BPQ020   | hypertension            | {Have you/Has SP} ever been told by a doctor or other health professional that {you/s/he} had hypertension, also called high blood pressure? |
-| URXCRS   | creatinine              | Creatinine, urine (umol/L)                                                                                                                   |
-| SLQ040   | snort                   | In the past 12 months, how often did {you/SP} snort, gasp, or stop breathing while {you were/s/he was} asleep?                               |
-| LBDTCSI  | total cholestrol levels | Total cholesterol (mmol/L)                                                                                                                   |
-| LBXHGB   | hemoglobin              | Hemoglobin (g/dL)                                                                                                                            |
-| LBXSASSI | AST                     | Aspartate aminotransferase                                                                                                                   |
-| SMQ040   | smoke                   | Do you now smoke cigarettes?                                                                                                                 |
-| BPXDI    | blood pressure          | Systolic and diastolic blood pressure taken as the average of the first three measurements                                                   |
+\begin{table}[H]
 
-```{r tabone, echo=FALSE}
-#tab.one <- as_kable_extra(vars, caption = "Variable names and descriptions", 
-#                 format = "latex", booktabs = TRUE) %>% 
-#  kable_styling(latex_options = c("HOLD_position") %>% 
-#  row_spec(c(0), bold = TRUE) %>% 
-#  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>% 
-#  column_spec(3, width = "8cm")
-#tab.one
+\caption{\label{tab:tabone}Variable descriptions}
+\centering
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{8cm}}
+\toprule
+\textbf{Variable} & \textbf{Name} & \textbf{Description}\\
+\midrule
+\textbf{SEQN} & \textbf{sequence number} & Respondent number\\
+\textbf{SDMVPSU} & \textbf{psu} & Masked variance unit pseudo-PSU variable for variance estimation\\
+\textbf{WTINT2YR} & \textbf{weights} & Full sample 2-year interview weights\\
+\textbf{SDMVSTRA} & \textbf{strata} & Masked variance unit pseudo-stratum variable for variance estimation\\
+\textbf{RIAGENDR} & \textbf{gender} & Respondent's number\\
+\addlinespace
+\textbf{RIDAGEYR} & \textbf{age (yrs)} & Respondent's age in years\\
+\textbf{DMDMARTL} & \textbf{marital status} & Marital status\\
+\textbf{INDFMIN2} & \textbf{income category} & Total family income (reported as a range value in dollars)\\
+\textbf{RIDRETH3} & \textbf{race} & Recode of reported race and Hispanic origin information, with Non-Hispanic Asian Category\\
+\textbf{DMDHHSZA} & \textbf{children <5} & Number of children aged 5 years or younger in the household\\
+\addlinespace
+\textbf{DMDEDUC2} & \textbf{education level} & What is the highest grade or level of school\{you have/SP has\} completed or the highest degree \{you have/s/he has\} received?\\
+\textbf{DMDCITZN} & \textbf{citizenship status} & \{Are you/Is SP\} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv\\
+\textbf{SLD012} & \textbf{sleep} & Number of hours usually sleep on weekdays or workdays\\
+\textbf{BMXBMI} & \textbf{bmi} & Body mass index\\
+\textbf{ALQ121} & \textbf{alcohol use} & In the past 12 months, how often did you drink any type of alcoholic beverage?\\
+\addlinespace
+\textbf{LBDHDDSI} & \textbf{hdl} & High density lipoprotein\\
+\textbf{LBDSALSI} & \textbf{albumin} & Albumin (g/L)\\
+\textbf{DIQ010} & \textbf{diabetes} & Have you ever been told by a doctor or health professional that you have diabetes or sugar diabetes?\\
+\textbf{BPQ020} & \textbf{hypertension} & \{Have you/Has SP\} ever been told by a doctor or other health professional that \{you/s/he\} had hypertension, also called high blood pressure?\\
+\textbf{URXCRS} & \textbf{creatinine} & Creatinine, urine (umol/L)\\
+\addlinespace
+\textbf{SLQ040} & \textbf{snort} & In the past 12 months, how often did \{you/SP\} snort, gasp, or stop breathing while \{you were/s/he was\} asleep?\\
+\textbf{LBDTCSI} & \textbf{total cholestrol levels} & Total cholesterol (mmol/L)\\
+\textbf{LBXHGB} & \textbf{hemoglobin} & Hemoglobin (g/dL)\\
+\textbf{LBXSASSI} & \textbf{AST} & Aspertate aminotransferase\\
+\textbf{SMQ040} & \textbf{smoke} & Do you now smoke cigarettes?\\
+\addlinespace
+\textbf{BPXDI} & \textbf{blood pressure} & Systolic and diastolic blood pressure taken as the average of the first three measurements\\
+\bottomrule
+\end{tabular}
+\end{table}
 
-#vars %>% kable(booktabs = T) %>% kable_styling(latex_options = "scale_down")
-
+```{r tabone, echo=FALSE, eval=FALSE}
+tab.one <- vars %>% 
+  kable(format = "latex", caption = "Variable descriptions", booktabs = TRUE) %>% 
+  kable_styling(latex_options = c("HOLD_position")) %>%
+  row_spec(c(0), bold = TRUE) %>%
+  column_spec(c(1, 2), bold = TRUE, width = "2cm") %>%
+  column_spec(3, width = "8cm")
+tab.one
 ```
@@
 
+## Model Evaluations
+
+Table \@ref(tab:tabmodel) summarizes the performance metrics across the different models fitted to the data in our investigation.
+
+\begin{table}
+
+\caption{\label{tab:tabmodel}Summary model performance measures.}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+  & df & AIC & BIC & MSE & $R^2$ & Adjusted $R^2$\\
+\midrule
+Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
+Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
+Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
+Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
+Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
+\addlinespace
+Model 6 & 41 & 15888.38 & 16117.55 & 173.69 & 0.36 & 0.34\\
+Model 7 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
+\bottomrule
+\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
+\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
+\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
+\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
+\end{tabular}
+\end{table}
+
+```{r tabmodel, eval=FALSE, echo=FALSE}
+## Model performance
+tabmodel <- model.checks %>% 
+  kable(format = "latex", caption = "Summary model performance measures.", digits = 2, booktabs = TRUE, col.names = c("df", "AIC", "BIC", "MSE", "R^2", "Adjusted R2"), escape = FALSE) %>% 
+  add_footnote(label = c("df = degrees of freedom", "AIC = Akaike Information Criteria", "BIC = Bayesian Information Criteria", "MSE = Mean Squared Error"))
+tabmodel
+```
+
 # Results
 
+## Descriptive statistics
+
 Table \@ref(tab:tabtwo) summarizes participant characteristics in the
-2015 - 2018 NHANES survey cycles stratified by gender. Among the
-participants, the proportion of males and females were 44.81% (n = 1378)
-and 55.19% (n = 1509), respectively. With ethnicity, the proportion of
-Mexican American, White and Black were 6.27%, 64.62% and 9.94%,
+2015 - 2018 NHANES survey cycles stratified by gender.
+The 2015-2016 survey cycle comprised 53.77% of the total analyzed sample whereas the 2017 - 2018 survey cycle comprised 46.23% of the total analyzed sample.
+Among the
+participants, the proportion of males and females were 53.76% (n = 1063)
+and 46.23% (n = 914), respectively. With race, the proportion of
+Mexican American, White and Black were 9.16%, 34.90% and 17.10%,
 respectively. Overall, the mean (SD) values for age, albumin, 
-creatinine, SBP, DBP, hemoglobin, TC, AST, HDL, BMI were 38.54 (17.72)
-years, 43.48 (3.58) g/L, 73.21 (19.73) umol/L, 115.33 (14.56) mmHg,
-68.47 (10.08) mmHg, 14.07(1.44)g/dL, 4.67 (1.01) mmol/L, 23.18 (13.01)
-IU/L, 1.59 (0.43) mmol/L, and 21.97(2.06) kg/m2, respectively. Among the
-participants, 66.01% were consumed alcohol drinkers, 2.61% were
-diabetes, 3.16% were hypertension, 6.72% have snort or stop breathing.
-17.99% were smokers. Sleep duration was divided into three groups, which
-were \< 6 h, 6--8 h, ≥ 8 h, each with a proportion of 6.26%, 40.53%,
-53.21%, respectively.
+SBP, DBP, hemoglobin, TC, AST, HDL, BMI were 43.79 (17.43)
+years, 43.04 (3.59) g/L, 118.61 (16.41) mmHg,
+69.60 (11.18) mmHg, 13.94 (1.46)g/dL, 4.81 (1.02) mmol/L, 23.47 (13.17)
+IU/L, 1.60 (0.43) mmol/L, and 22.08 (2.03) kg/m2, respectively. 
+Among the
+participants, 73.70% were alcohol drinkers, 5.06% were
+diabetic, 7.79% were hypertensive, 13.76% experienced snorting whereas
+21.60% were smokers. Sleep duration was divided into three groups, which
+were \< 6 h, 6--8 h, ≥ 8 h, each with a proportion of 7.89%, 40.06%,
+52.05%, respectively. Male and female participants differed significantly on all characteristics
+except citizenship status, and the cycle of the survey.
 
-Male and female participants differed significantly on characteristics
-including
-
-The univariate analysis of potential confounding factors is shown in
-Table 2.
-
-```{r tabtwo, echo=FALSE}
-#tabsum %>% 
-#  as_kable_extra(format = "latex", caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", digits = 2, longtable = T) %>% 
-#  kable_styling(latex_options = c("HOLD_position"))
-
+```{r tabtwo, echo=FALSE, message=FALSE, warning=FALSE}
 tabsum %>%
   as_kable_extra(booktabs = TRUE, 
-                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.", 
-                 longtable = TRUE) %>%
-  kable_styling(latex_options = "scale_down")
+                 caption = "Unweighted summary characteristics of the participants in the NHANES 2015 - 2018 survey cycles stratified by gender.
+BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase.
+SBP = Systolic blood pressure, DBP = Diastolic blood pressure.",
+                 longtable = TRUE)
 ```
 
-### Foward Stepwise Best Subset
+## Univariate regression analysis
 
-```{=tex}
-\begin{align*}
-\text{sbp} = &\, 85.65 + 0.47 \times \text{BMI} + 1.49 \times \text{total\_chol} - 0.00 \times \text{creatinine} \\
-&+ 3.06 \times \text{educ\_levelHighschool} + 2.00 \times \text{educ\_levelLess12grade} + 1.31 \times \text{educ\_levelsomCollege} \\
-&+ 4.59 \times \text{genderMale} + 0.47 \times \text{age\_yr} + 2.05 \times \text{marital\_statusLiving with partner} \\
-&- 0.12 \times \text{marital\_statusMarried} + 3.30 \times \text{marital\_statusNever married} \\
-&- 0.99 \times \text{marital\_statusSeparated} + 7.01 \times \text{marital\_statusWidowed} - 1.30 \times \text{sleep>8hrs} \\
-&- 2.82 \times \text{sleep6-8hrs} - 4.60 \times \text{raceMexican American} - 3.78 \times \text{raceOther} \\
-&- 5.03 \times \text{raceWhite income\_categoryLower-middle income} - 3.02 \times \text{income\_categoryMiddle income} \\
-&- 2.88 \times \text{income\_categoryUnknown/Refused} - 2.70 \times \text{income\_categoryUpper-middle income} \\
-&- 4.30 \times \text{income\_categoryVaried/High income}
-\end{align*}
+Table \@ref(tab:tabthree) presents the summary of univariate regression analysis results exploring the association between the exposures, selected potential confounding variables and systolic blood pressure.
+We found substantial associations between ...
+
+```{r tabthree, echo=FALSE}
+## univariate analysis
+sbp.univariate %>% 
+  as_kable_extra(format = "latex", booktabs = TRUE, digits = 2, caption = "Univariate analyses of exposures, potential confounding factors and systolic blood pressure.
+BMI = Body Mass Index; HDL = High Density Lipoprotein, TC = Total Cholestrol level; AST = aspertate aminotransferase.", longtable = TRUE)
 ```
-```{=tex}
-\begin{align*}
-\text{dbp} = &\, 53.55 + 0.22 \times \text{BMI} + 0.78 \times \text{total\_chol} + 0.85 \times \text{hemoglobin} + 0.02 \times \text{albumin} \\
-&+ 0.03 \times \text{AST} - 0.89 \times \text{diabetesNo} - 3.37 \times \text{diabetesYes} - 1.06 \times \text{children..51} \\
-&- 1.67 \times \text{children..52} - 1.83 \times \text{children..53 or more} - 1.90 \times \text{marital\_statusLiving with partner} \\
-&+ 0.00 \times \text{marital\_statusMarried} - 2.78 \times \text{marital\_statusNever married} + 0.91 \times \text{marital\_statusSeparated} \\
-&+ 1.72 \times \text{marital\_statusWidowed} - 2.40 \times \text{sleep>8hrs} - 2.37 \times \text{sleep6-8hrs} - 2.76 \times \text{raceMexican American} \\
-&+ 0.14 \times \text{raceOther} - 1.98 \times \text{raceWhite} - 2.16 \times \text{snortNot recorded} \\
-&- 0.57 \times \text{snortYes} - 2.33 \times \text{alcoholNo drinking} - 1.79 \times \text{alcoholNot recorded}
-\end{align*}
+
+## Multivariable Regression Analysis
+
+```{r tabfour, echo=FALSE}
+## Final selected multiple linear regression model
+tbl_regression(adj.sbp.int2,
+               label = list(bmi ~ "BMI",
+                            #hdl ~ "HDL",
+                            total_chol ~ "TC",
+                            #hemoglobin ~ "Hemoglobin",
+                            albumin ~ "Albumin",
+                            creatinine ~ "Creatinine",
+                            #hypertension ~ "Hypertension",
+                            #diabetes ~ "Diabetes",
+                            citizenship_status ~ "Citizenship",
+                            educ_level ~ "Education",
+                            children..5 ~ "Children > 5 yrs",
+                            age_yr ~ "Age (yrs)",
+                            marital_status ~ "Marital status",
+                            #cycle ~ "Survey cycle",
+                            sleep ~ "Sleep",
+                            race ~ "Race",
+                            smoke ~ "Smoking status",
+                            snort ~ "Snort",
+                            alcohol ~ "Alcohol",
+                            income_category ~ "Income",
+                            gender ~ "Gender"
+               )) %>% modify_header(label ~ "**Variable**") %>%
+  bold_labels() %>% 
+  as_kable_extra(format = "latex", digits = 2, caption = "Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP).
+BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)
+
 ```
-### LASSO
 
-```{=tex}
-\begin{align*}
-\text{sbp} = &\, 84.13 \times \text{(Intercept)} + 0.52 \times \text{bmi} + 0.79 \times \text{hdl} + 1.43 \times \text{total\_chol} + 0.04 \times \text{AST} \\
-&- 0.17 \times \text{diabetesNo} + 1.88 \times \text{diabetesYes} - 14.02 \times \text{citizenship\_statusDon't Know} \\
-&- 0.59 \times \text{citizenship\_statusNot a citizen of the US} - 12.06 \times \text{citizenship\_statusRefused} \\
-&+ 3.09 \times \text{educ\_levelHighschool} + 2.19 \times \text{educ\_levelLess12grade} + 1.16 \times \text{educ\_levelsomeCollege} \\
-&- 0.99 \times \text{children..51} - 1.17 \times \text{children..52} + 0.90 \times \text{children..53 or more} \\
-&+ 4.42 \times \text{genderMale} + 0.44 \times \text{age\_yr} + 1.77 \times \text{marital\_statusLiving with partner} \\
-&+ 0.01 \times \text{marital\_statusMarried} + 2.95 \times \text{marital\_statusNever married} - 1.05 \times \text{marital\_statusSeparated} \\
-&+ 7.19 \times \text{marital\_statusWidowed} - 1.47 \times \text{sleep>8hrs} - 2.95 \times \text{sleep6-8hrs} \\
-&- 4.11 \times \text{raceMexican American} - 3.43 \times \text{raceOther} - 5.14 \times \text{raceWhite} \\
-&+ 1.75 \times \text{smokeNot Smoking} + 0.22 \times \text{smokeSmoking} - 2.74 \times \text{snortNot recorded} \\
-&- 0.61 \times \text{snortYes} - 1.69 \times \text{alcoholNo drinking} - 0.28 \times \text{alcoholNot recorded} \\
-&- 1.62 \times \text{income\_categoryLower-middle income} - 3.04 \times \text{income\_categoryMiddle income} \\
-&- 2.64 \times \text{income\_categoryUnknown/Refused} - 2.71 \times \text{income\_categoryUpper-middle income} \\
-&- 4.54 \times \text{income\_categoryVaried/High income}
-\end{align*}
-```
-```{=tex}
-\begin{align*}
-\text{dbp} = &\, 47.99 + 0.24 \times \text{bmi} + 0.80 \times \text{total\_chol} + 0.58 \times \text{hemoglobin} + 0.17 \times \text{albumin} \\
-&+ 0.04 \times \text{AST} - 0.42 \times \text{diabetesNo} - 3.18 \times \text{diabetesYes} + 1.36 \times \text{citizenship\_statusDon't Know} \\
-&- 0.60 \times \text{citizenship\_statusNot a citizen of the US} - 1.78 \times \text{citizenship\_statusRefused} \\
-&- 0.11 \times \text{educ\_levelHighschool} - 0.92 \times \text{educ\_levelLess12grade} - 0.39 \times \text{educ\_levelsomeCollege} \\
-&- 0.99 \times \text{children..51} - 1.39 \times \text{children..52} - 2.08 \times \text{children..53 or more} \\
-&+ 0.59 \times \text{genderMale} + 0.01 \times \text{age\_yr} - 1.77 \times \text{marital\_statusLiving with partner} \\
-&+ 0.12 \times \text{marital\_statusMarried} - 2.67 \times \text{marital\_statusNever married} + 0.94 \times \text{marital\_statusSeparated} \\
-&+ 1.69 \times \text{marital\_statusWidowed} + 2.21 \times \text{cycle} - 2.21 \times \text{sleep>8hrs} \\
-&- 2.25 \times \text{sleep6-8hrs} - 2.08 \times \text{raceMexican American} + 0.31 \times \text{raceOther} \\
-&- 2.09 \times \text{raceWhite} + 0.44 \times \text{smokeNot Smoking} + 0.75 \times \text{smokeSmoking} \\
-&- 2.15 \times \text{snortNot recorded} - 0.50 \times \text{snortYes} - 1.28 \times \text{alcoholNo drinking} \\
-&- 1.17 \times \text{alcoholNot recorded}
-\end{align*}
-```
+## Subgroup analysis by gender
+
 # Discussion and conclusion

### 2023-12-11 data/scrip2.R

@@ -11,10 +11,13 @@ library(ggcorrplot)
 library(kableExtra)
 
 ## Load Data
-dt <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
+## dt <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
+dt <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/dt.csv")
 
 ## load the variable name file
-vars <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/vars.csv")
+## vars <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/vars.csv")
+## vars <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/vars.csv")
+
 dt$cycle <- as.factor(dt$cycle)
 ## some pre-processing
 dt <- dt %>% select(-X)
@@ -22,7 +25,8 @@ dt <- dt %>%
   mutate_if(is.character, as.factor)
 
 ## edit the citizenship levels: Citizen, not citizen/other
-## Marital status: single/never married, married or living with partner, divorced or separated, widowed 
+
+levels(dt$citizenship_status) <- c("Citizen", "Unknown", "Non-citizen", "Refused")
 
 ## summarize the data by gender
 tabsum <- dt %>%  
@@ -57,6 +61,7 @@ tabsum <- dt %>%
   modify_header(label ~ "**Variable**") %>%
   bold_labels() %>% 
   add_p()
+tabsum
 
 ####################################
 ## Statistical Modeling
@@ -103,37 +108,37 @@ sbp.univariate
 
 
 ## Univariate regressions (2) dbp
-dt.dbp <- dt %>% select(-c("seq_no", "psu", "strata", "sbp", "weights", "hypertension"))
-dbp.univariate <- tbl_uvregression(data = dt.dbp, y = dbp, method = lm,
-                                   hide_n = TRUE,
-                                   label = list(bmi ~ "BMI",
-                                                hdl ~ "HDL",
-                                                total_chol ~ "TC",
-                                                hemoglobin ~ "Hemoglobin",
-                                                albumin ~ "Albumin",
-                                                creatinine ~ "Creatinine",
-                                                #hypertension ~ "Hypertension",
-                                                diabetes ~ "Diabetes",
[…]

### 2023-12-11 manuscript/paper.Rmd
upd

@@
 # source file 
-# source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
-source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
+source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
+
+#source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
 ```
@@
 cornerstone extensively acknowledged and documented across the continuum
-of medical literature's data collection [@gurven2012does; @su2022association]. Elevated SBP stands as a
-precursor linked with a spectrum of critical health conditions, 
-encompassing cardiovascular diseases, stroke, and renal impairment, among
-others. This nexus between heightened SBP and these significant health
-adversities underscores the gravity of SBP as a predictive marker for
-adverse health outcomes. Noteworthy is the pivotal concern surrounding
-the trajectory of SBP in the natural aging process, a trajectory
-observed ubiquitously across diverse populations.
+of medical literature's data collection [@gurven2012does;
+@su2022association]. Elevated SBP stands as a precursor linked with a
+spectrum of critical health conditions, encompassing cardiovascular
+diseases, stroke, and renal impairment, among others. This nexus between
+heightened SBP and these significant health adversities underscores the
+gravity of SBP as a predictive marker for adverse health outcomes.
+Noteworthy is the pivotal concern surrounding the trajectory of SBP in
+the natural aging process, a trajectory observed ubiquitously across
+diverse populations.
 
 In western societies aged 40 years and above, a demonstrable pattern
-emerges, revealing an approximate elevation of 7 mmHg in SBP per decade among individuals above 40 years. 
-This discernible and consistent increment in SBP with
-advancing age accentuates its profound impact within the broader context
-of aging-related health dynamics. Similarly, while diastolic blood pressure
-(DBP), demonstrates a concurrent rise correlating with age,
-presents a substantially lower rate of increase when juxtaposed against
-the ascending trajectory observed in SBP [@gurven2012does; @su2022association]. 
-Thus, in light of the
+emerges, revealing an approximate elevation of 7 mmHg in SBP per decade
+among individuals above 40 years. This discernible and consistent
+increment in SBP with advancing age accentuates its profound impact
+within the broader context of aging-related health dynamics. Similarly,
+while diastolic blood pressure (DBP), demonstrates a concurrent rise
+correlating with age, presents a substantially lower rate of increase
+when juxtaposed against the ascending trajectory observed in SBP
+[@gurven2012does; @su2022association]. Thus, in light of the
 significance attached to SBP dynamics within aging populations and the
-intricate interplay between age, sleep, cholestrol levels and blood pressure alterations, our
-investigative focus is oriented towards the exploration of variables
-that exert potential influence on systolic blood pressure (SBP). This
-deliberate focus aims to elucidate the multifaceted nature of factors
-contributing to SBP variability, enabling a more comprehensive
-understanding of its determinants within the framework of health and
-aging.
+intricate interplay between age, sleep, cholestrol levels and blood
+pressure alterations, our investigative focus is oriented towards the
+exploration of variables that exert potential influence on systolic
+blood pressure (SBP). This deliberate focus aims to elucidate the
+multifaceted nature of factors contributing to SBP variability, enabling
+a more comprehensive understanding of its determinants within the
+framework of health and aging.
 
 <!-- Beyond age, the basal metabolic rate (BMI) emerges as a recognized -->
+
 <!-- contributor to systolic blood pressure (SBP), with individuals -->
+
 <!-- exhibiting higher BMI values demonstrating a propensity for elevated -->
+
 <!-- SBP. Gender also holds a discernible association with SBP dynamics. -->
+
 <!-- Empirical evidence in CITE demonstrates that women typically manifest -->
+
 <!-- lower SBP than men until the onset of menopause, after which women -->
+
 <!-- exhibit higher SBP levels compared to men. Additionally, heightened -->
+
 <!-- levels of cholesterol are closely linked to vascular damage. The -->
+
 <!-- accumulation of cholesterol deposits and calcium within arteries -->
+
 <!-- precipitates arterial stiffening, necessitating augmented cardiac -->
+
 <!-- efforts to circulate blood, potentially culminating in elevated blood -->
+
 <!-- pressure. This study endeavors to examine the interrelationships between -->
+
 <!-- age and cholesterol levels, emphasizing consideration of potential -->
+
 <!-- confounding variables. Table 1 comprehensively delineates the array of -->
+
 <!-- variables under scrutiny, ensuring a comprehensive analysis of their -->
+
 <!-- collective impact -->
@@
 (n = 5521). Individuals that reported being on anti-hypertensive
-medication were also excluded from further analyses (n = 2944). 
-Our final analyses were based on a sample of n = 1977 observations on 28 covariates.
+medication were also excluded from further analyses (n = 2944). Our
+final analyses were based on a sample of n = 1977 observations on 28
+covariates.
 
@@
 following questions: "Number of hours usually sleep on weekdays or
-workdays". We then categorized this variable into three groups, that is, \< 6
-h, 6--8 h, ≥ 8 h respectively, and used < 6 h as the reference
-group in our analysis. We also explored the association between total cholestrol (in mmol/L) level on systolic blood pressure including potential effect modification by other factors.
+workdays". We then categorized this variable into three groups, that is,
+\< 6 h, 6--8 h, ≥ 8 h respectively, and used \< 6 h as the reference
+group in our analysis. We also explored the association between total
+cholestrol (in mmol/L) level on systolic blood pressure including
+potential effect modification by other factors.
 
@@
 
+```{=tex}
 \begin{table}[H]
@@
 \end{table}
-
+```
 ```{r tabone, echo=FALSE, eval=FALSE}
@@
 
-Table \@ref(tab:tabmodel) summarizes the performance metrics across the different models fitted to the data in our investigation.
+Table \@ref(tab:tabmodel) summarizes the performance metrics across the
+different models fitted to the data in our investigation.
 
+```{=tex}
 \begin{table}
@@
 \end{table}
-
+```
 ```{r tabmodel, eval=FALSE, echo=FALSE}
@@
 Table \@ref(tab:tabtwo) summarizes participant characteristics in the
-2015 - 2018 NHANES survey cycles stratified by gender.
-The 2015-2016 survey cycle comprised 53.77% of the total analyzed sample whereas the 2017 - 2018 survey cycle comprised 46.23% of the total analyzed sample.
-Among the
-participants, the proportion of males and females were 53.76% (n = 1063)
-and 46.23% (n = 914), respectively. With race, the proportion of
-Mexican American, White and Black were 9.16%, 34.90% and 17.10%,
-respectively. Overall, the mean (SD) values for age, albumin, 
-SBP, DBP, hemoglobin, TC, AST, HDL, BMI were 43.79 (17.43)
-years, 43.04 (3.59) g/L, 118.61 (16.41) mmHg,
-69.60 (11.18) mmHg, 13.94 (1.46)g/dL, 4.81 (1.02) mmol/L, 23.47 (13.17)
-IU/L, 1.60 (0.43) mmol/L, and 22.08 (2.03) kg/m2, respectively. 
-Among the
-participants, 73.70% were alcohol drinkers, 5.06% were
-diabetic, 7.79% were hypertensive, 13.76% experienced snorting whereas
-21.60% were smokers. Sleep duration was divided into three groups, which
-were \< 6 h, 6--8 h, ≥ 8 h, each with a proportion of 7.89%, 40.06%,
-52.05%, respectively. Male and female participants differed significantly on all characteristics
-except citizenship status, and the cycle of the survey.
+2015 - 2018 NHANES survey cycles stratified by gender. The 2015-2016
+survey cycle comprised 53.77% of the total analyzed sample whereas the
+2017 - 2018 survey cycle comprised 46.23% of the total analyzed sample.
+Among the participants, the proportion of males and females were 53.76%
+(n = 1063) and 46.23% (n = 914), respectively. With race, the proportion
+of Mexican American, White and Black were 9.16%, 34.90% and 17.10%,
+respectively. Overall, the mean (SD) values for age, albumin, SBP, DBP,
+hemoglobin, TC, AST, HDL, BMI were 43.79 (17.43) years, 43.04 (3.59)
+g/L, 118.61 (16.41) mmHg, 69.60 (11.18) mmHg, 13.94 (1.46)g/dL, 4.81
+(1.02) mmol/L, 23.47 (13.17) IU/L, 1.60 (0.43) mmol/L, and 22.08 (2.03)
+kg/m2, respectively. Among the participants, 73.70% were alcohol
+drinkers, 5.06% were diabetic, 7.79% were hypertensive, 13.76%
+experienced snorting whereas 21.60% were smokers. Sleep duration was
+divided into three groups, which were \< 6 h, 6--8 h, ≥ 8 h, each with a
+proportion of 7.89%, 40.06%, 52.05%, respectively. Male and female
+participants differed significantly on all characteristics except
+citizenship status, and the cycle of the survey.
 
@@
 
-Table \@ref(tab:tabthree) presents the summary of univariate regression analysis results exploring the association between the exposures, selected potential confounding variables and systolic blood pressure.
-We found substantial associations between ...
+Table \@ref(tab:tabthree) presents the summary of univariate regression
+analysis results exploring the association between the exposures,
+selected potential confounding variables and systolic blood pressure. We
+found substantial associations between ...
 

### 2023-12-11 data/scrip2.R

@@ -11,11 +11,11 @@ library(ggcorrplot)
 library(kableExtra)
 
 ## Load Data
-## dt <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
-dt <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/dt.csv")
+dt <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
+##dt <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/dt.csv")
 
 ## load the variable name file
-## vars <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/vars.csv")
+vars <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/vars.csv")
 ## vars <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/vars.csv")
 
 dt$cycle <- as.factor(dt$cycle)
@@ -266,7 +266,7 @@ print(significant_sbp)
 ##-------------------------------------
 ## Systolic Blood Pressure
 ## Load variables selected by LASSO
-## lasso.coef.min <- read.csv("../SleepBP/data/lasso_res_sbp_min.csv")
+lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_sbp_min.csv")
 ## lasso.coef.min$x
 names.lasso.min <- c("bmi", "hdl", "total_chol","AST",
                      "creatinine", "diabetes", "citizenship_status",
@@ -289,7 +289,7 @@ modify_column_unhide(column = std.error)
 ## Diastolic Blood Pressure
 ##---------------------------------------------
 ## Load variables selected by LASSO
-## lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_dbp_min.csv")
+lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_dbp_min.csv")
 ## lasso.coef.min$x
 names.lasso.min <- c("bmi","total_chol", "hemoglobin", "albumin", "AST",
[…]

### 2023-12-11 manuscript/paper.Rmd
update

@@
 
-<!-- Beyond age, the basal metabolic rate (BMI) emerges as a recognized -->
-
-<!-- contributor to systolic blood pressure (SBP), with individuals -->
-
-<!-- exhibiting higher BMI values demonstrating a propensity for elevated -->
-
-<!-- SBP. Gender also holds a discernible association with SBP dynamics. -->
-
-<!-- Empirical evidence in CITE demonstrates that women typically manifest -->
-
-<!-- lower SBP than men until the onset of menopause, after which women -->
-
-<!-- exhibit higher SBP levels compared to men. Additionally, heightened -->
-
-<!-- levels of cholesterol are closely linked to vascular damage. The -->
-
-<!-- accumulation of cholesterol deposits and calcium within arteries -->
-
-<!-- precipitates arterial stiffening, necessitating augmented cardiac -->
-
-<!-- efforts to circulate blood, potentially culminating in elevated blood -->
-
-<!-- pressure. This study endeavors to examine the interrelationships between -->
-
-<!-- age and cholesterol levels, emphasizing consideration of potential -->
-
-<!-- confounding variables. Table 1 comprehensively delineates the array of -->
-
-<!-- variables under scrutiny, ensuring a comprehensive analysis of their -->
-
-<!-- collective impact -->
-
+```{=html}
+<!-- Beyond age, the basal metabolic rate (BMI) emerges as a recognized
+contributor to systolic blood pressure (SBP), with individuals
+exhibiting higher BMI values demonstrating a propensity for elevated
+SBP. Gender also holds a discernible association with SBP dynamics.
+Empirical evidence in CITE demonstrates that women typically manifest
+lower SBP than men until the onset of menopause, after which women
+exhibit higher SBP levels compared to men. Additionally, heightened
+levels of cholesterol are closely linked to vascular damage. The
+accumulation of cholesterol deposits and calcium within arteries
+precipitates arterial stiffening, necessitating augmented cardiac
+efforts to circulate blood, potentially culminating in elevated blood
+pressure. This study endeavors to examine the interrelationships between
+age and cholesterol levels, emphasizing consideration of potential
+confounding variables. Table 1 comprehensively delineates the array of
+variables under scrutiny, ensuring a comprehensive analysis of their
+collective impact. -->
+```
 # Methods
@@
 
+```{=html}
+<!-- 
 ### Exposures:
@@
 potential effect modification by other factors.
-
+-->
+```
 ### Covariates:
 
-In addition to the exposure and outcome variables, our analyses included
-the following as covariates: Race divided into four groups as Mexican
-American, white, black and other race. Alcohol consumption was grouped
-into drinking, no drinking, not recorded. Smoking status as smoking, not
-smoking, not recorded. Diabetes was defined as yes, no, borderline, or
-not recorded. Hypertension was defined as yes, no, or not recorded.
-Snorting was defined as yes, no, and not recorded. US citizenship status
-was defined as citizen by birth or naturalization, don't know, not a
-citizen, or refused to answer. Education level was grouped into four
-categories including graduate studies, high school, less than grade 12
-or some college. Additional covariates included marital status, gender,
-age, albumin, creatinine,hemoglobin, total cholesterol (TC), aspertate
+```{=html}
+<!-- Sleep duration on workdays was evaluated by the questionnaire with the
+following questions: "Number of hours usually sleep on weekdays or
+workdays". We then categorized this variable into three groups, that is,
+\< 6 h, 6--8 h, ≥ 8 h respectively, and used \< 6 h as the reference
+group in our analysis. -->
+```
+Our analyses included the following as covariates: Race divided into
+four groups as Mexican American, white, black and other race. Alcohol
+consumption was grouped into drinking, no drinking, not recorded.
+Smoking status as smoking, not smoking, not recorded. Diabetes was
+defined as yes, no, borderline, or not recorded. Hypertension was
+defined as yes, no, or not recorded. Snorting was defined as yes, no,
+and not recorded. US citizenship status was defined as citizen by birth
+or naturalization, don't know, not a citizen, or refused to answer.
+Education level was grouped into four categories including graduate
+studies, high school, less than grade 12 or some college. Additional
+covariates included marital status, gender, age, albumin,
+creatinine,hemoglobin, total cholesterol (TC), aspertate
 aminotransferase (AST), high-density lipoprotein (HDL), and body mass
@@
 smoking, race, hypertension, and snorting were analyzed as categorical
-variables. Table \@ref{tab:tabone} highlights the variable names and
-descriptions as utilized in this study.
+variables. Sleep duration on workdays was also recorded and was
+categorized into three groups, that is, \< 6 h, 6--8 h, ≥ 8 h
+respectively, and used \< 6 h as the reference group in our analysis.
+Table \@ref{tab:tabone} highlights the variable names and descriptions
+as utilized in this study.
 
@@
 
-## Model Evaluations
-
-Table \@ref(tab:tabmodel) summarizes the performance metrics across the
-different models fitted to the data in our investigation.
-
-```{=tex}
-\begin{table}
-
-\caption{\label{tab:tabmodel}Summary model performance measures.}
-\centering
-\begin{tabular}[t]{lrrrrrr}
-\toprule
-  & df & AIC & BIC & MSE & $R^2$ & Adjusted $R^2$\\
-\midrule
-Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
-Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
-Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
-Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
-Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
-\addlinespace
-Model 6 & 41 & 15888.38 & 16117.55 & 173.69 & 0.36 & 0.34\\
-Model 7 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
-\bottomrule
-\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
-\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
-\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
-\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
-\end{tabular}
-\end{table}
-```
-```{r tabmodel, eval=FALSE, echo=FALSE}
-## Model performance
-tabmodel <- model.checks %>% 
-  kable(format = "latex", caption = "Summary model performance measures.", digits = 2, booktabs = TRUE, col.names = c("df", "AIC", "BIC", "MSE", "R^2", "Adjusted R2"), escape = FALSE) %>% 
-  add_footnote(label = c("df = degrees of freedom", "AIC = Akaike Information Criteria", "BIC = Bayesian Information Criteria", "MSE = Mean Squared Error"))
-tabmodel
-```
-
 # Results
@@
 
+## Model Evaluations
+
+Table \@ref(tab:tabmodel) summarizes the performance metrics across the
+different models fitted to the data in our investigation.
+
+```{=tex}
+\begin{table}[H]
+
+\caption{\label{tab:tabmodel}Summary model performance measures.}
+\centering
+\begin{tabular}[t]{lrrrrrr}
+\toprule
+  & df & AIC & BIC & MSE & $R^2$ & Adjusted $R^2$\\
+\midrule
+Model 1 & 44 & 15888.48 & 16134.41 & 173.17 & 0.36 & 0.34\\
+Model 2 & 45 & 15849.26 & 16100.78 & 169.60 & 0.37 & 0.36\\
+Model 3 & 39 & 15883.76 & 16101.74 & 173.63 & 0.36 & 0.34\\
+Model 4 & 600 & 16156.48 & 19510.09 & 113.00 & 0.58 & 0.40\\
+Model 5 & 40 & 15848.93 & 16072.50 & 170.43 & 0.37 & 0.35\\
+\addlinespace
+Model 6 & 41 & 15888.38 & 16117.55 & 173.69 & 0.36 & 0.34\\
+Model 7 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
+\bottomrule
+\multicolumn{7}{l}{\textsuperscript{a} df = degrees of freedom}\\
+\multicolumn{7}{l}{\textsuperscript{b} AIC = Akaike Information Criteria}\\
+\multicolumn{7}{l}{\textsuperscript{c} BIC = Bayesian Information Criteria}\\
+\multicolumn{7}{l}{\textsuperscript{d} MSE = Mean Squared Error}\\
+\end{tabular}
+\end{table}
+```
+```{r tabmodel, eval=FALSE, echo=FALSE}
+## Model performance
+tabmodel <- model.checks %>% 
+  kable(format = "latex", caption = "Summary model performance measures.", digits = 2, booktabs = TRUE, col.names = c("df", "AIC", "BIC", "MSE", "R^2", "Adjusted R2"), escape = FALSE) %>% 
+  add_footnote(label = c("df = degrees of freedom", "AIC = Akaike Information Criteria", "BIC = Bayesian Information Criteria", "MSE = Mean Squared Error"))
+tabmodel
+```
+
 ## Subgroup analysis by gender
 
-# Discussion and conclusion
+# Conclussion
+
+## Discussion
+
+## Limitations
+
+Our study encountered a significant challenge with missing data,
+resulting in a notable reduction in our sample size. Despite initially
+having a sample of \$n=19,225\$, the final analysis was conducted on a
+considerably smaller sample of \$n=1,977\$ observations. This
+substantial reduction in the sample size raises concerns about the
+robustness and reliability of our findings. Addressing this issue
+becomes crucial to ensure the validity of our study outcomes.
+
+In response to this challenge, we propose the implementation of multiple
+imputation techniques. This approach involves generating multiple
+plausible sets of missing values based on observed data patterns. By
+imputing missing values using this method, we aim to enhance the
+completeness of our dataset. This strategy is intended to mitigate
+potential biases resulting from missing data and fortify the reliability
+and accuracy of our analytical outcomes.
 

### 2023-12-11 data/scrip2.R
update

@@ -78,7 +78,7 @@ design = svydesign(id = ~psu, weights = ~weights, strata = ~strata,
 
 ##-------------------------------
 ## Univariate regressions (1) sbp
-dt.sbp <- dt %>% select(-c("seq_no", "psu", "strata", "dbp", "weights", "hypertension"))
+dt.sbp <- dt %>% dplyr::select(-c("seq_no", "psu", "strata", "dbp", "weights", "hypertension"))
 sbp.univariate <- tbl_uvregression(data = dt.sbp, y = sbp, method = lm,
                                    hide_n = TRUE,
                                    label = list(bmi ~ "BMI",

### 2023-12-11 manuscript/paper.Rmd
update

@@
 
-```{=html}
-<!-- 
 ### Exposures:
@@
 potential effect modification by other factors.
--->
-```
+
 ### Covariates:
@@
 smoking, race, hypertension, and snorting were analyzed as categorical
-variables. Sleep duration on workdays was also recorded and was
-categorized into three groups, that is, \< 6 h, 6--8 h, ≥ 8 h
-respectively, and used \< 6 h as the reference group in our analysis.
-Table \@ref{tab:tabone} highlights the variable names and descriptions
-as utilized in this study.
+variables. Table \@ref{tab:tabone} highlights the variable names and
+descriptions as utilized in this study.
 
@@
 
+### Variable selection based on univariate regression
+
+Univariate regression analyses were performed on all potentially confounding covariates excluding hypertension.
+Model 1 was fitted to the data including the exposure variables as well as all potentially confounding covariates from univariate regression analysis.
+Model 2 was fitted to the same set of covariates including an interaction term between age and total cholesterol levels.
+Model 3 was fitted to the data including only covariates that had a statistically significant association with SBP in univariate analyses excluding HDL, hemoglobin, diabetes, survey cycle, and alcohol consumption.
+Model 4 included similar covariates as Model 3 but examined, in addition, all possible two-way interactions between the covariates.
+Model 5 included the same set of covariates as Model 3 in addition to an interaction term between age and total cholesterol levels.
+
+### Variable selection based on LASSO and best subsets
+
 With the purpose of identifying influential variables impacting blood
@@
 leading to the inclusion of 29 variables in the model.
+
+## Model performance evaluation
+
+All models fitted to the data in this investigation were nested and evaluations were based on information criteria including Akaike Information Criteria (AIC) and Bayesian Information Criteria (BIC).
+[BIC] Additionally, we also used performance measures including the mean squared error (MSE), r-squared, and adjusted r-squared.
+The likelihood ratio test (LRT) was used in competing model comparisons testing the null hypothesis that the nested models are equivalent, and the extra parameters in the larger model do not improve the fit significantly, that is, these terms have effect sizes equivalent to 0.
+\@ref(tab:tabmodel) summarizes the model performance metrics compared for all models implemented in this investigation.
+
+Here is where we add something....
+
+## compare model 5 with model 2 because they have competing BIC and Adj R^2
 
@@
 
-## Subgroup analysis by gender
+##
 

### 2023-12-11 manuscript/paper.Rmd
udpate

--- 2023-12-11 manuscript/paper.Rmd
@@
 
-With the purpose of identifying influential variables impacting blood
-pressure, a variable selection process was conducted. The initial
-approach involved employing best subset selection with 10-fold
-cross-validation to ascertain the optimal number of subsets. Notably,
-for systolic blood pressure, the forward best subset selection method
-discerned 13 variables, whereas diastolic blood pressure exhibited 19
-selected variables. However, a limitation inherent in forward best
-subset selection is its inability to eliminate previously selected
-features, potentially disregarding their relevance in light of newly
-added variables. Consequently, this methodology may yield sub-optimal
-variable selections due to its lack of adaptability.
+### Variable selection based on univariate regression
+
+### Variable selection based on LASSO and best subsets
+
+Furthermore, we conducted two distinct variable selection methodologies,
+similarly excluding hypertension. Model 6 was based on best subset
+selection with 10-fold cross-validation to ascertain the optimal number
+of subsets. Notably, for systolic blood pressure, the forward best
+subset selection method discerned 13 variables, whereas diastolic blood
+pressure exhibited 19 selected variables. However, a limitation inherent
+in forward best subset selection is its inability to eliminate
+previously selected features, potentially disregarding their relevance
+in light of newly added variables. Consequently, this methodology may
+yield sub-optimal variable selections due to its lack of adaptability.
 
@@
 ensure a negligible increase of no more than 1% in MSE, ultimately
-leading to the inclusion of 29 variables in the model.
+leading to the inclusion of 29 variables in the model. Model 7 consisted
+in variables selected by the LASSO regression. In addition, our study
+incorporated an alternative model that not only incorporated the
+variables identified through LASSO but also introduced an interaction
+term involving age and cholesterol levels; Model 8. This extended model
+aimed to explore potential combined effects between age and cholesterol
+on the outcome variable.
 

### 2023-12-11 data/scrip2.R

@@ -232,7 +232,7 @@ summary(adj.sbp)
 ## All two way-interactions from above
 ##--------------------------------------
 adj.sbp.int <- lm(sbp ~ (bmi + total_chol + gender + age_yr + marital_status + sleep + income_category + albumin + AST + creatinine + citizenship_status + educ_level + children..5 + race + smoke + snort + alcohol)^2 , data = dt.sbp)
-summary(adj.sbp.int)
+summary(adj.sbp.int) 
 
 ## sbp with total_chol and age
 adj.sbp.int2 <- lm(sbp ~ bmi + total_chol + gender + age_yr + marital_status + sleep + income_category + albumin + AST + creatinine + citizenship_status + educ_level + children..5 + race + smoke + snort + alcohol + total_chol:age_yr , data = dt.sbp)
@@ -248,7 +248,7 @@ summary(adj.sbp.int2)
 ## saved only results where p < 0.05
 
 ## sbp
-int.sbp <- lm(sbp ~ .^2, data = dt.sbp)
+int.sbp <- lm(sbp ~ .^2, data = dt.sbp)  ####check
 coefficients <- summary(int.sbp)$coefficients
 coef_df <- as.data.frame(coefficients)
 significant_sbp <- coef_df[coef_df$`Pr(>|t|)` < 0.05, ]
@@ -285,7 +285,13 @@ sum((predlasso-dt$sbp)^2)/nrow(dt)
 
 tbl_regression(modlasso.sbp.min, include = everything()) %>%
 modify_column_unhide(column = std.error)
- 
+
+
+##WITH interactin 
+modlasso.sbp.min.INT <- lm(sbp ~. + total_chol:age_yr ,data = dt %>% select(all_of(names.lasso.min),sbp)
+)
+summary(modlasso.sbp.min)
+predlasso.INT <- predict(modlasso.sbp.min.INT, newdata = dt %>% select(all_of(names.lasso.min)))
 ## Diastolic Blood Pressure
 ##---------------------------------------------
 ## Load variables selected by LASSO
@@ -329,7 +335,9 @@ modref.sbp <- lm(sbp ~. ,data = dt %>% select(all_of(names.reg),sbp))
 summary(modref.sbp)
 pred <- predict(modref.sbp, newdata = dt %>% select(all_of(names.reg)))
 sum((pred-dt$sbp)^2)/nrow(dt)
[…]

### 2023-12-11 manuscript/paper.Rmd

--- 2023-12-11 manuscript/paper.Rmd
@@
 
-With the purpose of identifying influential variables impacting blood
-pressure, a variable selection process was conducted. The initial
-approach involved employing best subset selection with 10-fold
-cross-validation to ascertain the optimal number of subsets. Notably,
-for systolic blood pressure, the forward best subset selection method
-discerned 13 variables, whereas diastolic blood pressure exhibited 19
-selected variables. However, a limitation inherent in forward best
-subset selection is its inability to eliminate previously selected
-features, potentially disregarding their relevance in light of newly
-added variables. Consequently, this methodology may yield sub-optimal
-variable selections due to its lack of adaptability.
+Furthermore, we conducted two distinct variable selection methodologies,
+similarly excluding hypertension. Model 6 was based on best subset
+selection with 10-fold cross-validation to ascertain the optimal number
+of subsets. Notably, for systolic blood pressure, the forward best
+subset selection method discerned 13 variables, whereas diastolic blood
+pressure exhibited 19 selected variables. However, a limitation inherent
+in forward best subset selection is its inability to eliminate
+previously selected features, potentially disregarding their relevance
+in light of newly added variables. Consequently, this methodology may
+yield sub-optimal variable selections due to its lack of adaptability.
 

### 2023-12-11 manuscript/paper.Rmd
update

@@
 All models fitted to the data in this investigation were nested and evaluations were based on information criteria including Akaike Information Criteria (AIC) and Bayesian Information Criteria (BIC).
-[BIC] Additionally, we also used performance measures including the mean squared error (MSE), r-squared, and adjusted r-squared.
+Additionally, we also used performance measures including the mean squared error (MSE), r-squared, and adjusted r-squared.
 The likelihood ratio test (LRT) was used in competing model comparisons testing the null hypothesis that the nested models are equivalent, and the extra parameters in the larger model do not improve the fit significantly, that is, these terms have effect sizes equivalent to 0.
 \@ref(tab:tabmodel) summarizes the model performance metrics compared for all models implemented in this investigation.
-
-Here is where we add something....
-
-## compare model 5 with model 2 because they have competing BIC and Adj R^2
 

### 2023-12-11 data/scrip2.R
up

@@ -402,10 +402,9 @@ R <- t(data.frame(sbp.full = summary(sbp.full)$r.squared,
                      adj.sbp = summary(adj.sbp)$r.squared,
                      adj.sbp.int = summary(adj.sbp.int)$r.squared,
                      adj.sbp.int2 = summary(adj.sbp.int2)$r.squared,
-                     modref.sbp = summary(modref.sbp)$r.squared),
+                     modref.sbp = summary(modref.sbp)$r.squared,
                      modlasso.sbp.min = summary(modlasso.sbp.min)$r.squared,
-                     modlasso.sbp.min.INT = summary(modlasso.sbp.min.INT)$r.squared
-)
+                     modlasso.sbp.min.INT = summary(modlasso.sbp.min.INT)$r.squared))
 colnames(R) <- "R^2"
 
 ## Adjusted R^2
@@ -414,10 +413,9 @@ AdjR <- t(data.frame(sbp.full = summary(sbp.full)$adj.r.squared,
                    adj.sbp = summary(adj.sbp)$adj.r.squared,
                    adj.sbp.int = summary(adj.sbp.int)$adj.r.squared,
                    adj.sbp.int2 = summary(adj.sbp.int2)$adj.r.squared,
-                   modref.sbp = summary(modref.sbp)$adj.r.squared),
+                   modref.sbp = summary(modref.sbp)$adj.r.squared,
                    modlasso.sbp.min = summary(modlasso.sbp.min)$adj.r.squared,
-                   modlasso.sbp.min.INT = summary(modlasso.sbp.min.INT)$adj.r.squared,
-)
+                   modlasso.sbp.min.INT = summary(modlasso.sbp.min.INT)$adj.r.squared))
 colnames(AdjR) <- "Adjusted R^2"
 
 

### 2023-12-11 manuscript/paper.Rmd
hi
up
update

@@
 
-Univariate regression analyses were performed on all potentially confounding covariates excluding hypertension.
-Model 1 was fitted to the data including the exposure variables as well as all potentially confounding covariates from univariate regression analysis.
-Model 2 was fitted to the same set of covariates including an interaction term between age and total cholesterol levels.
-Model 3 was fitted to the data including only covariates that had a statistically significant association with SBP in univariate analyses excluding HDL, hemoglobin, diabetes, survey cycle, and alcohol consumption.
-Model 4 included similar covariates as Model 3 but examined, in addition, all possible two-way interactions between the covariates.
-Model 5 included the same set of covariates as Model 3 in addition to an interaction term between age and total cholesterol levels.
+Univariate regression analyses were performed on all potentially
+confounding covariates excluding hypertension. Model 1 was fitted to the
+data including the exposure variables as well as all potentially
+confounding covariates from univariate regression analysis. Model 2 was
+fitted to the same set of covariates including an interaction term
+between age and total cholesterol levels. Model 3 was fitted to the data
+including only covariates that had a statistically significant
+association with SBP in univariate analyses excluding HDL, hemoglobin,
+diabetes, survey cycle, and alcohol consumption. Model 4 included
+similar covariates as Model 3 but examined, in addition, all possible
+two-way interactions between the covariates. This model is essentially overfit to the data. Model 5 included the same
+set of covariates as Model 3 in addition to an interaction term between
+age and total cholesterol levels.
 
@@
 
-All models fitted to the data in this investigation were nested and evaluations were based on information criteria including Akaike Information Criteria (AIC) and Bayesian Information Criteria (BIC).
-Additionally, we also used performance measures including the mean squared error (MSE), r-squared, and adjusted r-squared.
-The likelihood ratio test (LRT) was used in competing model comparisons testing the null hypothesis that the nested models are equivalent, and the extra parameters in the larger model do not improve the fit significantly, that is, these terms have effect sizes equivalent to 0.
-\@ref(tab:tabmodel) summarizes the model performance metrics compared for all models implemented in this investigation.
+All models fitted to the data in this investigation were nested and
+evaluations were based on information criteria including Akaike
+Information Criteria (AIC) and Bayesian Information Criteria (BIC). The
+BIC is favored over the AIC due to its increasing penalty with larger
+sample sizes, resulting in more stringent significance levels [@BIC].
+Additionally, we also used performance measures including the mean
+squared error (MSE), r-squared, and adjusted r-squared. The likelihood
+ratio test (LRT) was used in competing model comparisons testing the
+null hypothesis that the nested models are equivalent, and the extra
+parameters in the larger model do not improve the fit significantly,
+that is, these terms have effect sizes equivalent to 0.
+\@ref(tab:tabmodel) summarizes the model performance metrics compared
+for all models implemented in this investigation. The final model superiority was based on likelihood ratio tests.
 
@@
 
-## Model Evaluations
+## Model performance evaluation
 
@@
 ```
+
+Likelihood ratio tests showed a statistically significant difference between Model 1 and 2 with evidence of in favor of the latter ($\chi^2_{(1)} = 7064; p>0.05$).
+Model 2 which adjusted for all potentially confounding covariates in \@ref(tab:tabthree) including an interaction between cholestrol level and age was then compared to Model 3 which adjusted for only covariates we found significant in univariate analyses.
+Results here similarly showed evidence in favor of Model 2 ($\chi^2_{(1)} = 7979; p>0.05$).
+Comparing Model 2 to Model 3 revealed a non-statistically difference in model fit suggesting that Model 3, even though smaller performed comparably well.
+Model 3 was thus selected and compared to Model 4, 5, 6, and 7 with evidence based on the LRT favoring its choice as the best model ($p>0.05$).
+Additional regression diagnostics for Model 3 including residual versus fitted value plots, Q-Q plots for checking the normality of residuals, Cook's distance for outliers, residual and high leverage point plots did not reveal substantial deviations from linear regression assumptions.
+Model 3 was thus selected as the best model.
+Reported results are thus based on this model.
+
 ```{r tabmodel, eval=FALSE, echo=FALSE}
@@
 
-##
+## Discussion and conclusion
 
-# Conclussion
-
-## Discussion
-
-## Limitations
+### Study limitations
 
@@
 
-# References
+# Refences

### 2023-12-11 data/scrip2.R

@@ -449,11 +449,12 @@ anova.four
 anova.five <- anova(adj.sbp.int2, modref.sbp, test = "LRT")
 anova.five
 
-
+anova.six <- anova(adj.sbp.int2, modlasso.sbp.min.INT, test = "LRT")
+anova.six
 ##-----------------------------------
 ## Regression diagnostics
 par(mfrow = c(2,2))
 plot(adj.sbp.int2)
 
 
-# Having high cholesterol, especially high LDL (bad) cholesterol or low HDL (good) cholesterol, can increase the risk of atherosclerosis, which is the buildup of plaque in the arteries. Atherosclerosis can lead to heart attack, stroke, and other complications.
+# Having high cholesterol, especially high LDL (bad) cholesterol or low HDL (good) cholesterol, can increase the risk of atherosclerosis, which is the build up of plaque in the arteries. Atherosclerosis can lead to heart attack, stroke, and other complications.

### 2023-12-11 manuscript/paper.Rmd
hi

--- 2023-12-11 manuscript/paper.Rmd
@@
 
-Univariate regression analyses were performed on all potentially confounding covariates excluding hypertension.
-Model 1 was fitted to the data including the exposure variables as well as all potentially confounding covariates from univariate regression analysis.
-Model 2 was fitted to the same set of covariates including an interaction term between age and total cholesterol levels.
-Model 3 was fitted to the data including only covariates that had a statistically significant association with SBP in univariate analyses excluding HDL, hemoglobin, diabetes, survey cycle, and alcohol consumption.
-Model 4 included similar covariates as Model 3 but examined, in addition, all possible two-way interactions between the covariates.
-Model 5 included the same set of covariates as Model 3 in addition to an interaction term between age and total cholesterol levels.
+Univariate regression analyses were performed on all potentially
+confounding covariates excluding hypertension. Model 1 was fitted to the
+data including the exposure variables as well as all potentially
+confounding covariates from univariate regression analysis. Model 2 was
+fitted to the same set of covariates including an interaction term
+between age and total cholesterol levels. Model 3 was fitted to the data
+including only covariates that had a statistically significant
+association with SBP in univariate analyses excluding HDL, hemoglobin,
+diabetes, survey cycle, and alcohol consumption. Model 4 included
+similar covariates as Model 3 but examined, in addition, all possible
+two-way interactions between the covariates. Model 5 included the same
+set of covariates as Model 3 in addition to an interaction term between
+age and total cholesterol levels.
 
@@
 
-All models fitted to the data in this investigation were nested and evaluations were based on information criteria including Akaike Information Criteria (AIC) and Bayesian Information Criteria (BIC).
-Additionally, we also used performance measures including the mean squared error (MSE), r-squared, and adjusted r-squared.
-The likelihood ratio test (LRT) was used in competing model comparisons testing the null hypothesis that the nested models are equivalent, and the extra parameters in the larger model do not improve the fit significantly, that is, these terms have effect sizes equivalent to 0.
-\@ref(tab:tabmodel) summarizes the model performance metrics compared for all models implemented in this investigation.
+All models fitted to the data in this investigation were nested and
+evaluations were based on information criteria including Akaike
+Information Criteria (AIC) and Bayesian Information Criteria (BIC). The
+BIC is favored over the AIC due to its increasing penalty with larger
+sample sizes, resulting in more stringent significance levels [@BIC].
+Additionally, we also used performance measures including the mean
+squared error (MSE), r-squared, and adjusted r-squared. The likelihood
+ratio test (LRT) was used in competing model comparisons testing the
+null hypothesis that the nested models are equivalent, and the extra
+parameters in the larger model do not improve the fit significantly,
+that is, these terms have effect sizes equivalent to 0.
+\@ref(tab:tabmodel) summarizes the model performance metrics compared
+for all models implemented in this investigation.
 
@@
 
-Table \@ref(tab:tabmodel) summarizes the performance metrics across the
-different models fitted to the data in our investigation.
+Table \@ref(tab:tabmodel) presents an overview of various models applied
+to our dataset. Model 6, utilizing best subset selection, reveals
+notably high MSE, suggesting limited predictive efficacy. Models 1 and
+2, not including variable selection techniques, highlight the impact of
+incorporating the age-cholesterol interaction term, showing a reduction
+in MSE. Models 3, 4, and 5 included solely significant univariate
+regression variables. Model 5, with the lowest AIC and second lowest
+BIC, demonstrates an acceptable MSE. Model 4, with the lowest MSE and
+highest \$R\^2\$, implies predictive power, although its extensive terms
+raise concerns of potential overfitting. Model 3, based solely on
+significant univariate covariates, performs well but exhibits a higher
+MSE than Models 4 and 5, suggesting the probable influence of
+interaction terms. Models 7 and 8, employing LASSO, exhibit comparable
+performance metrics, with Model 8 displaying reduced BIC and MSE,
+attributed chiefly to the significant inclusion of an Age-Cholesterol
+interaction term.
 
@@
 \toprule
-  & df & AIC & BIC & MSE & $R^2$ & Adjusted $R^2$\\
+  & $\text{df}^a$ & $\text{AIC}^b$ & $\text{BIC}^c$ & $\text{MSE}^d$ & $R^2$ & Adjusted $R^2$\\
 \midrule
@@
 \addlinespace
-Model 6 & 41 & 15888.38 & 16117.55 & 173.69 & 0.36 & 0.34\\
-Model 7 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
+Model 6 & 25 & 15877.41 & 16017.15 & 2657.01 & 0.35 & 0.34\\
+Model 7 & 41 & 15888.38 & 16117.55 & 173.69  & 0.36 & 0.34\\
+Model 8 & 42 & 15849.19 & 16083.94 & 170.11  & 0.37 & 0.36\\
 \bottomrule
@@
 
-# Conclussion
+# Conclusion
 

### 2023-12-11 manuscript/paper.Rmd

--- 2023-12-11 manuscript/paper.Rmd
@@
 
-Table \@ref(tab:tabmodel) summarizes the performance metrics across the
-different models fitted to the data in our investigation.
+Table \@ref(tab:tabmodel) presents an overview of the performance measures of the various models applied
+to our data set. Model 6, utilizing best subset selection, revealed
+notably high MSE, suggesting limited predictive efficacy. Models 1 and
+2, not including variable selection techniques, highlight the impact of
+incorporating the age-cholesterol interaction term with notable reduction
+in MSE. Models 3, 4, and 5 included solely significant univariate
+regression variables. Model 5, with the lowest AIC and second lowest
+BIC, demonstrates an acceptable MSE. Model 4, with the lowest MSE and
+highest \$R\^2\$, demonstrated a high predictive power, although its extensive number of terms
+raise concerns of potential overfitting. Model 3, based solely on
+significant univariate covariates, performed well but exhibited a higher
+MSE than Models 4 and 5, suggesting the probable influence of
+interaction terms. Models 7 and 8, employing LASSO, exhibited comparable
+performance metrics, with Model 8 displaying reduced BIC and MSE,
+attributed chiefly to the significant inclusion of an age-cholesterol
+interaction term.
 
@@
 imputing missing values using this method, we aim to enhance the
-completeness of our dataset. This strategy is intended to mitigate
+completeness of our data set. This strategy is intended to mitigate
 potential biases resulting from missing data and fortify the reliability
@@
 
-# Refences
+# References

### 2023-12-11 manuscript/paper.Rmd
update

@@
 
-Table \@ref(tab:tabthree) presents the summary of univariate regression
+Univariate regression
 analysis results exploring the association between the exposures,
-selected potential confounding variables and systolic blood pressure. We
-found substantial associations between ...
+selected potential confounding variables showed substantial associations between BMI ($\beta = 0.95$ mmHg; 95%CI [0.60 - 1.3]; p<0.05), TC ($\beta = 3.6$ mmHg; 95%CI [2.9 - 4.3]; p<0.05), albumin ($\beta = -0.48$ mmHg; 95% CI [-0.68mmHg - -0.27]; p < 0.05), AST ($\beta = 0.12$ mmHg; 95%CI [0.07 - 0.18]; p <0.05), and age (years) ($\beta = 0.49$ mmHg; 95% CI [0.45 - 0.52]; p <0.05) and systolic blood pressure.
+Weekday sleep duration between 6 and 8 hours was associated with a 4.0 mmHg (95% CI [-6.8 - -1.2]) decrease in the average systolic blood pressure compared to 6 hours of sleep.
+On the other hand, more than 8 hours of sleep was associated with a decrease of 3.1 mmHg in systolic blood compared to 6 hours of sleep.
+Substantial associations were noted between citizenship, education, number of children below five years, gender, marital status, race, smoking status, snorting, alcohol, and income.
 
-```{r tabthree, echo=FALSE}
+```{r tabthree, echo=FALSE, eval=FALSE}
 ## univariate analysis

### 2023-12-11 manuscript/paper.Rmd

@@
 similar covariates as Model 3 but examined, in addition, all possible
-two-way interactions between the covariates. This model is essentially overfit to the data. Model 5 included the same
-set of covariates as Model 3 in addition to an interaction term between
-age and total cholesterol levels.
+two-way interactions between the covariates. This model is essentially
+overfit to the data. Model 5 included the same set of covariates as
+Model 3 in addition to an interaction term between age and total
+cholesterol levels.
 
@@
 \@ref(tab:tabmodel) summarizes the model performance metrics compared
-for all models implemented in this investigation. The final model superiority was based on likelihood ratio tests.
+for all models implemented in this investigation. The final model
+superiority was based on likelihood ratio tests.
 
@@
 
-## Multivariable Regression Analysis
-
-```{r tabfour, echo=FALSE}
-## Final selected multiple linear regression model
-tbl_regression(adj.sbp.int2,
-               label = list(bmi ~ "BMI",
-                            #hdl ~ "HDL",
-                            total_chol ~ "TC",
-                            #hemoglobin ~ "Hemoglobin",
-                            albumin ~ "Albumin",
-                            creatinine ~ "Creatinine",
-                            #hypertension ~ "Hypertension",
-                            #diabetes ~ "Diabetes",
-                            citizenship_status ~ "Citizenship",
-                            educ_level ~ "Education",
-                            children..5 ~ "Children > 5 yrs",
-                            age_yr ~ "Age (yrs)",
-                            marital_status ~ "Marital status",
-                            #cycle ~ "Survey cycle",
-                            sleep ~ "Sleep",
-                            race ~ "Race",
-                            smoke ~ "Smoking status",
-                            snort ~ "Snort",
-                            alcohol ~ "Alcohol",
-                            income_category ~ "Income",
-                            gender ~ "Gender"
-               )) %>% modify_header(label ~ "**Variable**") %>%
-  bold_labels() %>% 
-  as_kable_extra(format = "latex", digits = 2, caption = "Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP).
-BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)
-
-```
-
 ## Model performance evaluation
 
-Table \@ref(tab:tabmodel) presents an overview of the performance measures of the various models applied
-to our data set. Model 6, utilizing best subset selection, revealed
-notably high MSE, suggesting limited predictive efficacy. Models 1 and
-2, not including variable selection techniques, highlight the impact of
-incorporating the age-cholesterol interaction term with notable reduction
-in MSE. Models 3, 4, and 5 included solely significant univariate
-regression variables. Model 5, with the lowest AIC and second lowest
-BIC, demonstrates an acceptable MSE. Model 4, with the lowest MSE and
-highest \$R\^2\$, demonstrated a high predictive power, although its extensive number of terms
-raise concerns of potential overfitting. Model 3, based solely on
+Table \@ref(tab:tabmodel) presents an overview of the performance
+measures of the various models applied to our data set. Model 6,
+utilizing best subset selection, revealed notably high MSE, suggesting
+limited predictive efficacy. Models 1 and 2, not including variable
+selection techniques, highlight the impact of incorporating the
+age-cholesterol interaction term with notable reduction in MSE. Models
+3, 4, and 5 included solely significant univariate regression variables.
+Model 5, with the lowest AIC and second lowest BIC, demonstrates an
+acceptable MSE. Model 4, with the lowest MSE and highest \$R\^2\$,
+demonstrated a high predictive power, although its extensive number of
+terms raise concerns of potential overfitting. Model 3, based solely on
 significant univariate covariates, performed well but exhibited a higher
@@
 ```
+Likelihood ratio tests showed a statistically significant difference
+between Model 1 and 2 with evidence of in favor of the latter
+($\chi^2_{(1)} = 7064; p>0.05$). Model 2 which adjusted for all
+potentially confounding covariates in \@ref(tab:tabthree) including an
+interaction between cholestrol level and age was then compared to Model
+3 which adjusted for only covariates we found significant in univariate
+analyses. Results here similarly showed evidence in favor of Model 2
+($\chi^2_{(1)} = 7979; p>0.05$). Comparing Model 2 to Model 3 revealed a
+non-statistically difference in model fit suggesting that Model 3, even
+though smaller performed comparably well. Model 3 was thus selected and
+compared to Model 4, 5, 6, and 7 with evidence based on the LRT favoring
+its choice as the best model ($p>0.05$). Additional regression
+diagnostics for Model 3 including residual versus fitted value plots,
+Q-Q plots for checking the normality of residuals, Cook's distance for
+outliers, residual and high leverage point plots did not reveal
+substantial deviations from linear regression assumptions. Model 3 was
+thus selected as the best model. Reported results are thus based on this
+model.
 
-Likelihood ratio tests showed a statistically significant difference between Model 1 and 2 with evidence of in favor of the latter ($\chi^2_{(1)} = 7064; p>0.05$).
-Model 2 which adjusted for all potentially confounding covariates in \@ref(tab:tabthree) including an interaction between cholestrol level and age was then compared to Model 3 which adjusted for only covariates we found significant in univariate analyses.
-Results here similarly showed evidence in favor of Model 2 ($\chi^2_{(1)} = 7979; p>0.05$).
-Comparing Model 2 to Model 3 revealed a non-statistically difference in model fit suggesting that Model 3, even though smaller performed comparably well.
-Model 3 was thus selected and compared to Model 4, 5, 6, and 7 with evidence based on the LRT favoring its choice as the best model ($p>0.05$).
-Additional regression diagnostics for Model 3 including residual versus fitted value plots, Q-Q plots for checking the normality of residuals, Cook's distance for outliers, residual and high leverage point plots did not reveal substantial deviations from linear regression assumptions.
-Model 3 was thus selected as the best model.
-Reported results are thus based on this model.
+## Multivariable Regression Analysis
+
+```{r tabfour, echo=FALSE}
+## Final selected multiple linear regression model
+tbl_regression(adj.sbp.int2,
+               label = list(bmi ~ "BMI",
+                            #hdl ~ "HDL",
+                            total_chol ~ "TC",
+                            #hemoglobin ~ "Hemoglobin",
+                            albumin ~ "Albumin",
+                            creatinine ~ "Creatinine",
+                            #hypertension ~ "Hypertension",
+                            #diabetes ~ "Diabetes",
+                            citizenship_status ~ "Citizenship",
+                            educ_level ~ "Education",
+                            children..5 ~ "Children > 5 yrs",
+                            age_yr ~ "Age (yrs)",
+                            marital_status ~ "Marital status",
+                            #cycle ~ "Survey cycle",
+                            sleep ~ "Sleep",
+                            race ~ "Race",
+                            smoke ~ "Smoking status",
+                            snort ~ "Snort",
+                            alcohol ~ "Alcohol",
+                            income_category ~ "Income",
+                            gender ~ "Gender"
+               )) %>% modify_header(label ~ "**Variable**") %>%
+  bold_labels() %>% 
+  as_kable_extra(format = "latex", digits = 2, caption = "MODEL 5: Adjusted effect of sleep duration and total cholestrol on systolic blood pressure (SBP).
+BMI = Body Mass Index, TC = Total Cholestrol, AST = Aspertate aminotransferase.", booktabs = TRUE, longtable = TRUE)
+
+```
 

### 2023-12-11 manuscript/paper.Rmd
update

@@
 
-Univariate regression
+Table \@ref(tab:tabthree) presents the summary of univariate regression
 analysis results exploring the association between the exposures,
-selected potential confounding variables showed substantial associations between BMI ($\beta = 0.95$ mmHg; 95%CI [0.60 - 1.3]; p<0.05), TC ($\beta = 3.6$ mmHg; 95%CI [2.9 - 4.3]; p<0.05), albumin ($\beta = -0.48$ mmHg; 95% CI [-0.68mmHg - -0.27]; p < 0.05), AST ($\beta = 0.12$ mmHg; 95%CI [0.07 - 0.18]; p <0.05), and age (years) ($\beta = 0.49$ mmHg; 95% CI [0.45 - 0.52]; p <0.05) and systolic blood pressure.
-Weekday sleep duration between 6 and 8 hours was associated with a 4.0 mmHg (95% CI [-6.8 - -1.2]) decrease in the average systolic blood pressure compared to 6 hours of sleep.
-On the other hand, more than 8 hours of sleep was associated with a decrease of 3.1 mmHg in systolic blood compared to 6 hours of sleep.
-Substantial associations were noted between citizenship, education, number of children below five years, gender, marital status, race, smoking status, snorting, alcohol, and income.
+selected potential confounding variables and systolic blood pressure. We
+found substantial associations between ...
 
-```{r tabthree, echo=FALSE, eval=FALSE}
+```{r tabthree, echo=FALSE}
 ## univariate analysis

### 2023-12-11 manuscript/paper.Rmd
up

--- 2023-12-11 manuscript/paper.Rmd
@@
 
-```{r tabfour, echo=FALSE}
+Table \@ref(tab:multreg) shows the coefficients of sleep, age, total
+cholesterol and the interaction between total cholesterol and age
+
+```{=tex}
+
+\begin{table}[H]
+\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients for Model 5}
+\centering
+\begin{tabular}[t]{lrrr}
+\toprule
+  \textbf{Variable} & \textbf{Estimate} & \textbf{95\% CI}  & \textbf{p-value} \\
+\midrule
+\textbf{Intercept} & 102.55 &(88.42,116.69) & $<0.001$\\
+\textbf{Sleep} & 0.75 & (0.50, 1.00) & 0.002 \\
+$<6$hrs &  &  &  \\
+$>8$hrs & -1.4 & (-3.7,0.92) & 0.2\\
+$6-8$hrs & -3.0 & (-5.3,-0.67) &  0.012\\
+\textbf{Age} & -0.04& (-0.22, 0.13)	& 0.6 \\
+\textbf{TC} & -3.5	& (-5.2, -1.8) & <0.001 \\
+\textbf{TC*Age} & 0.11	& (0.07, 0.14) & <0.001 \\
+\bottomrule
+\end{tabular}
+\end{table}
+```
+
+```{r tabfour, echo=FALSE, eval=FALSE}
 ## Final selected multiple linear regression model

### 2023-12-11 data/scrip2.R
hi

@@ -456,3 +456,11 @@ plot(adj.sbp.int2)
 
 
 # Having high cholesterol, especially high LDL (bad) cholesterol or low HDL (good) cholesterol, can increase the risk of atherosclerosis, which is the build up of plaque in the arteries. Atherosclerosis can lead to heart attack, stroke, and other complications.
+
+
+
+
+
+
+
+

### 2023-12-11 manuscript/paper.Rmd
hi

@@
 
-Univariate regression
-analysis results exploring the association between the exposures,
-selected potential confounding variables showed substantial associations between BMI ($\beta = 0.95$ mmHg; 95%CI [0.60 - 1.3]; p<0.05), TC ($\beta = 3.6$ mmHg; 95%CI [2.9 - 4.3]; p<0.05), albumin ($\beta = -0.48$ mmHg; 95% CI [-0.68mmHg - -0.27]; p < 0.05), AST ($\beta = 0.12$ mmHg; 95%CI [0.07 - 0.18]; p <0.05), and age (years) ($\beta = 0.49$ mmHg; 95% CI [0.45 - 0.52]; p <0.05) and systolic blood pressure.
-Weekday sleep duration between 6 and 8 hours was associated with a 4.0 mmHg (95% CI [-6.8 - -1.2]) decrease in the average systolic blood pressure compared to 6 hours of sleep.
-On the other hand, more than 8 hours of sleep was associated with a decrease of 3.1 mmHg in systolic blood compared to 6 hours of sleep.
-Substantial associations were noted between citizenship, education, number of children below five years, gender, marital status, race, smoking status, snorting, alcohol, and income.
+Univariate regression analysis results exploring the association between
+the exposures, selected potential confounding variables showed
+substantial associations between BMI ($\beta = 0.95$ mmHg; 95%CI [0.60 -
+1.3]; p\<0.05), TC ($\beta = 3.6$ mmHg; 95%CI [2.9 - 4.3]; p\<0.05),
+albumin ($\beta = -0.48$ mmHg; 95% CI [-0.68mmHg - -0.27]; p \< 0.05),
+AST ($\beta = 0.12$ mmHg; 95%CI [0.07 - 0.18]; p \<0.05), and age
+(years) ($\beta = 0.49$ mmHg; 95% CI [0.45 - 0.52]; p \<0.05) and
+systolic blood pressure. Weekday sleep duration between 6 and 8 hours
+was associated with a 4.0 mmHg (95% CI [-6.8 - -1.2]) decrease in the
+average systolic blood pressure compared to 6 hours of sleep. On the
+other hand, more than 8 hours of sleep was associated with a decrease of
+3.1 mmHg in systolic blood compared to 6 hours of sleep. Substantial
+associations were noted between citizenship, education, number of
+children below five years, gender, marital status, race, smoking status,
+snorting, alcohol, and income.
 
@@
 
-Table \@ref(tab:multreg) shows the coefficients of sleep, age, total
-cholesterol and the interaction between total cholesterol and age
+Table 1 displays the estimated coefficients derived from a multiple
+regression analysis involving variables. We only show sleep during
+weekdays, age, total cholesterol, and the interaction effect between
+total cholesterol and age. Notably, the findings reveal that an increase
+of one year in age is associated with a marginal decrease in systolic
+blood pressure (SBP) by an amount of 0.04mmHg. However, this observed
+reduction is deemed statistically insignificant within the confines of
+this model. Conversely, an increase of one unit in total cholesterol is
+linked to a more pronounced reduction in SBP by a factor of 3.5mmHg,
+which attains statistical significance. Furthermore, the interaction
+between age and total cholesterol yields a statistically significant
+coefficient estimate of 0.11. This interaction effect suggests that as
+both age and total cholesterol concurrently increase by one unit, there
+is an associated elevation in SBP by a factor of 0.11mmHg. Additionally,
+the coefficients attributed to sleep duration bear analogous
+interpretations to those observed in the univariate scenario.
+Specifically, weekday sleep durations ranging between 6 and 8 hours
+demonstrate an average decrease of 3.0mmHg in systolic blood pressure in
+comparison to the reference duration of 6 hours of sleep. Conversely,
+durations exceeding 8 hours exhibit a decrease of 1.4mmHg in systolic
+blood pressure as compared to the 6-hour reference duration.
 
@@
 ```
-
 ```{r tabfour, echo=FALSE, eval=FALSE}

### 2023-12-11 manuscript/paper.Rmd

--- 2023-12-11 manuscript/paper.Rmd
@@
 ---
-title: "Weekday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population"
+title: "Workday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population"
 thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
@@
 Sleep duration on workdays was evaluated by the questionnaire with the
-following questions: "Number of hours usually sleep on weekdays or
+following questions: "Number of hours usually sleep on Workdays or
 workdays". We then categorized this variable into three groups, that is,
@@
 <!-- Sleep duration on workdays was evaluated by the questionnaire with the
-following questions: "Number of hours usually sleep on weekdays or
+following questions: "Number of hours usually sleep on Workdays or
 workdays". We then categorized this variable into three groups, that is,
@@
 ```
-Our analyses included the following as covariates: Race divided into
+Our analyses included the following as covariates based on previous literature [@su2022association]: Race divided into
 four groups as Mexican American, white, black and other race. Alcohol
@@
 \textbf{DMDCITZN} & \textbf{citizenship status} & \{Are you/Is SP\} a citizen of the United States? [Information about citizenship is being collected by the U.S. Public Health Serv\\
-\textbf{SLD012} & \textbf{sleep} & Number of hours usually sleep on weekdays or workdays\\
+\textbf{SLD012} & \textbf{sleep} & Number of hours usually sleep on Workdays or workdays\\
 \textbf{BMXBMI} & \textbf{bmi} & Body mass index\\
@@
 
-Univariate regression
-analysis results exploring the association between the exposures,
-selected potential confounding variables showed substantial associations between BMI ($\beta = 0.95$ mmHg; 95%CI [0.60 - 1.3]; p<0.05), TC ($\beta = 3.6$ mmHg; 95%CI [2.9 - 4.3]; p<0.05), albumin ($\beta = -0.48$ mmHg; 95% CI [-0.68mmHg - -0.27]; p < 0.05), AST ($\beta = 0.12$ mmHg; 95%CI [0.07 - 0.18]; p <0.05), and age (years) ($\beta = 0.49$ mmHg; 95% CI [0.45 - 0.52]; p <0.05) and systolic blood pressure.
-Weekday sleep duration between 6 and 8 hours was associated with a 4.0 mmHg (95% CI [-6.8 - -1.2]) decrease in the average systolic blood pressure compared to 6 hours of sleep.
-On the other hand, more than 8 hours of sleep was associated with a decrease of 3.1 mmHg in systolic blood compared to 6 hours of sleep.
-Substantial associations were noted between citizenship, education, number of children below five years, gender, marital status, race, smoking status, snorting, alcohol, and income.
+Univariate regression analysis results exploring the association between
+the exposures, selected potential confounding variables showed
+substantial associations between BMI ($\beta = 0.95$ mmHg; 95%CI [0.60 -
+1.3]; p\<0.05), TC ($\beta = 3.6$ mmHg; 95%CI [2.9 - 4.3]; p\<0.05),
+albumin ($\beta = -0.48$ mmHg; 95% CI [-0.68mmHg - -0.27]; p \< 0.05),
+AST ($\beta = 0.12$ mmHg; 95%CI [0.07 - 0.18]; p \<0.05), and age
+(years) ($\beta = 0.49$ mmHg; 95% CI [0.45 - 0.52]; p \<0.05) and
+systolic blood pressure. Workday sleep duration between 6 and 8 hours
+was associated with a 4.0 mmHg (95% CI [-6.8 - -1.2]) decrease in the
+average systolic blood pressure compared to 6 hours of sleep. On the
+other hand, more than 8 hours of sleep was associated with a decrease of
+3.1 mmHg in systolic blood compared to 6 hours of sleep. Substantial
+associations were noted between citizenship, education, number of
+children below five years, gender, marital status, race, smoking status,
+snorting, alcohol, and income.
 
@@
 
-Table \@ref(tab:multreg) shows the coefficients of sleep, age, total
-cholesterol and the interaction between total cholesterol and age
+Table \@ref(tab:multreg) displays the estimated coefficients derived from a multiple
+regression analysis involving variables. We only show sleep during
+Workdays, age, total cholesterol, and the interaction effect between
+total cholesterol and age. Notably, the findings reveal that an increase
+of one year in age is associated with a decrease in systolic
+blood pressure (SBP) by an amount of 0.04 mmHg (95%CI [-0.22 - 0.13], p = 0.60). 
+Conversely, an increase of one unit in total cholesterol is
+linked to a significant reduction in SBP by a factor of 3.5 mmHg (95%CI [-5.2, -1.8], p = 0.01). Furthermore, the interaction
+between age and total cholesterol yielded a statistically significant
+coefficient estimate suggesting that as
+both age and total cholesterol concurrently increase by one unit, there
+is an associated elevation in SBP by a factor of 0.11 mmHg (95%CI [0.07 - 0.14], p = 0.001). Additionally,
+the coefficients attributed to sleep duration bear analogous
+interpretations to those observed in the univariate scenario.
+Specifically, workday sleep durations ranging between 6 and 8 hours
+demonstrate a significant average decrease of 3.0 mmHg (95%CI [-5.3 - -0.67],  p = 0.012) in systolic blood pressure in
+comparison to the reference duration of 6 hours of sleep. Conversely,
+durations exceeding 8 hours exhibit a decrease of 1.4mmHg (95%CI [-3.7- 0.92] p = 0.2]) in systolic
+blood pressure as compared to the 6-hour reference duration.
 
@@
 \begin{table}[H]
-\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients for Model 5}
+\caption{\label{tab:multreg} Sleep, Age and Cholesterol Coefficients based on Model 5.
+These results are adjusted for other significant potentially confounding factors based on univariate regression analysis.}
 \centering
@@
 
+The current study investigated the
+intricate interplay between age, sleep, cholestrol levels and systolic blood
+pressure adjusting for other potentially confounding covariates informed by literature.
+We performed comprehensive model comparisons to explore the association between age, total cholestrol levels, sleep durations and systolic blood pressure.
+We find based on the best model, that work day sleep durations ranging between 6 and 8 hours and more than 8 hours are associated with a decrease in systolic blood pressure relative to the 6 hour reference group.
+Additionally, our study reports substantial effect modification of the effect of total cholestrol levels on systolic blood pressure by age whe adjusted for selected confounders within the study population.
+Previous studies exploring sleep and systolic blood pressure have found similar associations [@su2022association].
+
 ### Study limitations

### 2023-12-11 manuscript/paper.Rmd
hi

@@
 ```
-Our analyses included the following as covariates based on previous literature [@su2022association]: Race divided into
-four groups as Mexican American, white, black and other race. Alcohol
-consumption was grouped into drinking, no drinking, not recorded.
-Smoking status as smoking, not smoking, not recorded. Diabetes was
-defined as yes, no, borderline, or not recorded. Hypertension was
-defined as yes, no, or not recorded. Snorting was defined as yes, no,
-and not recorded. US citizenship status was defined as citizen by birth
-or naturalization, don't know, not a citizen, or refused to answer.
-Education level was grouped into four categories including graduate
-studies, high school, less than grade 12 or some college. Additional
-covariates included marital status, gender, age, albumin,
-creatinine,hemoglobin, total cholesterol (TC), aspertate
-aminotransferase (AST), high-density lipoprotein (HDL), and body mass
-index (BMI). Details about these variables can be found at
+Our analyses included the following as covariates based on previous
+literature [@su2022association]: Race divided into four groups as
+Mexican American, white, black and other race. Alcohol consumption was
+grouped into drinking, no drinking, not recorded. Smoking status as
+smoking, not smoking, not recorded. Diabetes was defined as yes, no,
+borderline, or not recorded. Hypertension was defined as yes, no, or not
+recorded. Snorting was defined as yes, no, and not recorded. US
+citizenship status was defined as citizen by birth or naturalization,
+don't know, not a citizen, or refused to answer. Education level was
+grouped into four categories including graduate studies, high school,
+less than grade 12 or some college. Additional covariates included
+marital status, gender, age, albumin, creatinine,hemoglobin, total
+cholesterol (TC), aspertate aminotransferase (AST), high-density
+lipoprotein (HDL), and body mass index (BMI). Details about these
+variables can be found at
 <https://wwwn.cdc.gov/nchs/nhanes/search/default.aspx>. Age, albumin,
@@
 
-Table \@ref(tab:multreg) displays the estimated coefficients derived from a multiple
-regression analysis involving variables. We only show sleep during
-Workdays, age, total cholesterol, and the interaction effect between
-total cholesterol and age. Notably, the findings reveal that an increase
-of one year in age is associated with a decrease in systolic
-blood pressure (SBP) by an amount of 0.04 mmHg (95%CI [-0.22 - 0.13], p = 0.60). 
-Conversely, an increase of one unit in total cholesterol is
-linked to a significant reduction in SBP by a factor of 3.5 mmHg (95%CI [-5.2, -1.8], p = 0.01). Furthermore, the interaction
+Table \@ref(tab:multreg) displays the estimated coefficients derived
+from a multiple regression analysis involving variables. We only show
+sleep during Workdays, age, total cholesterol, and the interaction
+effect between total cholesterol and age. Notably, the findings reveal
+that an increase of one year in age is associated with a decrease in
+systolic blood pressure (SBP) by an amount of 0.04 mmHg (95%CI [-0.22 -
+0.13], p = 0.60). Conversely, an increase of one unit in total
+cholesterol is linked to a significant reduction in SBP by a factor of
+3.5 mmHg (95%CI [-5.2, -1.8], p = 0.01). Furthermore, the interaction
 between age and total cholesterol yielded a statistically significant
-coefficient estimate suggesting that as
-both age and total cholesterol concurrently increase by one unit, there
-is an associated elevation in SBP by a factor of 0.11 mmHg (95%CI [0.07 - 0.14], p = 0.001). Additionally,
-the coefficients attributed to sleep duration bear analogous
-interpretations to those observed in the univariate scenario.
+coefficient estimate suggesting that as both age and total cholesterol
+concurrently increase by one unit, there is an associated elevation in
+SBP by a factor of 0.11 mmHg (95%CI [0.07 - 0.14], p = 0.001).
+Additionally, the coefficients attributed to sleep duration bear
+analogous interpretations to those observed in the univariate scenario.
 Specifically, workday sleep durations ranging between 6 and 8 hours
-demonstrate a significant average decrease of 3.0 mmHg (95%CI [-5.3 - -0.67],  p = 0.012) in systolic blood pressure in
-comparison to the reference duration of 6 hours of sleep. Conversely,
-durations exceeding 8 hours exhibit a decrease of 1.4mmHg (95%CI [-3.7- 0.92] p = 0.2]) in systolic
-blood pressure as compared to the 6-hour reference duration.
+demonstrate a significant average decrease of 3.0 mmHg (95%CI [-5.3 -
+-0.67], p = 0.012) in systolic blood pressure in comparison to the
+reference duration of 6 hours of sleep. Conversely, durations exceeding
+8 hours exhibit a decrease of 1.4mmHg (95%CI [-3.7- 0.92] p = 0.2]) in
+systolic blood pressure as compared to the 6-hour reference duration.
 
@@
 
-The current study investigated the
-intricate interplay between age, sleep, cholestrol levels and systolic blood
-pressure adjusting for other potentially confounding covariates informed by literature.
+<!-- The current study investigated the intricate interplay between age, sleep, cholestrol levels and systolic blood pressure adjusting for other potentially confounding covariates informed by literature.
 We performed comprehensive model comparisons to explore the association between age, total cholestrol levels, sleep durations and systolic blood pressure.
@@
 Previous studies exploring sleep and systolic blood pressure have found similar associations [@su2022association].
+-->
+
+The present investigation explote into the settle interplay among age,
+sleep patterns, cholesterol levels, and systolic blood pressure, while
+accounting for various potentially confounding factors gleaned from
+pertinent literature. Our study conducted an exhaustive series of model
+comparisons to elucidate the intricate relationship between age, total
+cholesterol levels, sleep durations, and systolic blood pressure.
+
+In light of our findings derived from the best model based on the
+performance measures in Table \@ref(tab:tabmodel), we observed that
+weekday sleep durations spanning between 6 and 8 hours, as well as
+durations exceeding 8 hours, exhibit a decrease in systolic blood
+pressure in comparison to the reference group of individuals sleeping
+for 6 hours. Furthermore, our investigation revealed a noteworthy effect
+modification regarding the influence of total cholesterol levels on
+systolic blood pressure when adjusted for specific confounding variables
+within the studied cohort, especially in relation to age.
+
+Our study contributes to the existing body of knowledge by
+substantiating the associations uncovered in previous research
+concerning sleep patterns and systolic blood pressure
+[@su2022association]. This correlation underscores the significance of
+sleep parameters as influential factors in understanding blood pressure
+dynamics.
 

### 2023-12-12 manuscript/paper.Rmd
up

@@
 ---
-title: "Workday sleep, total cholestrol and the modifying effect of age on systolic blood pressure: An analysis of the 2015-2018 non-obese NHANES population"
+title: "Age modifies the effect of total cholesterol level on systolic blood pressure: Analysis of evidence from the 2015 - 2018 NHANES data"
 thanks: "Replication files are available on the author's Github account (http://github.com/okutse/sleepBP). **Current version**: `r format(Sys.time(), '%B %d, %Y')`"
@@
 selection with 10-fold cross-validation to ascertain the optimal number
-of subsets. Notably, for systolic blood pressure, the forward best
-subset selection method discerned 13 variables, whereas diastolic blood
-pressure exhibited 19 selected variables. However, a limitation inherent
-in forward best subset selection is its inability to eliminate
-previously selected features, potentially disregarding their relevance
-in light of newly added variables. Consequently, this methodology may
-yield sub-optimal variable selections due to its lack of adaptability.
+of subsets. Best subset selection was selected by identifying the best
+model that contains a given number of predictors yielding lowest RSS.
+Notably, for systolic blood pressure, the forward best subset selection
+method discerned 13 variables, whereas diastolic blood pressure
+exhibited 19 selected variables. However, a limitation inherent in
+forward best subset selection is its inability to eliminate previously
+selected features, potentially disregarding their relevance in light of
+newly added variables. Consequently, this methodology may yield
+sub-optimal variable selections due to its lack of adaptability.
 

### 2023-12-15 manuscript/paper.Rmd
updte
final update

@@
 author:
-- Amos Okutse, Brown University
-- Monic Colon-Vargas, Brown University
-abstract: "**Introduction:** Lorem ipsum dolor sit amet, consectetur adipiscing elit. Donec sit amet libero justo. Pellentesque eget nibh ex. Aliquam tincidunt egestas lectus id ullamcorper.
+- Amos Okutse$^1$
+- Monica Colon-Vargas$^1$
+- $^1$School of Public Health, Brown University, Providence, RI.
+abstract: | 
+  **Introduction:** Systolic blood pressure (SBP) is a precursor linked with a spectrum of critical health conditions.
+This nexus between heightened SBP and these significant health adversities underscores the gravity of SBP as a predictive marker for adverse health outcomes.
 
-**Methods:** Proin tellus orci, posuere sed cursus at, bibendum ac odio.
-Nam consequat non ante eget aliquam.
-Nulla facilisis tincidunt elit.
-Nunc hendrerit pellentesque quam, eu imperdiet ipsum porttitor ut.
-Interdum et malesuada fames ac ante ipsum primis in faucibus.
-Suspendisse potenti.
-Duis vitae nibh mauris.
-Duis nec sem sit amet ante dictum mattis.
-Suspendisse diam velit, maximus eget commodo at, faucibus et nisi.
-Ut a pellentesque eros, sit amet suscipit eros.
-Nunc tincidunt quis risus suscipit vestibulum.
-Quisque eu fringilla massa."
+  **Objective:** To explore the intricate interplay between SBP dynamics within non-obese aging populations with self-reported workday sleep duration, total cholesterol levels, and the total cholesterol-age dependence adjusting for selected potentially confounding factors.
+
+  **Methods:** Cross-sectional data spanning two survey cycles (2015 – 2018) was obtained from the National Health and Nutrition Examination Survey (NHANES) database.
+We performed multivariable regressions to examine the interplay between age, total cholesterol levels, self-reported workday sleep duration, and SBP.
+Potentially confounding covariates adjusted for in analyses were selected using univariate regressions, best-subsets selection, and the Least Absolute Shrinkage and Selection Operator (LASSO).
+Model performance evaluation was based on information criteria, mean squared error (MSE), adjusted R-squared, and likelihood ratio tests.
+
+  **Results:** Our results are based on a complete-case analysis of data on n = 1977 participants.
+We found a statistically significant association between workday sleep duration and SBP comparing $6 – 8$ hours sleep category to < $6$ hour sleep category [$\beta$ = -3.0; 95\% Confidence Interval (CI): -5.3,-0.67), p = 0.012] and a substantial effect modification of age on the effect of total cholesterol levels on SBP [$\beta$ = 0.11; 95\%CI: 0.07 - 0.14;  p <0.001].
+
+  **Conclusion:** Workday sleep duration between 6-8 hours is associated with an optimal decrease in SBP.
+Age significantly modifies the effect of total cholesterol levels on SBP, adjusting for other potentially confounding factors.
+
+  ***Keywords:*** *Systolic blood pressure*, *Sleep duration*, *Age*, *Total cholestrol level*, *Effect modification.*
 keywords: "pandoc, r markdown, knitr"
-date: "  `r format(Sys.time(), '%d %B, %Y')` "
+#date: "  `r format(Sys.time(), '%d %B, %Y')` "
 geometry: margin=1in
-#mainfont: cochineal
+mainfont: cochineal
 header-includes:
@@
 # source file 
-source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
+#source(file = "C:/Users/monic/OneDrive/Desktop/SleepBP/data/scrip2.R")
 
-#source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
+source("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/scrip2.R")
 ```
@@
 In western societies aged 40 years and above, a demonstrable pattern
-emerges, revealing an approximate elevation of 7 mmHg in SBP per decade
+emerges, revealing an approximate elevation of about 7 mmHg in SBP per decade
 among individuals above 40 years. This discernible and consistent
-increment in SBP with advancing age accentuates its profound impact
-within the broader context of aging-related health dynamics. Similarly,
+increase in SBP with advancing age accentuates its profound impact
+within the broader context of age-related health dynamics. Similarly,
 while diastolic blood pressure (DBP), demonstrates a concurrent rise
@@
 selected features, potentially disregarding their relevance in light of
-newly added variables. Consequently, this methodology may yield
-sub-optimal variable selections due to its lack of adaptability.
+newly added variables. Foward best subset deos not outperforms shringake
+methods when minimizeing prediction errors [@hastie2009elements] .
+Consequently, this methodology may yield sub-optimal variable selections
+due to its lack of adaptability.
 
@@
 SBP = Systolic blood pressure, DBP = Diastolic blood pressure.",
-                 longtable = TRUE)
+                 longtable = TRUE) %>% 
+  kable_styling(font_size = 8)
 ```
@@
 ($\chi^2_{(1)} = 7064; p>0.05$). Model 2 which adjusted for all
-potentially confounding covariates in \@ref(tab:tabthree) including an
+potentially confounding covariates including an
 interaction between cholestrol level and age was then compared to Model
@@
 \textbf{Intercept} & 102.55 &(88.42,116.69) & $<0.001$\\
-\textbf{Sleep} & 0.75 & (0.50, 1.00) & 0.002 \\
+\textbf{Sleep} &  &  &  \\
 $<6$hrs &  &  &  \\
@@
 
-## Discussion and conclusion
+# Discussion and conclusion
 
-<!-- The current study investigated the intricate interplay between age, sleep, cholestrol levels and systolic blood pressure adjusting for other potentially confounding covariates informed by literature.
-We performed comprehensive model comparisons to explore the association between age, total cholestrol levels, sleep durations and systolic blood pressure.
-We find based on the best model, that work day sleep durations ranging between 6 and 8 hours and more than 8 hours are associated with a decrease in systolic blood pressure relative to the 6 hour reference group.
-Additionally, our study reports substantial effect modification of the effect of total cholestrol levels on systolic blood pressure by age whe adjusted for selected confounders within the study population.
-Previous studies exploring sleep and systolic blood pressure have found similar associations [@su2022association].
--->
-
-The present investigation explote into the settle interplay among age,
-sleep patterns, cholesterol levels, and systolic blood pressure, while
-accounting for various potentially confounding factors gleaned from
-pertinent literature. Our study conducted an exhaustive series of model
+The present investigation explores the subtle interplay between age,
+workday sleep patterns, cholesterol levels, and systolic blood pressure, while
+accounting for various potentially confounding factors informed by literature. We conducted an exhaustive series of model
 comparisons to elucidate the intricate relationship between age, total
-cholesterol levels, sleep durations, and systolic blood pressure.
-
-In light of our findings derived from the best model based on the
+cholesterol levels, workday sleep duration, and systolic blood pressure. In light of our findings derived from the best model based on the
 performance measures in Table \@ref(tab:tabmodel), we observed that
@@
 sleep parameters as influential factors in understanding blood pressure
-dynamics.
+dynamics [@fuchs2020high].
+In this way, the study explores factors that bear substantial weight in understanding disease risk within the aging US population.
+Findings from this study could inform public health policies and interventions towards healthy diet as well as sleep patterns, personalized medicine, and healthy aging promotion.
 
-### Study limitations
-
-Our study encountered a significant challenge with missing data,
-resulting in a notable reduction in our sample size. Despite initially
-having a sample of \$n=19,225\$, the final analysis was conducted on a
-considerably smaller sample of \$n=1,977\$ observations. This
-substantial reduction in the sample size raises concerns about the
-robustness and reliability of our findings. Addressing this issue
-becomes crucial to ensure the validity of our study outcomes.
-
-In response to this challenge, we propose the implementation of multiple
+While our study used a representative sample of the US population, missing data on variables of interest in analyses presented a significant challenge.
+Our results are based on complete case analysis where all individuals with missing data on the variables of interest were excluded.
+While complete case analysis ensures that the analysis is based on comprehensive and complete information for each participant, the notable reduction in sample size has potential to result in low statistical power and bias when the assumption of missing completely at random does not hold [@baraldi2010introduction].
+Future studies can address this challenge through such methods as multiple
 imputation techniques. This approach involves generating multiple
-plausible sets of missing values based on observed data patterns. By
-imputing missing values using this method, we aim to enhance the
-completeness of our data set. This strategy is intended to mitigate
-potential biases resulting from missing data and fortify the reliability
-and accuracy of our analytical outcomes.
+plausible sets of missing values based on observed data patterns [@baraldi2010introduction].
 

### 2023-12-15 data/scrip2.R

@@ -11,11 +11,11 @@ library(ggcorrplot)
 library(kableExtra)
 
 ## Load Data
-dt <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
-##dt <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/dt.csv")
+## dt <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/dt.csv")
+dt <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/dt.csv")
 
-## load the variable name file
-vars <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/vars.csv")
+## load the variable name file [no need to load these names]
+## vars <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/vars.csv")
 ## vars <- read.csv("/Users/aokutse/Desktop/PhD/Fall 2023/PHP2601 Linear Models/SleepBP/data/vars.csv")
 
 dt$cycle <- as.factor(dt$cycle)
@@ -266,7 +266,7 @@ print(significant_sbp)
 ##-------------------------------------
 ## Systolic Blood Pressure
 ## Load variables selected by LASSO
-lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_sbp_min.csv")
+## lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_sbp_min.csv")
 ## lasso.coef.min$x
 names.lasso.min <- c("bmi", "hdl", "total_chol","AST",
                      "creatinine", "diabetes", "citizenship_status",
@@ -295,7 +295,7 @@ predlasso.INT <- predict(modlasso.sbp.min.INT, newdata = dt %>% select(all_of(na
 ## Diastolic Blood Pressure
 ##---------------------------------------------
 ## Load variables selected by LASSO
-lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_dbp_min.csv")
+## lasso.coef.min <- read.csv("C:/Users/monic/OneDrive/Desktop/SleepBP/data/lasso_res_dbp_min.csv")
 ## lasso.coef.min$x
[…]
